### 2024-08-05 tex/writeup.tex
-

\documentclass[12pt,a4paper]{article}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{booktabs}

\title{Investigating the Impact of Macroeconomic Indicators 
on the FTSE 250 Index Using Machine Learning Methods}
\author{Mateen Khan}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
    This is the abstract of your paper. It should be a brief summary of your research.
\end{abstract}

\section{Introduction}

The stock market has played a central role in the economy and political ascent of many European states such as Britain and the Netherlands ever since its inception during the Age of Discovery. Accordingly, any instabilities that have arisen in the stock market throughout history have had far-reaching destabilising effects on the economy; the Mississippi Bubble (1717-1720) was an early example of the dangers that stock markets carry, where the French colonial administration in America went bankrupt as a result of a speculative bubble; with the growing interconnectedness of markets, the stock market as an institution has also contributed significantly to global disasters such as the Great Depression and the Global Financial Crisis of 2008/9. Thus, economic policymakers and administrators must take into careful consideration both the impacts of their macroeconomic policies, and the general trend of the economy to ensure that any instability is corrected before it may have a destabilising effect. In addition, investors and traders are indirectly affected by the fluctuations of economic factors by virtue of their effect on companies. Therefore, examining the determinants of the stock market’s movements is important both for informing economic policymakers and financial services companies. Over time the FTSE 250 has gained a reputation for being a more accurate mirror of the UK economy than its higher-cap counterpart the FTSE 100, mainly because unlike the latter its overseas share of earnings is roughly a half - which is similar to the S\&P 500, instead of the nearly $80\%$ of the FTSE $100$'s earnings being overseas. This is also reflected in the constitutions of the respective indices, with the FTSE 250 historically commensurating more greatly with the structure of the UK economy. Within this context, the study has a twofold aim:
\begin{enumerate}
    \item To investigate which macroeconomic variables influence the stock market price significantly, and quantify the effects of the following selection of relevant macroeconomic indicators on the FTSE 250, chosen as a more representative measure of the UK stock market. An Auto-regressive Delayed Lag (ARDL) model is utilised, in conjunction with an Error Correction Model, over the January 1993 - March 2024 period: interest rate, inflation, USD exchange rate, and M3 money supply. 
    \item To investigate the efficacy of machine learning methods, in particular an RBF neural network (RBF-NN), in predicting stock price movements based on the same macroeconomic indicators. 
\end{enumerate}

\section{Literature Review}

There is a large and growing body of research investigating the impact of macroeconomic variables on the stock market; empirical research supports our selection of macroeconomic determinants:

Louis Bachelier’s \textit{Théorie de la Spéculation} is generally considered to be one of the earliest contributions to the theory of financial markets and randomness.-source- Building on the concepts of Brownian motion, he introduced the random walk theory for stock prices, arguing that future price movements are independent of past movements. 
This laid the foundation for modern portfolio theory in the form of 
Markowitz’s conception of risk and return Sharpe’s Capital Asset Pricing 
Model (1964) - source. Eugene Fama used these foundations to develop his 
Efficient Market Hypothesis in 1970, which argues that asset prices fully 
reflect all available information; Fama outlined three forms of the EMH - 
weak-form, semi-strong form and strong form EMH:
\begin{enumerate}
    \item Weak-form: Future stock prices cannot be predicted based on past prices, implying prices follow a random-walk - Louis Bachelier. Current stock prices fully reflect all information contained in past price movements. This rules out technical analysis as a viable tool in the modelling of stock price fluctuations.
    \item Semi-strong form: Stock prices quickly adjust to reflect all publicly available information including not just price data (which is covered by the weak form EMH)  but also financial statements and reports, macroeconomic changes, political changes, and industry trends. This rules out fundamental analysis as a viable tool in the modelling of stock price fluctuations.
    \item Strong-form: Stock markets are perfectly efficient, stock prices reflect all information, both public and private. This includes all publicly available information covered by the semi-strong form, insider information, and even future events that are not yet known to the public. This implies that it is impossible to enjoy a consistent advantage in predicting stock prices.
\end{enumerate}

The consistent success of investors such as Warren Buffett of Berkshire Hathaway and Jim Simons of Renaissance Technologies in the decades after the development of the EMH have challenged its central argument, and have shown the possibility of achieving consistent returns against the market. Even the creator of the hypothesis conceded in Fama (1991) that many stock returns are in fact predictable from previous values. Furthermore, the empirical research is not in agreement on whether the London Stock Exchange is efficient at any of the three levels. Rounaghi and Nassir Zadeh (2016) gives evidence that the FTSE100 and FTSE250 are efficient over the period 2007 - 2013, and Libberton (2010) also gives evidence that the FTSE100 is weak-form efficient over 1995 - 2007. However, Borges (2010) gives evidence to reject the EMH for both the French and UK stock markets over the period 1993 - 2007; the findings of Ullah and Asghar (2023) support Borges (2010) and reject EMH for the FTSE 100 over the period 2012 - 2020. Additionally, Bhavsar (2015) rejects the EMH for both the FTSE 100 and FTSE 250 over the 13-year period January 2002 - December 2014. Some studies such as Rosini and Shenai (2020) found that the efficiency of the FTSE100 and FTSE250 varied over time, giving support to Andrew Lo’s Adaptive Market Hypothesis. The empirical findings then suggest that it is not impossible to determine, to some degree, movements in the price of stock indices on the LSE using macroeconomic variables.

The Fisher hypothesis, developed by the neoclassicist economist Irving Fisher, states that the nominal interest rate will adjust to accommodate any changes in expected inflation (Frank, Robert; Bernanke, Ben; Antonovics, Kate; Heffetz, Ori. Principles of Macroeconomics. McGraw-Hill. pp. 138–139). The implications of this in the stock market are that firms anticipate inflation correctly and adjust prices accordingly, and that the market is efficient, quickly incorporating inflation into prices. It was tested in the US stock market by Jaffe and Mandelker (1976) when they examined the relationship between monthly inflation and stock returns; they would find a negative relationship between inflation and stock returns over the period January 1951 - December 1971. In support of Jaffe and Mandelker is a large array of empirical research, including but not limited to, Bodie (1976) who found that stocks were not an effective hedge against inflation, as they were affected negatively by inflation; Fama and Schwert (1977) built on Bodie (1976), finding that over a roughly 20 year period stocks did not provide an effective hedge against inflation, and returns were negatively affected by inflation in this period; Mogdiliani and Cohn (1979) sought to root this phenomenon in economic theory, and came up with the Inflation Illusion Hypothesis, which posits that investors fail to correctly account for the impact of inflation on nominal earnings and interest rates, and consequently stock prices when they are valuing stocks, primarily through overestimating the correct discount rate in the Discounted Cash Flow valuation model (link in appendix?). Campbell and Vuolteenaho (2004) found empirical evidence to support this in the US, suggesting that nearly $80\%$ of stock-market time-series variation is caused by the level of inflation. Firth (1979) and Gultekin (1983) found that the relationship between nominal stock returns and inflation in the UK is positive, and that real returns on UK stocks have remained relatively stable even as inflation varied, in support of the generalised Fisher hypothesis. Hassan (2008) later found evidence against this when he investigated the relationship between stock returns and inflation in the UK, finding a significantly positive relationship between the two. Additionally, Cunado et. al. found no evidence of stocks acting as effective hedges over a roughly two-century period in the UK, providing further evidence against the Fisher hypothesis in the UK.

Interest rates are by definition the cost of borrowing. Therefore, higher interest rates increase the cost of borrowing for firms, which means higher expenses for financing operations and capital investments, in turn reducing profitability and adversely affecting public perceptions of future growth prospects; they also present increased borrowing costs for consumers, in turn leading to reduced spending and lower overall demand for the goods and services of companies, and thus profitability. Furthermore, interest rate rises present a higher opportunity cost, this is captured by the flight-to-quality or flight-to-safety phenomenon: investors move their investments from relatively volatile assets such as stocks to ‘safe’ assets such as bonds, which results in a negative correlation between stock and bond returns; owing to the inverse relationship between the bank rate and bond prices and thus positive relation with bond yields because bond yields = annual coupon payments (fixed) / bond price, this suggests that a rise in the bank rate will make bonds a more attractive investment than stocks, which again posits a negative relationship between stock prices and interest rates. - Provide sources for both theoretic claims -. Thus, we would expect interest rates to negatively affect stock prices, and our expectations are satisfied by the empirical research. Alam and Uddin (2009) investigated the impact of the monthly interest rate on the stock market price in Australia, Bangladesh, Canada, Chile, Colombia, Germany, Italy, Jamaica, Japan, Malaysia, Mexico, Philippine, S. Africa, Spain, and Venezuela over the period January 1988 to March 2003 and found a significant inverse relationship between the two. Talla (2013) also found evidence of a negative relationship between the interest rate and Stockholm Stock Exchange price change. Pilinkus and Boguslauskas (2009) found increases in the interest rate led to a decline in the Lithuanian stock market index value. Bernanke and Kuttner (2004)’s findings also indicated the considerable impact of unanticipated interest rate fluctuations on the stock market price. Asgharian et. al (2019) found evidence to support the flight-to-quality phenomenon, particularly in highly-developed capital markets such as the UK’s stock market. Malika (2023) used a Non-Linear ARDL model to find that the interest rate shows a significantly negative long-run relationship with the UK stock market price.

Classical economic theory outlines the relationship between exchange rates and stock prices using flow and portfolio balance models, the former outlining the negative effect of currency depreciations on the cash flow of firms operating internationally, and in turn their stock price, and the latter describing investor propensity in the face of expected currency depreciations to redistribute their investment from domestic to foreign assets, thus decreasing demand and lowering stock prices. In addition, Dornbusch and Fisher (1980) showed that variations in the exchange rate affect future cash flows, which in turn affect the Present Value of a stock and thus stock prices based on the Discounted Cash Flow model. Additionally, Branson et. al (1977) in turn established a bidirectional relation between the exchange rate and stock prices, essentially that the demand for money in each country depends on the performance of assets, including the stock market, in that country’s currency, and conversely the demand for money is determined by real wealth, which is itself partly influenced by stock market returns. Aggarwal (1981) also studied the relationship between exchange rates and stock prices, using the monthly prices of the U.S. capital market and the floating exchange rates of the dollar for the period 1974 to 1978. The results showed a significantly positive correlation between stock prices and the exchange rate. The effect on less developed capital markets was naturally different, with Vanita and Khushboo (2015) reporting a significantly negative relationship between the exchange rate and stock prices of Russia, India and South Africa; however, both supported Branson’s thesis.  More recently, Aydemir and Demirhan (2009) also found evidence to support Branson’s hypothesis in the Turkish Stock Exchange, namely that a bidirectional causal relationship between exchange rate and Turkish stock indices does exist. Furthermore, HT Wong (2022) found evidence of cointegration between the exchange rate and UK, Japanese, Malaysian, Philippine, Singaporean, Korean, German, Hong Kong and Indonesian stock prices. Javangwe and Takawira (2022) further found, through investigation of the South African stock market, that the exchange rate exerts a negative effect on the stock market price in the short run, and a positive effect in the long run.

One of the earliest papers investigating the effect of money supply on the stock market price was Palmer (1970), who examined the effect of M1 money supply on the S$\&$P 425 Industrial Stock Index and found that ‘primary changes (defined as a trend occurring over a period of
months) in the nation's money stock may motivate the private sector to adjust its wealth portfolios in such a manner as to yield predictable movements in the prices of corporate securities,’ positing a definite relation between the money supply and stock market price. Further research by Homa and Jaffe (1971) indicated that the price of a common stock is determined significantly by the risk-free rate of interest, which itself, according to the liquidity preference theory of the monetarist school of economics, a function of money supply, as more money circulating in the economy means more liquidity available for banks, thus affecting the cost of borrowing (interest) - source -. They found that the money supply has a positive impact on the average level of stock prices. This was disputed by Pesando (1974) who claimed that the inability of the model used in this study to produce accurate forecasts of the stock price was sufficient evidence against a relationship between the money supply and stock market prices. Baks and Kramer (1999), supported by Conover et. al (1999) findings, initially found significant evidence of the impact of money supply on stock markets across G-7 countries:  Canada, France, Germany, Italy, Japan, the UK and the US. This empirically supports the economic theory that increases in money supply generally raise the value of personal asset portfolios, which leads individuals to reallocate excess cash to stocks, in turn affecting stock prices. In the face of the drastic shock to the global economy in 2008, many central banks including the USCB and the Bank of England utilised Quantitative Easing (QE) for the first time. As a macroeconomic tool, QE involves the creation of digital money by the central bank to engage in large-scale asset purchasing to stimulate the economy - stock and other asset markets would theoretically be affected immediately by such a policy. Picha (2017) utilised a portfolio balance model to investigate the effect of QE, measured by the M2 money supply, on the S$\&$P 500 using Johansen’s cointegration methodology and found that the money supply indeed exerted a effect on the latter, and the fact that the adjustment to the long-run equilibrium took 6-months suggested that it was quite moderate. Moreover, cointegration between the money supply and S$\&$P 500 over the long-run was established, adding to a large extant literature that supports long-run dependence of stock price on money supply. Included are Mukherjee and Naka (1995), Wongbangpo and Sharma (2002) and Lee (2006) who posit the long-run dependence of Japanese, New Zealand and ASEAN-5 indices respectively on the M1 money supply; furthermore, Bahloul et. al. (2016) establish that M3 money supply significantly influences developed stock indices such as the UK, USA, and Japan over the long-run. Synek (2024) found long term dependence of stock indices such as the FTSE100 and TSX on the M2 money supply.

There have been many studies, excluding those already mentioned, that have investigated the impact of multiple macroeconomic variables on the stock market. Demir (2018), for example, used the ARDL model to investigate the impact of quarterly economic growth, interest rate, real effective exchange rate, crude oil prices, FDI, and FPI on the Borsa Istanbul 100 (BIST-100). The study found that economic growth, real effective exchange rate, portfolio investments, and foreign direct investments raised the stock market index value while the interest rate and crude oil prices negatively affected it. Khan and Khan (2018) dealt with the case of the Karachi Stock Exchange, investigating the effect of the interest rate, inflation rate, M2 money supply, USD exchange rate, economic activity, and exports on the KSE-100. Their findings indicate that the Karachi Stock Exchange, in the long run, was significantly positively affected by the money supply, and significantly negatively affected by the exchange rate, and interest rate; in the short run, all of the variables were found to be insignificant except for the exchange rate, which was again negatively cointegrated with the stock price. In the case of the London Stock Exchange, Asprem (1989) examined the relationship between various European stock indices, asset portfolios and macroeconomic variables in ten European countries, the author’s results showing that employment, inflation, imports and interest rate are inversely related to stock prices. The relationships between stock prices and macroeconomic variables were strongest in Germany, Switzerland and the UK, suggesting that these variables have a significant influence on stock indices in the UK. Additionally, Olomu (2015) examined monthly inflation, industrial production index, M1 money supply, effective exchange rate, and interest rate effect on the FTSE 100 and found that inflation and exchange rate showed a positive relationship with the FTSE100 over the long run, whereas the industrial production index, money supply and interest rate showed negative long-run relationships. Olawale et. al (2014) found that significant long-run relationships existed between FTSE100 returns and the industrial production index, interest rate, and inflation in the UK, and in the short-run, industrial production index, short-term interest rates, and unemployment rates have no significant causal link with FTSE100 returns; they also used M3 money supply as a proxy for Quantitative Easing, finding a significant positive impact on stock returns in the US, but a negative one in the UK. Finally, Marshall and Vasilev (2022) investigated the London Stock Exchange, and found that the FTSE 100 is sensitive to interest rate and exchange rate fluctuations but not sensitive to the employment rate, real GDP, gold prices or public sector budget deficit.

The accelerating development of Machine learning (ML) presents a challenge to the EMH and traditional economic theories through its ability to field advanced computational tools capable of uncovering patterns and inefficiencies in the market that were previously thought to be obscured by the ‘noise’ and randomness posited by the EMH, and beyond the capability of standard econometric models. ML models are able to continually improve over time as they learn from new data, something which regular econometric models are incapable of doing, in particular Neural Networks (NN) have been used extensively for forecasting time-series data, including asset volatility, volume, and prices and future macroeconomic variable movements.

Recurrent Neural Networks (RNN) are a type of neural network especially suited for sequential data due to their feedback loops which enable them to use both current and past inputs, therefore allowing information to persist. This allows RNNs to learn and take into account recent trends when training and making predictions. However, there is an intricacy in how RNNs perform optimisation during training which causes RNNs to lose memory in the long run; of course in many cases we desire the model to retain long-run dependencies, so the Long Short-Term Memory (LSTM) network was developed by Hochreiter and Schmidhuber and introduced in 1997. LSTMs have been used to great effect in many-time series forecasting problems, including stock price prediction, with some studies such as insert and insert achieving insert metric of insert however they are not suitable for small, noisy datasets - such as those involving economic data - and tend to overfit the training data (Foster et. al. 1991).

Radial Basis Function Neural Networks (RBF-NN) are a type of artificial 
neural network that use radial basis functions at each node in the 
activation layer of the network. They were developed in 1988 by 
Broomhead and Lowe at the Royal Signals and Radar Establishment. 
They have since been employed in a variety of contexts such as 
hydrological forecasting (Chang et. al 2001), biomedical engineering 
(Acikkar et. al 2001) and time-series analysis (insert), with success. 
Whilst neural networks trained on small datasets often display unstable 
behaviour during performance (LeBaron and Weigend 1998), The advantage of 
RBF-NNs lies in their strong tolerance for input noise, being able to 
operate effectively in extremely volatile financial time-series 
environments (Cafferate et. al 2019), and their strong ability to 
generalise on such data (Sharkawy 2020) as well as their efficacy on 
relatively small datasets (Kosarac et. al 2022). RBF-NNs have been used 
considerably for stock price forecasting. Esfandyari et. al (2001) 
supports this, using a sample size of 70 and 29 to forecast daily NASDAQ 
index values using an Artificial Neural Network; they achieved test $R^2$
statistics as small as 83$\%$ and large as 98$\%$. Cao et. al (2005) 
used an RBF-NN based on the Optimal Partition Algorithm (OPA) to 
forecast S$\&$P 500 prices, using the index values from January 1993 to 
December 1995. Kumar et al. (2019) use an RBFN, with the Back-propagation 
Algorithm (BPA) used instead to tune parameters; their results demonstrate 
that the RBF-NN outperforms a standard feed-forward neural network with a 
single hidden layer. Che et. al (2014) also found that the RBF-NN, on a 
small training dataset, was able to predict 45 days in advance the value 
of the Shanghai Composite Index, with a mean prediction error of 
1.4387$\%$, suggesting that the RBF-NN is competent in time-series 
forecasting using small training datasets. Aunjum et. al (2021) were 
able to predict crude oil prices using macroeconomic and political news 
using an RBF-NN, with an $R^2$ of 86$\%$. Recently, Abotaleb et.al (2024)
used quarterly data from 2001 to 2023 to model the influence of economic 
policies on the quarterly prices of various stock markets including the 
FTSE100 and Dow Jones price using an RBF-NN, finding that the model 
achieved relative forecast error rates as low as 6.8$\%$ and an error of 
27.2$\%$ in the case of the FTSE100. Thus, we find that the RBF-NN is a 
competent tool in the sphere of financial time-series forecasting, and 
is able to operate effectively within the constraints placed by our 
problem domain and data.

\section{Data}

For all of the research, monthly data between October 1992, when the FTSE250 index first opened, and March 2024 was utilised. 
To account for the fact that the Bank of England’s Monetary Policy Committee does not meet on a monthly basis, the interest rate data was 
adjusted using the following recipe:
\begin{enumerate}
    \item If the interest rate was unchanged through the month, the previous set rate was used.
    \item If the rate changed from $i_1,\ldots,i_n$ on day $d_j$, for $1\leq j\leq n$ of a month with $x$ days, then a weighted average $W_{int}$ was calculated using (1) - insert hyperlink - to more accurately account for the effects of the interest rate over the month. 
    \item The value of INT in would accordingly be $W_{int}$ for those months in which the rate changed. 
\end{enumerate}

\begin{align}
    W_{int} = i_1 \frac{d_{j_1}}{x} + \cdots +i_n\frac{d_{j_2}}{x}
\end{align}

\begin{table}[h!]
    \centering
    \caption{Data sources}
    \begin{tabular}{lll}
        \toprule
        \textbf{Variable} & \textbf{Notation} & \textbf{Source} \\
        \midrule
        FTSE 250 Index & FTSE250 & Yahoo Finance (link) \\
        Consumer Price Index & CPI & Office for National Statistics (ONS) (link) \\
        USD/GBP Exchange Rate (£) & EXCHG & Federal Reserve of St Louis (link)\\
        Interest Rate ($\%$) & INT & Bank of England (link) \\
        Money Supply (M3) (millions £) & M3 & Bank of England (link) \\
        \bottomrule
    \end{tabular}
\end{table}

Table 2 displays some basic descriptive statistics on the raw data. Insert analysis of descriptive statistics and historical behaviour.

\begin{table}[h!]
    \centering
    \caption{Descriptive statistics}
    \begin{tabular}{lllll}
        \toprule
        \textbf{Variable} & \textbf{Mean} & \textbf{SD} &  \textbf{Min.} & \textbf{Max.}\\
        \midrule
        FTSE 250 Index &  1 & 1 & 1 & 1 \\
        CPI Index &  1 & 1 & 1 & 1 \\
        USD/GBP Exchange Rate &  1 & 1 & 1 & 1 \\
        Interest rate &  1 & 1 & 1 & 1 \\
        Money Supply (M3) &  1 & 1 & 1 & 1 \\
        \bottomrule
    \end{tabular}
\end{table}

\section{ARDL Bounds Test}

\section{RBF-NN}

\section{Conclusion}

This is the conclusion.

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}

### 2024-08-05 ardl/adrl-bounds.R

@@ -1,0 +1,125 @@
+library(tseries)
+library(vars)
+library(lmtest)
+library(ARDL)
+library(dynlm)
+
+columns_to_transform <- c("CPI", "INT", "UNEM", "EXCHG", "FTSE.250")
+data[columns_to_transform] <- lapply(data[columns_to_transform], function(x) log(x))
+
+adf.test(data$UNEM)
+dunem = diff(data$UNEM)
+adf.test(dunem)
+
+adf.test(data$INT)
+dint = diff(data$INT)
+adf.test(dint)
+
+adf.test(data$EXCHG)
+dexchg = diff(data$EXCHG)
+adf.test(dexchg)
+
+
+adf.test(data$CPI)
+dcpi = diff(data$CPI)
+adf.test(dcpi)
+
+max_lags <- 9
+results <- list()
+
+for (p in 1:max_lags) {
+  model <- ardl(FTSE.250 ~ UNEM + INT + CPI + EXCHG, data = data, order = c(p,p,p,p,p))
+  
+  aic <- AIC(model)
+  sc <- BIC(model)
+  
+  bg_test <- bgtest(model, order = p)
+  
+  results[[p]] <- list(
+    AIC = aic,
+    SC = sc,
+    Serial_Correlation_p_value = bg_test$p.value
+  )
+}
+
+results_df <- do.call(rbind, lapply(results, as.data.frame))
+print(results_df)
+
+
+
+VARselect(data$FTSE.250)
+VARselect(data$CPI)
+VARselect(data$UNEM)
+VARselect(data$INT)
+VARselect(data$EXCHG)
+
+model3 <- ardl(FTSE.250 ~ UNEM + INT + CPI + EXCHG, data = data, order = c(4,7,7,7,7))
+summary(model)
+
+test3 <- bounds_f_test(model3, case=3)
+test3
+
+model4 <- ardl(FTSE.250 ~ INT + CPI + EXCHG, data = data, order = c(4,7,7,7))
+summary(model)
+
+test4 <- bounds_f_test(model4, case=2)
+test4
+
+install.packages("dLagM")
+library(dLagM)
+
+coefs <- coef(model4)
+
+
+
+# LR_coeff = sum(short-run coefficients of the variable) / (1 - sum(autoregressive coefficients))
+LR_INT <- sum(coefs[grepl("INT", names(coefs))]) / (1 - sum(coefs[grepl("FTSE.250", names(coefs))]))
[…]

### 2024-08-05 ardl/r-code-m3.R

@@ -1,0 +1,133 @@
+library(tseries)
+library(vars)
+library(lmtest)
+library(ARDL)
+library(dynlm)
+
+columns_to_transform <- c("CPI", "INT", "UNEM", "EXCHG", "M2","M3", "FTSE.250")
+monthly.data.m3[columns_to_transform] <- lapply(monthly.data.m3[columns_to_transform], function(x) log(x))
+
+month
+adf.test(data$UNEM)
+dunem = diff(data$UNEM)
+adf.test(dunem)
+
+adf.test(data$INT)
+dint = diff(data$INT)
+adf.test(dint)
+
+adf.test(data$EXCHG)
+dexchg = diff(data$EXCHG)
+adf.test(dexchg)
+
+
+adf.test(data$CPI)
+dcpi = diff(data$CPI)
+adf.test(dcpi)
+
+adf.test(monthly.data.m3$M3)
+dm3 = diff(monthly.data.m3$M3)
+adf.test(dm3)
+
+max_lags <- 9
+results <- list()
+
+for (p in 1:max_lags) {
+  model <- ardl(FTSE.250 ~ INT + CPI + EXCHG + M3, data = monthly.data.m3, order = c(p,p,p,p,p))
+  
+  aic <- AIC(model)
+  sc <- BIC(model)
+  
+  bg_test <- bgtest(model, order = p)
+  
+  results[[p]] <- list(
+    AIC = aic,
+    SC = sc,
+    Serial_Correlation_p_value = bg_test$p.value
+  )
+}
+
+results_df <- do.call(rbind, lapply(results, as.data.frame))
+print(results_df)
+
+
+
+VARselect(data$FTSE.250)
+VARselect(data$CPI)
+VARselect(data$UNEM)
+VARselect(data$INT)
+VARselect(monthly.data.m3$M3)
+VARselect(data$EXCHG)
+
+model3 <- ardl(FTSE.250 ~ INT + CPI + EXCHG + M2, data = data, order = c(5,4,7,6,7))
+summary(model)
+
+test3 <- bounds_f_test(model3, case=2)
+test3
+
+model5 <- ardl(FTSE.250 ~ INT + CPI + EXCHG +M3, data = monthly.data.m3, order = c(5,4,7,7,4))
+summary(model5)
+
+
+test5 <- bounds_f_test(model5, case=2)
+test5
+
+install.packages("dLagM")
+library(dLagM)
+
+coefs <- coef(model5)
+
+
+
[…]

### 2024-08-05 rbf-nn/imq-gridsearch.py

@@ -1,0 +1,81 @@
+import numpy as np
+import pandas as pd
+from sklearn.gaussian_process import GaussianProcessRegressor
+from sklearn.gaussian_process.kernels import Kernel, NormalizedKernelMixin
+from sklearn.metrics import mean_squared_error, r2_score
+from scipy.stats import zscore
+from sklearn.model_selection import GridSearchCV, TimeSeriesSplit
+
+import matplotlib.pyplot as plt
+
+# Custom Generalized Inverse Multiquadric Kernel
+class GeneralizedInverseMultiquadricKernel(Kernel, NormalizedKernelMixin):
+    def __init__(self, length_scale=1.0, sigma=1.0, power=1.0):
+        self.length_scale = length_scale
+        self.sigma = sigma
+        self.power = power
+
+    def __call__(self, X, Y=None):
+        X = np.atleast_2d(X) / self.length_scale
+        if Y is None:
+            Y = X
+        else:
+            Y = np.atleast_2d(Y) / self.length_scale
+        dists = np.sum((X[:, np.newaxis, :] - Y[np.newaxis, :, :]) ** 2, axis=2)
+        return (dists + self.sigma ** 2) ** (-self.power)
+
+    def diag(self, X):
+        return np.ones(X.shape[0])
+
+    def is_stationary(self):
+        return True
+
+# Load and preprocess the data
+data = pd.read_csv('monthly-data-m3.csv')
+data.set_index('month', inplace=True)
+data_normalized = data.apply(zscore)
+data_normalized = data_normalized.drop(columns=['M2'])
+data_normalized['FTSE-250'] = data_normalized['FTSE-250'].diff()
+data_normalized['CPI'] = data_normalized['CPI'].diff()
+data_normalized['INT'] = data_normalized['INT'].diff()
+data_normalized['EXCHG'] = data_normalized['EXCHG'].diff()
+data_normalized['M3'] = data_normalized['M3'].diff()
+data_normalized = data_normalized.iloc[1:]
+
+# Define features and target variable
+X = data_normalized[['CPI', 'INT', 'EXCHG', 'M3']]
+Y = data_normalized['FTSE-250']
+
+# Define the parameter grid for hyperparameter tuning
+param_grid = {
+    'kernel__length_scale': [0.1, 1.0, 5, 10.0],
[…]

### 2024-08-06 ardl/ardl.R
update

@@ -3,53 +3,53 @@ library(vars)
 library(lmtest)
 library(ARDL)
 library(dynlm)
+library(dplyr)
+library(dLagM)
+library(strucchange)
+library(zoo)
 
-columns_to_transform <- c("CPI", "INT", "UNEM", "EXCHG", "M2","M3", "FTSE.250")
+
+columns_to_transform <- c("CPI", "INT", "UNEM", "EXCHG", "M2", "M3", "FTSE.250")
 monthly.data.m3[columns_to_transform] <- lapply(monthly.data.m3[columns_to_transform], function(x) log(x))
 
-month
 adf.test(data$UNEM)
-dunem = diff(data$UNEM)
+dunem <- diff(data$UNEM)
 adf.test(dunem)
 
 adf.test(data$INT)
-dint = diff(data$INT)
+dint <- diff(data$INT)
 adf.test(dint)
 
 adf.test(data$EXCHG)
-dexchg = diff(data$EXCHG)
+dexchg <- diff(data$EXCHG)
 adf.test(dexchg)
 
 
 adf.test(data$CPI)
-dcpi = diff(data$CPI)
+dcpi <- diff(data$CPI)
 adf.test(dcpi)
 
 adf.test(monthly.data.m3$M3)
-dm3 = diff(monthly.data.m3$M3)
+dm3 <- diff(monthly.data.m3$M3)
 adf.test(dm3)
 
 max_lags <- 9
 results <- list()
 
 for (p in 1:max_lags) {
-  model <- ardl(FTSE.250 ~ INT + CPI + EXCHG + M3, data = monthly.data.m3, order = c(p,p,p,p,p))
-  
+  model <- ardl(FTSE.250 ~ INT + CPI + EXCHG + M3, data = monthly.data.m3, order = c(p, p, p, p, p))
+
   aic <- AIC(model)
   sc <- BIC(model)
-  
+
   bg_test <- bgtest(model, order = p)
-  
+
   results[[p]] <- list(
     AIC = aic,
     SC = sc,
     Serial_Correlation_p_value = bg_test$p.value
   )
 }
-
-results_df <- do.call(rbind, lapply(results, as.data.frame))
-print(results_df)
-
 
 
 VARselect(data$FTSE.250)
@@ -59,27 +59,18 @@ VARselect(data$INT)
 VARselect(monthly.data.m3$M3)
 VARselect(data$EXCHG)
 
[…]

### 2024-08-09 tex/paper.tex
writeup update

--- 2024-08-05 tex/writeup.tex
@@
 \usepackage{booktabs}
+\usepackage{multirow}
 
@@
     \item To investigate which macroeconomic variables influence the stock market price significantly, and quantify the effects of the following selection of relevant macroeconomic indicators on the FTSE 250, chosen as a more representative measure of the UK stock market.
-An Auto-regressive Delayed Lag (ARDL) model is utilised, in conjunction with an Error Correction Model, over the January 1993 - March 2024 period: interest rate, inflation, USD exchange rate, and M3 money supply.
+An Auto-regressive Distributed Lag (ARDL) model is utilised, in conjunction with an Error Correction Model, over the January 1993 - March 2024 period: interest rate, inflation, USD exchange rate, and M3 money supply.
     \item To investigate the efficacy of machine learning methods, in particular an RBF neural network (RBF-NN), in predicting stock price movements based on the same macroeconomic indicators. 
@@
         \midrule
-        FTSE 250 Index & FTSE250 & Yahoo Finance (link) \\
+        FTSE 250 Index & FTSE$_{250}$ & Yahoo Finance (link) \\
         Consumer Price Index & CPI & Office for National Statistics (ONS) (link) \\
@@
 
-Table 2 displays some basic descriptive statistics on the raw data. Insert analysis of descriptive statistics and historical behaviour.
+Randomtext
+Randomtext
+Randomtext
+Randomtext
+Randomtext
+RandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+
+Table 2 displays some basic descriptive statistics on the raw data. 
+Insert analysis of descriptive statistics and historical behaviour, 
+brief notes on why the indicators are relevant
 
@@
 
-\section{ARDL Bounds Test}
+Sample size discussion
 
-\section{RBF-NN}
+RandomtextRandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+RandomtextRandomtextRandomtextRandomtext
+\section{ARDL Model}
+
+Based on the empirical research and characteristics of the UK economy, we 
+define the stock market index function as:
+
+\begin{equation}
+    FTSE_{250} = f\biggl(CPI, EXCHG, INT, M3\biggr)
+\end{equation}
+
+The study uses an autoregressive distributed lag (ARDL) model to determine 
+whether a long-term dependence exists for the FTSE250 and our macroeconomic 
+indicators, and if so to quantify the long-run and short-run relationships
+between the former and the latter (Panopoulou and Pittis 2004).
+
+\subsection{Methodology}
+
+Before constructing the ARDL model we first apply a natural logarithmic transform
+on the variables, this is done because in a log-transformed model,
+the coefficients can be interpreted as elasticities. Put simply, 
+a coefficient in a model, where both the dependent and independent variables
+are in logarithmic form, represents the percentage change in the 
+dependent variable for a one percent change in an independent variable, 
+which furnishes us with easily interpretable results. Having done this,
+we must then check at which level the variables are stationary (sidenote) at.
+Stationarity refers to the statistical property of a time-series where 
+its mean, variance, and autocorrelation structure do not change over time; 
+either a variable is stationary at level ($I(0)$), requiring no alteration,
+or it is stationary after taking the first difference ($I(1)$). 
+Non-stationarity leads to spurious regression, where relationships that 
+are not actually present between two time-series are produced. Various
+techniques exist for stationarity testing, however in this study the 
+Augmented Dickey-Fuller (ADF) unit root test, developed by (link), was used.
+
+As the ARDL model involves lagged values of the dependent and independent 
+variables, we are required to determine the optimum lag for the variables 
+that minimises error; this was done using the Schwarz
+Information Criterion.
+
+The implicit form (2) is represented in the ARDL model as:
+
+\begin{align*}
+    \Delta \text{FTSE}_{250_t} &= \alpha_0 + \sum_{i=1}^{p} \alpha_i \Delta \text{FTSE}_{250_{t-i}} + \sum_{j_{1}=0}^{q_1} \beta_{1j_{1}} \Delta \text{CPI}_{t-j_{1}} + \sum_{j_{2}=0}^{q_2} \beta_{2j_{2}} \Delta \text{INT}_{t-j_{2}} \\
+                               & \sum_{j_{3}=0}^{q_3} \beta_{3j_{3}} \Delta \text{EXCHG}_{t-j_{3}} + \sum_{j_{4}=0}^{q_4} \beta_{4j_{4}} \Delta \text{M3}_{t-j_{4}} + \varepsilon_t + \phi_{1} FTSE_{250_{t-1}} \\
+                               & + \phi_{2} CPI_{t-1} + \phi_{3} INT_{t-1} +\phi_4 EXCHG_{t-1} + \phi_5 M3_{t-1}
+\end{align*}
+
+Here, the intercept term, denoting so and so, is given by $\alpha_0$; the 
+remaining $\alpha_i$ terms represent the short-run relationships 
+between the variables, and $\phi_j$ represent the long-run relationships.
+
+In the ARDL bounds test, the null and alternate hypothesis are:
+
+\begin{align*}
+    H_{0}: \phi_i &= \phi_j\\
+    H_{1}: \phi_i &\neq \phi_j
+\end{align*}
+
+for, $i,j = 1,2,3,4,5$, with $i\neq j$. The F-statistic is computed for the joint
+significance of the lagged variables and compared against the critical values
+given by Pesaran et. al. (2001) to determine whether cointegration exists, 
+and at what significance level. If cointegration exists, then we estimate 
+the long-run coefficients $\kappa_i$ by estimating the cointegration equation:
+
+\begin{align}
+    FTSE_{250} = \kappa_0 + \kappa_1 CPI + \kappa_2 INT + \kappa_3 EXCHG + \kappa_4 M3 + \varepsilon 
+\end{align}
+
+Then, an Error Correction Model (ECM) is constructed:
+
+\begin{align*}
+    \Delta \text{FTSE}_{250_t} &= \alpha_0 + \sum_{i=1}^{p} \alpha_i \Delta \text{FTSE}_{250_{t-i}} + \sum_{j_{1}=0}^{q_1} \beta_{1j_{1}} \Delta \text{CPI}_{t-j_{1}} + \sum_{j_{2}=0}^{q_2} \beta_{2j_{2}} \Delta \text{INT}_{t-j_{2}} \\
+                               & \sum_{j_{3}=0}^{q_3} \beta_{3j_{3}} \Delta \text{EXCHG}_{t-j_{3}} + \sum_{j_{4}=0}^{q_4} \beta_{4j_{4}} \Delta \text{M3}_{t-j_{4}} + \gamma\text{ECT}_{t-1} + \varepsilon_t
+\end{align*}
+
+where the error correction term ECT is found by calculating the residual 
+produced by (3):
+
+\begin{align*}
+    ECT_{t-1} = FTSE_{250_{t-1}} - \biggl(\kappa_0 + \kappa_1 CPI + \kappa_2 INT + \kappa_3 EXCHG + \kappa_4 M3 + \varepsilon \biggr)
+\end{align*}
+
+The parameter $\gamma$ describes the speed at which short-run deviations from the long-run equilibrium
+are adjusted. A negative $\gamma$ implies that there exists a correction mechanism that converges the deviation 
+to the long-run equilibrium. To estimate the short-run coefficients, $\alpha_i$ and $\beta_k$, as 
+well as the short-run correction speed $\gamma$ we use OLS regression.
+
+\subsection{Results}
+
+We began by applying a natural logarithmic transform on the variables, and 
+then conducting the ADF unit root test on the data. The null hypothesis of the 
+ADF Unit Root test is that of non-stationarity, and the alternate hypothesis 
+is that the variables are stationary\footnote{All of the variables are tested in logarithmic form. The asterisks *, **, ***, and **** denote statistical significance
+at the 1$\%$, 2.5$\%$, 5$\%$ and 10$\%$ significance levels.}.
+
+\begin{table}[h!]
+    \centering
+    \caption{Augmented Dickey-Fuller (ADF) Unit Root Test}
+    \begin{tabular}{lccc}
+        \toprule
+        \textbf{Variable} & \textbf{Level ADF} & \textbf{First Difference ADF} & \textbf{Stationarity} \\
+        \midrule
+        FTSE$_{250}$ & $-3.2646^{****}$ & $-6.6892^{*}$ & At first difference \\
+        CPI          & $-1.4594$ & $-4.705^{*}$ & At first difference \\
+        EXCHG        & $-2.2467$ & $-7.3855^{*}$ & At first difference \\
+        INT          & $-2.0946$ & $-4.1156^{*}$ & At first difference \\
+        M3           & $-0.70827$ & $-6.3806^{*}$ & At first difference \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+Subsequently, the optimum lag of the variables was calculated using the 
+Schwarz Information Criterion, to obtain the optimum model 
+of ARDL(5, 4, 7, 7, 4). The resulting F-statistic of $4.0798$ was compared against the critical values
+given by Pesaran et. al. (2001), providing evidence of cointegration 
+at the $10\%$, $5\%$ and $2.5\%$ significance levels, but not at the $1\%$ 
+level.
+
+\begin{table}[h!]
+    \centering
+    \caption{ARDL Bounds Test Critical Values}
+    \begin{tabular}{lcccc}
+        \toprule
+        \textbf{Order} & \textbf{10$\%$} & \textbf{5$\%$} & \textbf{2.5$\%$} & \textbf{1$\%$} \\
+        \midrule
+        I(0) & $2.20$ & $2.56$ & $2.88$ & $3.29$ \\
+        I(1) & $3.09$ & $3.49$ & $3.87$  & $4.37$ \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+The long-run and short-run coefficients were then estimated, and presented 
+alonsgide the diagnostic test results which confirm that the model does not
+display autocorrelation, heteroskedascity, non-normality and the cumulative 
+sum of residuals and squared residuals are both stable, indicating that the 
+relationship between the variables has not changed over time. The error
+correction term is also given, and it is negative, as expected.
+
+\begin{table}[h!]
+    \centering
+    \caption{Coefficients Estimates}
+    \begin{tabular}{llll}
+        \toprule
+        \textbf{Variable} & \textbf{Long-Run Coefficient} & \textbf{Variable} & \textbf{Short-Run Coefficient} \\
+        \midrule
+        FTSE$_{250}$ & N/A & FTSE$_{250}$ $(-5)$  & $-0.033089$ \\
+        CPI & $3.157179$ & CPI $(-4)$ & $-0.921235$ \\
+        INT & $-0.1341097$ & INT $(-7)$ & $-0.034572^{****}$\\
+        EXCHG &  $-0.890146$ & EXCHG $(-7)$ & $0.122318$ \\
+        M3 & $-0.1141114$ & M3 $(-4)$ & $-0.527798^{***}$ \\
+         & & ECT $(-1)$ & $-0.002$ \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+\begin{table}[h!]
+    \centering
+    \caption{Diagnostic Tests}
+    \begin{tabular}{llll}
+        \toprule
+        \textbf{Condition} & \textbf{Test} & \textbf{p-value} & \textbf{Result} \\
+        \midrule
+        Autocorrelation & Box-Ljung Test & $0.9926$ & No autocorrelation \\
+        Heteroskedascity & Breusch-Pagan Test & $0.3635$ & Homoskedascity \\
+        Normality & Shapiro-Wilk Test & $3.889$e-$08$ & Normality \\
+        CUSUM & CUSUM Test & N/A & Stable \\
+        CUSUM$^2$ & CUSUM$^2$ Test & N/A & Stable \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+\section{RBF Neural Network}
+
+\subsection{Model Architecture}
+
+The exact interpolation problem, which forms the motivation for RBF-NNs, 
+involves mapping a $d$-dimensional input space $\boldsymbol{\mathrm{x}}$ t
+o a one-dimensional target space $t$. The dataset consists of $N$ input
+ vectors $\boldsymbol{\mathrm{x_n}}$, along with corresponding target 
+ values $t_n$. The objective is to find a function
+  $h(\boldsymbol{\mathrm{x}})$ such that:
+
+\begin{align}
+  h(\boldsymbol{\mathrm{x_n}}) = t_n, \quad \text{for } n = 1, \ldots, N
+\end{align}
+
+The radial basis function approach (Powell, 1987) 
+introduces a set of $N$ basis functions, one for each 
+data point, which are functions of the distance
+$||\boldsymbol{\mathrm{x}} - \boldsymbol{\mathrm{x_n}}||$, where
+the basis function $\phi(\cdot)$ is some non-linear function. The output of the mapping is then expressed as a linear combination of these basis functions:
+
+\begin{align}
+    h(\boldsymbol{\mathrm{x}}) = \sum_{n=1}^{N} w_n \phi(||\boldsymbol{\mathrm{x}} - \boldsymbol{\mathrm{x_n}}||)
+\end{align}
+
+By introducing several modifications to the exact interpolation 
+procedure, we obtain the RBF-NN
+model first constructed in (Broomhead and Lowe, 1988). 
+The RBF-NN mapping takes the following form:
+
+\begin{align}
+    y(\boldsymbol{\mathrm{x}}) =  \sum_{j=1}^{M} w_{j} \phi_j(\boldsymbol{\mathrm{x}}) + w_{0}
+\end{align}
+
+where the radial basis function $\phi$ is the Inverse Multi-Quadric:
+
+\[
+\phi(r) = \left(r^2 + \sigma^2 \right)^{-\beta}
+\]
+
+where $r = \|\boldsymbol{\mathrm{x}} - \boldsymbol{\mathrm{c}}\|$ is the Euclidean distance between the input vector 
+\(\boldsymbol{\mathrm{x}}\) and the center of the basis function 
+\(\boldsymbol{\mathrm{c}}\). The constant $\sigma$ represents the
+spread of the basis function at each centre.
+-
+
+An image of the network should be added, include courtesy of Bishop.
+
+\subsection{Methodology}
+
+The methodology should include a reference to the ARDL test
+so that we are linking everythnig together, and then go through why we must normalise the data for the RBF
+- one of the papers in my bookmarks talks about this. After this it should 
+go through how this was implemented in Python, and how we calculated the
+optimum hyperparameters, and which test error metric we utilised. Also explain
+the long-term and short-term prediction part.
+
+\subsection{Results}
+
+Present the results for the long term and short
 
@@
 
-This is the conclusion.
-
+What did this study achieve?
+How may it inform policymakers and investors?
+Areas for future work to build upon.
 \bibliographystyle{plainnat}

### 2024-08-11 tex/paper.tex
update

@@
 Moreover, cointegration between the money supply and S$\&$P 500 over the long-run was established, adding to a large extant literature that supports long-run dependence of stock price on money supply.
+In the same study, it was found that there did not exist a significant relationship between the money supply and the UK stock market.
 Included are Mukherjee and Naka (1995), Wongbangpo and Sharma (2002) and Lee (2006) who posit the long-run dependence of Japanese, New Zealand and ASEAN-5 indices respectively on the M1 money supply; furthermore, Bahloul et.
@@
 
-For all of the research, monthly data between October 1992, when the FTSE250 index first opened, and March 2024 was utilised. 
-To account for the fact that the Bank of England’s Monetary Policy Committee does not meet on a monthly basis, the interest rate data was 
-adjusted using the following recipe:
+For the scope of this research, monthly data spanning from October 1992—marking the inception of the FTSE 250 index—through to March 2024 has been utilised, giving us 
+altogether a dataset of 379 observations.
+
+During the data collection process, it was necessary to deal with an issue raised by the fact that the Bank of England’s Monetary Policy Committee (MPC) does not convene on a monthly basis, and are not obligated to announce a rate on a regular basis.
+Thus, it was imperative to address the resulting irregularities in the interest rate data; to ensure a more accurate representation of interest rate fluctuations over time, the data was adjusted using the following algorithm:
 \begin{enumerate}
@@
     \item If the rate changed from $i_1,\ldots,i_n$ on day $d_j$, for $1\leq j\leq n$ of a month with $x$ days, then a weighted average $W_{int}$ was calculated using (1) - insert hyperlink - to more accurately account for the effects of the interest rate over the month.
-    \item The value of INT in would accordingly be $W_{int}$ for those months in which the rate changed. 
+    \item The value of INT would accordingly be $W_{int}$ for those months in which the rate changed. 
 \end{enumerate}
@@
         \midrule
-        FTSE 250 Index & FTSE$_{250}$ & Yahoo Finance (link) \\
-        Consumer Price Index & CPI & Office for National Statistics (ONS) (link) \\
-        USD/GBP Exchange Rate (£) & EXCHG & Federal Reserve of St Louis (link)\\
-        Interest Rate ($\%$) & INT & Bank of England (link) \\
-        Money Supply (M3) (millions £) & M3 & Bank of England (link) \\
+        FTSE 250 Index & FTSE$_{250}$ & Yahoo Finance \\
+        Consumer Price Index & CPI & Office for National Statistics (ONS) \\
+        USD/GBP Exchange Rate (£) & EXCHG & Federal Reserve of St Louis \\
+        Interest Rate ($\%$) & INT & Bank of England \\
+        Money Supply (M3) (millions £) & M3 & Bank of England \\
         \bottomrule
@@
 
-Randomtext
-Randomtext
-Randomtext
-Randomtext
-Randomtext
-RandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-
-Table 2 displays some basic descriptive statistics on the raw data. 
-Insert analysis of descriptive statistics and historical behaviour, 
-brief notes on why the indicators are relevant
+Table 1 presents the sources of the data; the original datasets are available on request. 
+The precedent for the selection of these variables as some of the most important 
+factors in the fluctuation of the stock market has already been set in the literature, however, 
+a brief explanation of why the M3 measure of the money supply was chosen is 
+required. The M3 measure includes a broader range of liquid assets 
+compared to M2, encompassing all the components of M2 plus large time 
+deposits, institutional money market funds, and other larger liquid assets
+making it a more comprehensive indicator of the money supply's impact on economic activity and financial markets.
 
@@
         \midrule
-        FTSE 250 Index &  1 & 1 & 1 & 1 \\
-        CPI Index &  1 & 1 & 1 & 1 \\
-        USD/GBP Exchange Rate &  1 & 1 & 1 & 1 \\
-        Interest rate &  1 & 1 & 1 & 1 \\
-        Money Supply (M3) &  1 & 1 & 1 & 1 \\
+        FTSE 250 Index &  11,013 & 6,162 & 2,520 & 24,102 \\
+        CPI Index &  88.4 & 18.0 & 62.8 & 134 \\
+        USD/GBP Exchange Rate (£) &  0.659 & 0.085 & 0.483 & 0.883 \\
+        Interest rate ($\%$) &  3.21 & 2.52 & 0.1 & 8.4 \\
+        Money Supply (M3) (millions £) &  1,861,958 & 971,112 & 494,181 & 3,704,873 \\
         \bottomrule
@@
 
-Sample size discussion
+Over the period covered by our dataset, the FTSE 250 index has demonstrated substantial volatility with an average value of 11,013.27, fluctuating between 2,519.9 and 24,102.2. 
+This volatility reflects significant market shifts driven by events such as the 
+Dotcom bubble burst in 2000, the global financial crisis of 2008, and more recent geopolitical uncertainties such as Brexit, the COVID-19 pandemic, and the Russia-Ukraine conflict. However, 
+despite these fluctuations, the index has shown considerable growth, 
+indicative of a generally upward trend in the UK stock market over the 
+long term.
 
-RandomtextRandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
-RandomtextRandomtextRandomtextRandomtext
+The USD/GBP exchange rate has exhibited moderate variability, 
+ranging from 0.4831 to 0.8834, suggesting periods of both 
+appreciation and depreciation of the Sterling. The little variability 
+there is can be
+attributed to major events such as the 2008 financial crisis and Brexit, 
+which both greatly affected the value of the Pound.
+
+The Consumer Price Index (CPI), with an average of 88.36, has 
+experienced significant inflationary fluctuations, ranging from 
+62.8 to 133.5. These fluctuations highlight varying price levels and 
+reflect periods of both high inflation, particularly in the late 2000s 
+near the global financial crisis, and also during 2021-22 as the Russia-Ukraine
+conflict destabilised energy and food markets, immediately after the devastation 
+of supply chains by the COVID-19 pandemic; and also low inflation in the 
+years of austerity in the UK and the global recession that onset during 
+the COVID19 pandemic.
+
+Interest rates, averaging $3.21\%$, have varied widely from $0.1\%$ to $8.4\%$, with a standard deviation of $2.52\%$, reflecting the 
+increasing role of monetary policy in the UK. The Bank of England was forced to adjust interest rates 
+many times in response to worsening economic conditions, examples being the steep rate cuts during the 2008 crisis to stimulate the economy and the gradual rate hikes in response to subsequent inflationary pressures.
+
+The money supply (M3), averaging 1,861,958.47, has shown substantial variability, ranging from 494,181 to 3,704,873, 
+in large part influenced by the various Quantitative Easing (QE) programs implemented by the Bank of England 
+in response to the aforementioned crises. These programs aimed to inject liquidity into the economy, thus inflating the overall money supply.
+
 \section{ARDL Model}
@@
         CPI & $3.157179$ & CPI $(-4)$ & $-0.921235$ \\
-        INT & $-0.1341097$ & INT $(-7)$ & $-0.034572^{****}$\\
-        EXCHG &  $-0.890146$ & EXCHG $(-7)$ & $0.122318$ \\
+        INT & $-0.1341097^{**}$ & INT $(-7)$ & $-0.034572^{****}$\\
+        EXCHG &  $-0.890146^{***}$ & EXCHG $(-7)$ & $0.122318$ \\
         M3 & $-0.1141114$ & M3 $(-4)$ & $-0.527798^{***}$ \\
-         & & ECT $(-1)$ & $-0.002$ \\
+        Intercept & $-3.447767$ & ECT $(-1)$ & $-0.16^{**}$ \\
         \bottomrule
@@
 \end{table}
+
+The long term coefficients show that the interest rate and exchange rate
+interest rate have significant long term effects on the stock market price
+while money supply and inflation are statistically insignificant. Neifar (2023) 
+proposes that this is because deflation, that is the negative movements in the 
+CPI, is a statistically significant factor in the movement of the UK stock 
+market, but positive movements (inflation) are not. Nevertheless, the positive
+coefficient of 3.16, and the negative short-term adjustment, is in line with the expectation from the literature that 
+inflation exerts a positive effect on the stock market in the long-term, but a negative 
+one in the short-run.
+
+The insignificant long-term effect of the money supply 
+was noted by Picha (2017) and appears to be confirmed here. There are a few explanations
+for why this may be.
+
+Unsurprisingly, the USD/GBP exchange rate is significantly related to the FTSE 250's movements. It 
+should not be forgotten that, although it is considered to be a more domestic indicator of the LSE, 
+at times $50\%$ of the FTSE 250's earnings have come from abroad, indicating that 
+the value of the Pound has an important impact on this stock market. As the USD has been 
+the global reserve currency for many decades, the depreciation of the Pound 
+against it by $1\%$ has an impact of $-0.89\%$ on the FTSE 250. The findings 
+are in line with Neifar (2023), Liang and Willett (2015), and Khan and Khan (2018).
+
+The interest rate was also found to have significant negative long-run effects on the 
+stock market, affirming the findings of the extant literature ().
+The mechanisms by which this occurs were discussed earlier in the literature review.
+
+Turning to the short-run dynamics, we find that the error correction term 
+is negative and statistically significant ensuring that long-run equilibrium 
+can be achieved. The adjustment speed of $0.16$ indicates that 
+roughly $16\%$ of the fluctuation in the previous month's price, caused by the 
+macroeconomic variable movements, is adjusted back to the long-run equilibrium in the current month.
+However, only increases in the interest rate and money supply are able to 
+significantly contribute towards this correction, with $1\%$ increases in the former 
+leading to a $-0.03\%$ change in the FTSE250 price 7-months later, and in the case of the latter
+a $-0.53\%$ change after a 4-month lag.
 
@@
 spread of the basis function at each centre.
--
-
-An image of the network should be added, include courtesy of Bishop.
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{bishopnn.png}
+    \caption{Courtesy of C. Bishop (1995)}
+    \label{fig:bishopnn}
+\end{figure}
 
@@
 
-The methodology should include a reference to the ARDL test
-so that we are linking everythnig together, and then go through why we must normalise the data for the RBF
-- one of the papers in my bookmarks talks about this. After this it should 
-go through how this was implemented in Python, and how we calculated the
-optimum hyperparameters, and which test error metric we utilised. Also explain
-the long-term and short-term prediction part.
+The ARDL Bounds Test confirmed that our variable selection was sensible by virtue
+of its establishment of cointegration. The RBF-NN, it is hoped, will be able
+to capture intricate non-linear patterns in the data that the econometric model
+was unable to, and accordingly be able to predict with accuracy the future price
+movements. The RBF-NN was utilised for both short-term and long-term 
+predictions. For short-term forecasting, 
+the model was tested over a 5-month horizon, based on the recommendation of the
+ARDL model, while the long-term predictions spanned 57 months. 
+We differenced the data based on the recommendations of the 
+unit root tests conducted earlier to achieve stationarity.
+
+We began by normalising the data, using a z-score normalisation. This is because
+neural networks work best with data
+that is between a specified range; it ensures that the data is roughly uniformly
+distributed between the network inputs and outputs 
+(Mendelsohn, 1993, Preprocessing Data for Neural Networks). Following this 
+we had to program the RBF-NN into Python. As no specific RBF package exists
+in Python, we used the kernel regression approach outlined by 
+Bishop, using the GaussianProcessRegressor (GPR) package in Python.
+
+\subsubsection{The Gaussian Process Regressor}
+
+The Scikit GaussianProcessRegressor is based on the 
+standard linear regression 
+model, equipped with Gaussian noise,
+
+\begin{align} 
+    f(x) = x^{T}w, \quad{y = f(x) + \varepsilon}
+\end{align}
+
+where $y$ differs from the function values $f(x)$ by noise which is assumed
+to follow an independent and identically distributed Gaussian distribution
+$\mathcal{N}_1$
+\begin{align*}
+    \varepsilon \sim \mathcal{N}_1 \bigl(0, \sigma_{n}^2\bigr)
+\end{align*}
+This assumption together with the model produces the 
+probability density of the observations given the inputs, 
+which is factored over cases in the training set, 
+permissible because of the independence assumption,
+
+\begin{align}
+    p(y|X,w) = \prod_{i=1}^{n} p(y_i|x_i ,w) = \mathcal{N}\bigl(X^T, \sigma_{n}^2 I\bigr)
+\end{align}
+
+In the Bayesian formalism we need to specify a prior over the parameters, expressing our beliefs about the prior
+parameters before we look at the observations. A covariance function or kernel
+is specified according to the characteristics of the data; in our model this 
+is the generalised inverse multiquadric
+
+$$ k(x,x') = \biggl( ||x-x'|| + \sigma^2\biggr)^{-\beta} $$
+and used to construct the covariance matrix $K$, which along with 
+a mean of $0$ forms the distribution of the the weights 
+$$ w \sim \mathcal{N}\bigl(0, K\bigr)$$
+By Bayes' Theorem,
+
+\begin{align}
+    p(w|y, X) = \frac{p(y|X,w) p(w)}{p(y|X)}
+\end{align}
+Using this, the posterior distribution is then obtained
+
+\begin{align}
+    p(w|y, X) \sim \mathcal{N}\bigl(\frac{A^{-1}Xy}{\sigma_{n}^{2}}, A^{-1}\bigr)
+\end{align}
+where the covariance matrix $A^{-1} = K^{-1} + \frac{XX^T}{\sigma_{n}^2}$.
+To make predictions on a test case all possible $w$
+values, weighted by their posterior probability, are averaged to obtain the
+predictive distribution, upon which predicted outputs are computed.
+
+\subsubsection{The Network}
+
+A $4$-dimensional input space $X$
+passes inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
+and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \), through the first
+of two hidden layers, in which the covariance function $(x,x')$
+is applied, the goal being to determine the distribution of the weights. Prior to this,
+the optimal hyperparameters in the covariance function, which plays an analogous
+role to the RBF basis function, are calculated using GridSearch Cross Validation; these hyperparameters are:
+\begin{enumerate}
+    \item $\alpha$: a regularisation parameter which represents a small quantity
+    that is added to the diagonals of the covariance matrix to ensure it is 
+    invertible. 
+    \item $\beta$: the power in the covariance function, which controls the shape of the function.
+    \item $\sigma$: a parameter which controls how much of the noise in the data the kernel accounts for; naturally, 
+    it moderates the risk of overfitting and underfitting.
+    \item Length scale: a parameter that controls the influence - analogous to 
+    the width each RBF neuron possesses - that each kernel (analogous to node) is to have.
+\end{enumerate}
+following this the model uses the training observation $(y,X)$ to find
+the posterior predictive distribution, and with it the predicted output, after 
+averaging over all possible $w$, weighted by their likelihood according to (10).
 
@@
 
-Present the results for the long term and short
+The model returned a mean absolute error (MAE)  MAE $=0.16$ and 
+$R^2 = 9\%$ for the long term prediction.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{short-term.png}
+    \caption{}
+    \label{fig:}
+\end{figure}
+
+The interpretation of this is that, historically, the variations in the 
+Consumer Price Index, interest rate, USD/GBP exchange rate, and M3 money 
+supply have explained or accounted for $9\%$ of the fluctuations of the
+FTSE250 value; in the context of a phenomenon as complex as the stock market,
+which is influenced by an inumerable quantity of factors such as investor sentiment
+- inherently irrational, 
+political instability, geopolitical shocks, corporate governance, and financial regulation, this is 
+unsurprising. Furthermore, the selection of economic variables in this study
+is by no means exhaustive; commodity prices, international trade dynamics,
+economic growth, and many other variables also exert a non-insignificant 
+effect on the stock market.
+
+The short-term prediction yielded MAE = $0.07$ and 
+$R^2 = 61\%$, for a 5-month time horizon.
+The significantly better performance of the model in the short-run
+indicates that the selected macroeconomic variables possesses
+a stronger predictive power over shorter periods. This was highlighted 
+earlier in the ARDL model, with the interest rate and money supply significantly
+impacting the FTSE250 price. In the short-term, which we define as 5 months, 
+as per the optimum lag for the FTSE250 in the ARDL model, the market movements
+are more closely tied to recent deviations in the economy, because theoretically
+the short-run is not exposed to the various unforeseen external shocks and structural changes
+that dilute the relationship between these variables and the stock market in the long-run.
 
@@
 
-What did this study achieve?
-How may it inform policymakers and investors?
-Areas for future work to build upon.
+What did this study achieve: it tested the findings of previous research on the 
+UK stock market, via the FTSE250, and was able to...
+
+How may it inform policymakers and investors: the effect of the 
+Areas for future work to build upon: consider daily data with some kind of 
+stochastic interpolation to get daily data and try other networks. 
+Other variables. Autoregress on 
+FTSE itself in rbf, as this was not in scope of the project.
+
 \bibliographystyle{plainnat}

### 2024-08-11 ardl/ardl.R

@@ -29,9 +29,13 @@ adf.test(data$CPI)
 dcpi <- diff(data$CPI)
 adf.test(dcpi)
 
-adf.test(monthly.data.m3$M3)
-dm3 <- diff(monthly.data.m3$M3)
+adf.test(data$M3)
+dm3 <- diff(data$M3)
 adf.test(dm3)
+
+adf.test(data$FTSE.250)
+dftse <- diff(data$FTSE.250)
+adf.test(dftse)
 
 max_lags <- 9
 results <- list()
@@ -87,9 +91,19 @@ monthly.data.m3$CointEq <- with(monthly.data.m3, FTSE.250 -
     long_run_intercept))
 
 
-ecm_model <- lm(diff(FTSE.250) ~ diff(INT) + diff(CPI) + diff(EXCHG) + diff(M3) + na.omit(lag(CointEq, 1)), data = monthly.data.m3)
+
+monthly.data.m3$ECT <- monthly.data.m3$FTSE.250 - monthly.data.m3$CointEq
+
+ecm_model <- dynlm(
+  diff(FTSE.250) ~ 
+    lag(diff(FTSE.250), 5) + lag(diff(CPI), 4) + lag(diff(INT), 7)  + lag(diff(EXCHG), 7) + lag(diff(M3), 4) + na.omit(lag(ECT, 1)), data = monthly.data.m3)
 summary(ecm_model)
 
+
+LR_INT
+LR_CPI
+LR_EXCHG
+LR_M3
 
 residuals <- residuals(model5)
 fitted_values <- fitted(model5)

### 2024-08-12 tex/paper.tex
introduction

@@
 \usepackage{multirow}
+\usepackage[numbers]{natbib}
+\usepackage{natbib}
+\usepackage{parskip}
+\usepackage{xcolor} \color{black}
 
@@
 \author{Mateen Khan}
-\date{\today}
+\date{}
 
@@
 \begin{abstract}
-    This is the abstract of your paper. It should be a brief summary of your research.
+    Investment decisions and economic policy are highly influenced by 
+    macroeconomic variables and the stock market. Econometric models
+    have been the predominant tools in this context. The aim of this study is to analyse the impacts of some especially important macroeconomic
+    indicators on the FTSE250 
+    (Financial Times Stock Exchange 250 Index), using monthly data 
+    from  October 1993 to April 2024 using an ARDL model, obtaining 
+    quantifiable estimates for the long-term and short-term impacts of the variables
+    on the FTSE250. The efficacy of Radial Basis Function 
+    Neural Networks (RBF-NN) in this problem domain is then investigated. We find that 
+    the RBF-NN is a competent short-term forecasting tool, and is able to 
+    provide an alternative approach to standard econometric models in measuring the long-term influence of
+    macroeconomic variables on the stock market.
 \end{abstract}
@@
 
-The stock market has played a central role in the economy and political ascent of many European states such as Britain and the Netherlands ever since its inception during the Age of Discovery.
-Accordingly, any instabilities that have arisen in the stock market throughout history have had far-reaching destabilising effects on the economy; the Mississippi Bubble (1717-1720) was an early example of the dangers that stock markets carry, where the French colonial administration in America went bankrupt as a result of a speculative bubble; with the growing interconnectedness of markets, the stock market as an institution has also contributed significantly to global disasters such as the Great Depression and the Global Financial Crisis of 2008/9.
-Thus, economic policymakers and administrators must take into careful consideration both the impacts of their macroeconomic policies, and the general trend of the economy to ensure that any instability is corrected before it may have a destabilising effect.
-In addition, investors and traders are indirectly affected by the fluctuations of economic factors by virtue of their effect on companies.
-Therefore, examining the determinants of the stock market’s movements is important both for informing economic policymakers and financial services companies.
-Over time the FTSE 250 has gained a reputation for being a more accurate mirror of the UK economy than its higher-cap counterpart the FTSE 100, mainly because unlike the latter its overseas share of earnings is roughly a half - which is similar to the S\&P 500, instead of the nearly $80\%$ of the FTSE $100$'s earnings being overseas.
-This is also reflected in the constitutions of the respective indices, with the FTSE 250 historically commensurating more greatly with the structure of the UK economy.
-Within this context, the study has a twofold aim:
+The stock market has played a central role in the economy and 
+political ascent of many states ever since its inception. 
+Accordingly, any instabilities that have arisen in the stock market 
+throughout history have had far-reaching destabilising effects on the 
+economy; the Mississippi Bubble (1717-1720) was an early example of the 
+dangers that stock markets carry, where the French colonial administration 
+in America went bankrupt as a result of a speculative bubble. Wth the
+growing interconnectedness of markets, the stock market as an institution 
+has also contributed significantly to global disasters such as the Great 
+Depression and the Global Financial Crisis of 2008/9. Thus, economic 
+policymakers and administrators must take into careful consideration 
+both the impacts of their macroeconomic policies on the stock market.
+
+Investors and traders are also affected by the 
+fluctuations of economic factors due to their effect on firms.
+The ability to forecast market trends allows 
+investors to optimise their trading strategies, allowing them to make
+informed decisions about when to buy or sell assets. 
+Additionally, in a globalised economy, the interconnectedness of 
+markets means that events in one region may have significant effects worldwide, 
+making it even more critical for investors to stay ahead of potential market movements.
+The traditional methods that have been leveraged in this context have been econometric, 
+however machine learning methods have emerged as powerful alternatives capable
+of capturing complex relationships in noisy financial data, and thereby providing more
+accurate forecasts.
+
+We utilise the FTSE 250 as our measure of the UK stock market, ecause unlike the 
+FTSE100 its overseas share of earnings is 
+roughly $50\%$, similar to the S\&P 500. In contrast, nearly $80\%$ of the FTSE $100$'s earnings 
+are from overseas. This is reflected in the constitution of the respective indices, with 
+the FTSE 250 historically resembling the structure 
+of the UK economy more closely.
+
+Thus, the study has a twofold aim:
 \begin{enumerate}
@@
 An Auto-regressive Distributed Lag (ARDL) model is utilised, in conjunction with an Error Correction Model, over the January 1993 - March 2024 period: interest rate, inflation, USD exchange rate, and M3 money supply.
-    \item To investigate the efficacy of machine learning methods, in particular an RBF neural network (RBF-NN), in predicting stock price movements based on the same macroeconomic indicators. 
+    \item To investigate the efficacy of machine learning methods, in particular a radial basis function neural network, in predicting stock price movements based on the same macroeconomic indicators.
 \end{enumerate}
@@
 
-There is a large and growing body of research investigating the impact of macroeconomic variables on the stock market; empirical research supports our selection of macroeconomic determinants:
-
-Louis Bachelier’s \textit{Théorie de la Spéculation} is generally considered to be one of the earliest contributions to the theory of financial markets and randomness.-source- Building on the concepts of Brownian motion, he introduced the random walk theory for stock prices, arguing that future price movements are independent of past movements.
-This laid the foundation for modern portfolio theory in the form of 
-Markowitz’s conception of risk and return Sharpe’s Capital Asset Pricing 
-Model (1964) - source. Eugene Fama used these foundations to develop his 
-Efficient Market Hypothesis in 1970, which argues that asset prices fully 
+The Efficient Market Hypothesis in 1970 \cite{fama1970} argues that asset prices fully 
 reflect all available information; Fama outlined three forms of the EMH - 
@@
 
-The consistent success of investors such as Warren Buffett of Berkshire Hathaway and Jim Simons of Renaissance Technologies in the decades after the development of the EMH have challenged its central argument, and have shown the possibility of achieving consistent returns against the market.
-Even the creator of the hypothesis conceded in Fama (1991) that many stock returns are in fact predictable from previous values.
-Furthermore, the empirical research is not in agreement on whether the London Stock Exchange is efficient at any of the three levels.
-Rounaghi and Nassir Zadeh (2016) gives evidence that the FTSE100 and FTSE250 are efficient over the period 2007 - 2013, and Libberton (2010) also gives evidence that the FTSE100 is weak-form efficient over 1995 - 2007.
+The consistent success of investors such as Warren Buffett of 
+Berkshire Hathaway and Jim Simons of Renaissance Technologies 
+in the decades after the development of the EMH have challenged 
+its central argument, and have shown the possibility of achieving 
+consistent returns against the market. Even the creator of the 
+hypothesis conceded that many stock returns are in fact predictable from 
+previous values \cite{fama1991}. Furthermore, the empirical research is not in a
+greement on whether the London Stock Exchange is efficient at any of the three levels. 
+Rounaghi and Zadeh (2016) gives evidence that the FTSE100 and FTSE250 are efficient over the period 2007 - 2013, and Libberton (2010) also gives evidence that the FTSE100 is weak-form efficient over 1995 - 2007.
 However, Borges (2010) gives evidence to reject the EMH for both the French and UK stock markets over the period 1993 - 2007; the findings of Ullah and Asghar (2023) support Borges (2010) and reject EMH for the FTSE 100 over the period 2012 - 2020.
@@
 
-In the Bayesian formalism we need to specify a prior over the parameters, expressing our beliefs about the prior
-parameters before we look at the observations. A covariance function or kernel
+As this is a Bayesian model we need to specify a \textit{prior distribution} over the parameters, 
+expressing our beliefs about the prior
+parameters before we examine the observations. A covariance function or kernel
 is specified according to the characteristics of the data; in our model this 
@@
 and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \), through the first
-of two hidden layers, in which the covariance function $(x,x')$
+layer, in which the covariance function $(x,x')$
 is applied, the goal being to determine the distribution of the weights. Prior to this,
@@
 
-The model returned a mean absolute error (MAE)  MAE $=0.16$ and 
-$R^2 = 9\%$ for the long term prediction.
+The model returned a mean absolute error of MAE $=0.16$ and 
+$R^2 = 9\%$ for the long term prediction, outperforming the ARDL-Error Correction model by 
+$2.4\%$.
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{short-term.png}
-    \caption{}
-    \label{fig:}
+    \includegraphics[width=0.8\textwidth]{long-term-monthly.png}
+    \caption{Long-term predicted vs monthly real FTSE250 price changes}
+    \label{fig:lmonthly}
 \end{figure}
@@
 earlier in the ARDL model, with the interest rate and money supply significantly
-impacting the FTSE250 price. In the short-term, which we define as 5 months, 
+impacting the FTSE250 price.
+
+In the short-term, which we define as 5 months, 
 as per the optimum lag for the FTSE250 in the ARDL model, the market movements
@@
 the short-run is not exposed to the various unforeseen external shocks and structural changes
-that dilute the relationship between these variables and the stock market in the long-run.
+that dilute the relationship between these variables and the stock market in the 
+long-run. Essentially, as we stretch our prediction window, we 
+introduce more uncertainty and unknowns, which translates into a 
+lower predictive quality.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{long-term-daily.png}
+    \caption{Long-term predicted vs daily real FTSE250 price changes}
+    \label{fig:ldaily}
+\end{figure}
+
+Furthermore, as the time horizon extends into the long-run, fundamental factors 
+begin to matter a lot more. The compound effect of earnings, cash flows and
+managerial efficiencies begins to outweigh the 
+influence of fluctuating macroeconomic sentiment. Thus, in the long-term
+we see a different picture altogether.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{short-term-monthly.png}
+    \caption{Short-term predicted vs monthly real FTSE250 price changes}
+    \label{fig:smonthly}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{short-term-daily.png}
+    \caption{Short-term predicted vs daily real FTSE250 price changes}
+    \label{fig:sdaily}
+\end{figure}
 
@@
 
-What did this study achieve: it tested the findings of previous research on the 
-UK stock market, via the FTSE250, and was able to...
+This study first tested the findings of previous econometric research on the 
+UK stock market, relating the impact of some especially
+relevant macroeconomic factors on the FTSE250.
 
-How may it inform policymakers and investors: the effect of the 
-Areas for future work to build upon: consider daily data with some kind of 
-stochastic interpolation to get daily data and try other networks. 
-Other variables. Autoregress on 
-FTSE itself in rbf, as this was not in scope of the project.
+Stock markets provide important information on the economy, 
+and economic policymakers are vigilant in tracking the fluctuations 
+in these markets, and try to forecast the 
+forthcoming fluctuations based on policy changes that affect the macroeconomy.
+Within this context, the study examined the 
+impacts of the interest rate, USD/GBP
+exchange rate, the consumer price index, and 
+M3 money supply market activity on
+the FTSE250 index over the October 1993 – April 2024 period.
+The results suggest a cointegrated long-run
+relationship between the macroeconomic variables and the stock market index. 
+The long-run coefficients imply that
+increases in the interest rate and depreciation in the exchange rate
+significantly depress the FTSE250 value. 
+Policymakers dealing with forecasting the UK stock markets
+should consider the factors in the study, and to raise the stock
+market's value, a lower interest rate regime should be adopted and 
+the Pound should be made more competitive.
+
+The the study also investigated the efficacy of machine learning methods in 
+the field of stock-econometric forecasting, and returned good results, providing investors
+with an accurate short-term forecast tool, and also providing economists
+with a quantitative non-econometric measure of the 
+long-term influence of the selected macroeconomic 
+variables on the FTSE250 that outperforms a standard econometric model.
+
+Further research should endeavour to include other macroeconomic variables such as 
+foreign direct and portfolio investment, commodity prices and economic growth.
+Falling outside the immediate scope of the current study, 
+an autoregresive model could also be tried, including the past values of the 
+FTSE250 in the input space to provide a more accurate model. In addition, 
+a more granular data set could be utilised by applying a 
+Brownian bridge algorithm to generate daily macroeconomic data and 
+subsequently testing the RBF-NN against other neural networks 
+such as auto-regressive LSTMs and GRUs, as well as other machine learning methods suited for 
+forecasting such as Support Vector Machines (SVM).
 
 \bibliographystyle{plainnat}
-\bibliography{references}
+\bibliography{paper}
 

### 2024-08-13 tex/paper.tex
introduction + literature review

@@
-\documentclass[12pt,a4paper]{article}
+\documentclass[11pt,a4paper]{article}
 \usepackage{amsmath,amssymb}
@@
 \usepackage{setspace}
+\usepackage{tikz}
 \usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
 \usepackage{booktabs}
@@
 \usepackage{parskip}
-\usepackage{xcolor} \color{black}
+\usepackage{tablefootnote}
+\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
+\newcommand{\citeboth}[1]{\citeauthor{#1} \citep{#1}}
+
+\usepackage[
+    colorlinks=true,      % Enable colored links
+    linkcolor=green,       % Color for internal links (e.g., \ref, \pageref)
+    citecolor=blue,       % Color for citations
+    urlcolor=blue,        % Optional: Set PDF metadata
+]{hyperref}
 
@@
 of capturing complex relationships in noisy financial data, and thereby providing more
-accurate forecasts.
+accurate forecasts. Although econometric methods are still valuable in informing us
+of the relevance of various macroeconomic variables.
 
@@
 the FTSE 250 historically resembling the structure 
-of the UK economy more closely.
+of the UK economy more closely than the FTSE100.
 
@@
     \item To investigate which macroeconomic variables influence the stock market price significantly, and quantify the effects of the following selection of relevant macroeconomic indicators on the FTSE 250, chosen as a more representative measure of the UK stock market.
-An Auto-regressive Distributed Lag (ARDL) model is utilised, in conjunction with an Error Correction Model, over the January 1993 - March 2024 period: interest rate, inflation, USD exchange rate, and M3 money supply.
+An Auto-regressive Distributed Lag (ARDL) model is utilised, in conjunction with an Error Correction Model (ARDL-ECM) over the January 1993 - March 2024 period: interest rate, inflation, USD exchange rate, and M3 money supply.
     \item To investigate the efficacy of machine learning methods, in particular a radial basis function neural network, in predicting stock price movements based on the same macroeconomic indicators.
@@
 
-The Efficient Market Hypothesis in 1970 \cite{fama1970} argues that asset prices fully 
+\subsection{Stock Market-Macroeconomy Relationship}
+
+There are a number of works which investigate the impact of macroeconomic variables
+on the stock market.
+
+\subsubsection{Inflation}
+
+The Fisher hypothesis states that nominal interest rates adjust to expected 
+inflation, implying market efficiency as firms correctly anticipate inflation 
+and adjust prices. This was tested this in the US stock market by \citeboth{jaffe1976}, and
+\citeboth{bodie1976} who both found a negative relationship between inflation and stock returns.
+To account for this,
+\citeboth{mogdiliani1979} proposed the Inflation Illusion Hypothesis, 
+arguing that investors overestimate the discount rate in valuations, leading to mispricing. 
+\citeboth{gultekin1983} and \citeboth{firth1979}, however, found that the relationship between nominal stock returns and inflation in the UK is positive, and that real returns on UK stocks have remained stable even as inflation varied.
+Later, \citeboth{hasan2008} provided evidence against this, challenging the Fisher hypothesis.
+
+\subsubsection{Interest Rate}
+
+Interest rates, being the cost of borrowing, directly impact firms by increasing financing costs, reducing profitability, and lowering growth prospects.
+For consumers, higher rates raise borrowing costs, leading to reduced spending and demand, which further lowers corporate profitability.
+The empirical research supports this theory: \citeboth{alam2009} found a significant inverse relationship between interest rates and stock prices across multiple countries from 1988 to 2003; 
+\citeboth{talla2013} observed a negative relationship between interest rates and stock prices on the Stockholm Stock Exchange; 
+\citeboth{demir2019} reported similar findings for the Turkish Stock Exchange.
+In the case of the UK, \citeboth{neifar2023} used a Non-Linear ARDL model to find a significantly negative long-run relationship between interest rates and the UK stock market.
+Higher interest rates also encourage a shift from stocks to safer
+assets like bonds, this is known as the flight-to-quality phenomenon, where rising bond yields make bonds more attractive compared to stocks.
+\citeboth{asgharian2016} provided evidence for this, particularly in developed markets like the UK.
+This typically results in a negative correlation between stock prices and interest rates.
+
+\subsubsection{Exchange Rate}
+
+Classical economic theory explains the relationship between exchange rates and stock prices through flow and portfolio balance models.
+The flow model suggests that currency depreciations negatively impact the cash flow of international firms, reducing their stock prices.
+The portfolio balance model indicates that expected currency depreciations lead investors to shift from domestic to foreign assets, decreasing demand for domestic stocks and lowering their prices.
+Moreover, \citeboth{branson1977} established a bidirectional relation between the exchange rate and stock prices: the demand for money in each country depends on the performance of assets in that country’s currency, and conversely the demand for money is determined by real wealth, itself partly influenced by stock market returns.
+\citeboth{aydemir2009} confirmed this, in the case of the Turkish Stock Exchange. 
+\citeboth{khan2018} also found that the exchange rate showed a positive relationship with the Karachi Stock Exchange over the long run, and a negative relationship in the short-run. 
+\citeboth{wong2022} demonstrated cointegration between exchange rates and stock prices across multiple countries, including the UK.
+
+\subsubsection{Money Supply}
+
+One of the foundational studies on the relationship between money supply and 
+stock market prices was \citeboth{palmer1970}, who analysed the impact of M1 money supply on the S$\&$P 425 Industrial Stock Index. 
+He posited that trends in the nation's money stock could lead the 
+private sector to adjust wealth portfolios, resulting in 
+predictable movements in stock prices. \citeboth{homa1971} further argued that 
+stock prices are significantly influenced by the risk-free interest rate, 
+which is, according to the monetarist liquidity preference theory, a 
+function of money supply. They found that an increase in money supply 
+positively impacts stock prices. This was disputed by \citeboth{pesando1974}, 
+who critiqued the methodology used in their study. This was disputed by \citeboth{pícha2017},
+who investigated the effect of the post-2008 quantitative easing programs on the US and UK stock markets, finding 
+that the former was significantly affected by the M2 money supply. However, he found that in this period 
+the UK stock market was not significantly affected.
+However, later research utilising more advanced models and different data than \citeboth{palmer1970}, such as \citeboth{bahloul2017} and 
+\citeboth{synek2024} found that the money supply does significantly influences developed stock indices, including the UK.
+
+\subsection{Forecasting}
+
+\subsubsection{Efficient Market Hypothesis}
+
+The Efficient Market Hypothesis proposed in \citeboth{fama1970} argues that asset prices fully 
 reflect all available information; Fama outlined three forms of the EMH - 
@@
 \begin{enumerate}
-    \item Weak-form: Future stock prices cannot be predicted based on past prices, implying prices follow a random-walk - Louis Bachelier.
-Current stock prices fully reflect all information contained in past price movements.
-This rules out technical analysis as a viable tool in the modelling of stock price fluctuations.
-    \item Semi-strong form: Stock prices quickly adjust to reflect all publicly available information including not just price data (which is covered by the weak form EMH)  but also financial statements and reports, macroeconomic changes, political changes, and industry trends.
-This rules out fundamental analysis as a viable tool in the modelling of stock price fluctuations.
+    \item Weak-form: Future stock prices cannot be predicted based on past prices.
+    \item Semi-strong form: Stock prices quickly adjust to reflect all publicly available information including macroeconomic changes.  
     \item Strong-form: Stock markets are perfectly efficient, stock prices reflect all information, both public and private.
-This includes all publicly available information covered by the semi-strong form, insider information, and even future events that are not yet known to the public.
-This implies that it is impossible to enjoy a consistent advantage in predicting stock prices.
 \end{enumerate}
 
-The consistent success of investors such as Warren Buffett of 
-Berkshire Hathaway and Jim Simons of Renaissance Technologies 
-in the decades after the development of the EMH have challenged 
-its central argument, and have shown the possibility of achieving 
-consistent returns against the market. Even the creator of the 
-hypothesis conceded that many stock returns are in fact predictable from 
-previous values \cite{fama1991}. Furthermore, the empirical research is not in a
+The consistent success of various high-profile investors in the decades 
+after the development of the EMH have challenged 
+its central argument, something acknowledged in the revised formulation of the EMH 
+in \citeboth{fama1991}. Furthermore, the empirical research is not in a
 greement on whether the London Stock Exchange is efficient at any of the three levels. 
-Rounaghi and Zadeh (2016) gives evidence that the FTSE100 and FTSE250 are efficient over the period 2007 - 2013, and Libberton (2010) also gives evidence that the FTSE100 is weak-form efficient over 1995 - 2007.
-However, Borges (2010) gives evidence to reject the EMH for both the French and UK stock markets over the period 1993 - 2007; the findings of Ullah and Asghar (2023) support Borges (2010) and reject EMH for the FTSE 100 over the period 2012 - 2020.
-Additionally, Bhavsar (2015) rejects the EMH for both the FTSE 100 and FTSE 250 over the 13-year period January 2002 - December 2014.
-Some studies such as Rosini and Shenai (2020) found that the efficiency of the FTSE100 and FTSE250 varied over time, giving support to Andrew Lo’s Adaptive Market Hypothesis.
-The empirical findings then suggest that it is not impossible to determine, to some degree, movements in the price of stock indices on the LSE using macroeconomic variables.
+\citeboth{libberton2010} and \citeboth{rounaghi} give evidence of weak-form efficiency,
+however, \citeboth{borges2010}, \citeboth{asghar2023} and \citeboth{bhavsar2015} al provide evidence to reject the EMH. 
+Some studies such as \citeboth{rosini2020} even found that the 
+efficiency of the FTSE100 and FTSE250 varied over time. The empirical findings then suggest 
+that it is not impossible to determine movements in the 
+price of stock indices on the FTSE250 using macroeconomic variables.
 
-The Fisher hypothesis, developed by the neoclassicist economist Irving Fisher, states that the nominal interest rate will adjust to accommodate any changes in expected inflation (Frank, Robert; Bernanke, Ben; Antonovics, Kate; Heffetz, Ori.
-Principles of Macroeconomics.
-McGraw-Hill.
-pp.
-138–139).
-The implications of this in the stock market are that firms anticipate inflation correctly and adjust prices accordingly, and that the market is efficient, quickly incorporating inflation into prices.
-It was tested in the US stock market by Jaffe and Mandelker (1976) when they examined the relationship between monthly inflation and stock returns; they would find a negative relationship between inflation and stock returns over the period January 1951 - December 1971.
-In support of Jaffe and Mandelker is a large array of empirical research, including but not limited to, Bodie (1976) who found that stocks were not an effective hedge against inflation, as they were affected negatively by inflation; Fama and Schwert (1977) built on Bodie (1976), finding that over a roughly 20 year period stocks did not provide an effective hedge against inflation, and returns were negatively affected by inflation in this period; Mogdiliani and Cohn (1979) sought to root this phenomenon in economic theory, and came up with the Inflation Illusion Hypothesis, which posits that investors fail to correctly account for the impact of inflation on nominal earnings and interest rates, and consequently stock prices when they are valuing stocks, primarily through overestimating the correct discount rate in the Discounted Cash Flow valuation model (link in appendix?).
-Campbell and Vuolteenaho (2004) found empirical evidence to support this in the US, suggesting that nearly $80\%$ of stock-market time-series variation is caused by the level of inflation.
-Firth (1979) and Gultekin (1983) found that the relationship between nominal stock returns and inflation in the UK is positive, and that real returns on UK stocks have remained relatively stable even as inflation varied, in support of the generalised Fisher hypothesis.
-Hassan (2008) later found evidence against this when he investigated the relationship between stock returns and inflation in the UK, finding a significantly positive relationship between the two.
-Additionally, Cunado et.
-al.
-found no evidence of stocks acting as effective hedges over a roughly two-century period in the UK, providing further evidence against the Fisher hypothesis in the UK.
+\subsubsection{Machine Learning Methods for Forecasting}
 
-Interest rates are by definition the cost of borrowing.
-Therefore, higher interest rates increase the cost of borrowing for firms, which means higher expenses for financing operations and capital investments, in turn reducing profitability and adversely affecting public perceptions of future growth prospects; they also present increased borrowing costs for consumers, in turn leading to reduced spending and lower overall demand for the goods and services of companies, and thus profitability.
-Furthermore, interest rate rises present a higher opportunity cost, this is captured by the flight-to-quality or flight-to-safety phenomenon: investors move their investments from relatively volatile assets such as stocks to ‘safe’ assets such as bonds, which results in a negative correlation between stock and bond returns; owing to the inverse relationship between the bank rate and bond prices and thus positive relation with bond yields because bond yields = annual coupon payments (fixed) / bond price, this suggests that a rise in the bank rate will make bonds a more attractive investment than stocks, which again posits a negative relationship between stock prices and interest rates.
-- Provide sources for both theoretic claims -.
-Thus, we would expect interest rates to negatively affect stock prices, and our expectations are satisfied by the empirical research.
-Alam and Uddin (2009) investigated the impact of the monthly interest rate on the stock market price in Australia, Bangladesh, Canada, Chile, Colombia, Germany, Italy, Jamaica, Japan, Malaysia, Mexico, Philippine, S.
-Africa, Spain, and Venezuela over the period January 1988 to March 2003 and found a significant inverse relationship between the two.
-Talla (2013) also found evidence of a negative relationship between the interest rate and Stockholm Stock Exchange price change.
-Pilinkus and Boguslauskas (2009) found increases in the interest rate led to a decline in the Lithuanian stock market index value.
-Bernanke and Kuttner (2004)’s findings also indicated the considerable impact of unanticipated interest rate fluctuations on the stock market price.
-Asgharian et.
-al (2019) found evidence to support the flight-to-quality phenomenon, particularly in highly-developed capital markets such as the UK’s stock market.
-Malika (2023) used a Non-Linear ARDL model to find that the interest rate shows a significantly negative long-run relationship with the UK stock market price.
-
-Classical economic theory outlines the relationship between exchange rates and stock prices using flow and portfolio balance models, the former outlining the negative effect of currency depreciations on the cash flow of firms operating internationally, and in turn their stock price, and the latter describing investor propensity in the face of expected currency depreciations to redistribute their investment from domestic to foreign assets, thus decreasing demand and lowering stock prices.
-In addition, Dornbusch and Fisher (1980) showed that variations in the exchange rate affect future cash flows, which in turn affect the Present Value of a stock and thus stock prices based on the Discounted Cash Flow model.
-Additionally, Branson et.
-al (1977) in turn established a bidirectional relation between the exchange rate and stock prices, essentially that the demand for money in each country depends on the performance of assets, including the stock market, in that country’s currency, and conversely the demand for money is determined by real wealth, which is itself partly influenced by stock market returns.
-Aggarwal (1981) also studied the relationship between exchange rates and stock prices, using the monthly prices of the U.S.
-capital market and the floating exchange rates of the dollar for the period 1974 to 1978.
-The results showed a significantly positive correlation between stock prices and the exchange rate.
-The effect on less developed capital markets was naturally different, with Vanita and Khushboo (2015) reporting a significantly negative relationship between the exchange rate and stock prices of Russia, India and South Africa; however, both supported Branson’s thesis.
-More recently, Aydemir and Demirhan (2009) also found evidence to support Branson’s hypothesis in the Turkish Stock Exchange, namely that a bidirectional causal relationship between exchange rate and Turkish stock indices does exist.
-Furthermore, HT Wong (2022) found evidence of cointegration between the exchange rate and UK, Japanese, Malaysian, Philippine, Singaporean, Korean, German, Hong Kong and Indonesian stock prices.
-Javangwe and Takawira (2022) further found, through investigation of the South African stock market, that the exchange rate exerts a negative effect on the stock market price in the short run, and a positive effect in the long run.
-
-One of the earliest papers investigating the effect of money supply on the stock market price was Palmer (1970), who examined the effect of M1 money supply on the S$\&$P 425 Industrial Stock Index and found that ‘primary changes (defined as a trend occurring over a period of
-months) in the nation's money stock may motivate the private sector to adjust its wealth portfolios in such a manner as to yield predictable movements in the prices of corporate securities,’ positing a definite relation between the money supply and stock market price.
-Further research by Homa and Jaffe (1971) indicated that the price of a common stock is determined significantly by the risk-free rate of interest, which itself, according to the liquidity preference theory of the monetarist school of economics, a function of money supply, as more money circulating in the economy means more liquidity available for banks, thus affecting the cost of borrowing (interest) - source -.
-They found that the money supply has a positive impact on the average level of stock prices.
-This was disputed by Pesando (1974) who claimed that the inability of the model used in this study to produce accurate forecasts of the stock price was sufficient evidence against a relationship between the money supply and stock market prices.
-Baks and Kramer (1999), supported by Conover et.
-al (1999) findings, initially found significant evidence of the impact of money supply on stock markets across G-7 countries:  Canada, France, Germany, Italy, Japan, the UK and the US.
-This empirically supports the economic theory that increases in money supply generally raise the value of personal asset portfolios, which leads individuals to reallocate excess cash to stocks, in turn affecting stock prices.
-In the face of the drastic shock to the global economy in 2008, many central banks including the USCB and the Bank of England utilised Quantitative Easing (QE) for the first time.
-As a macroeconomic tool, QE involves the creation of digital money by the central bank to engage in large-scale asset purchasing to stimulate the economy - stock and other asset markets would theoretically be affected immediately by such a policy.
-Picha (2017) utilised a portfolio balance model to investigate the effect of QE, measured by the M2 money supply, on the S$\&$P 500 using Johansen’s cointegration methodology and found that the money supply indeed exerted a effect on the latter, and the fact that the adjustment to the long-run equilibrium took 6-months suggested that it was quite moderate.
-Moreover, cointegration between the money supply and S$\&$P 500 over the long-run was established, adding to a large extant literature that supports long-run dependence of stock price on money supply.
-In the same study, it was found that there did not exist a significant relationship between the money supply and the UK stock market.
-Included are Mukherjee and Naka (1995), Wongbangpo and Sharma (2002) and Lee (2006) who posit the long-run dependence of Japanese, New Zealand and ASEAN-5 indices respectively on the M1 money supply; furthermore, Bahloul et.
-al.
-(2016) establish that M3 money supply significantly influences developed stock indices such as the UK, USA, and Japan over the long-run.
-Synek (2024) found long term dependence of stock indices such as the FTSE100 and TSX on the M2 money supply.
-
-There have been many studies, excluding those already mentioned, that have investigated the impact of multiple macroeconomic variables on the stock market.
-Demir (2018), for example, used the ARDL model to investigate the impact of quarterly economic growth, interest rate, real effective exchange rate, crude oil prices, FDI, and FPI on the Borsa Istanbul 100 (BIST-100).
-The study found that economic growth, real effective exchange rate, portfolio investments, and foreign direct investments raised the stock market index value while the interest rate and crude oil prices negatively affected it.
-Khan and Khan (2018) dealt with the case of the Karachi Stock Exchange, investigating the effect of the interest rate, inflation rate, M2 money supply, USD exchange rate, economic activity, and exports on the KSE-100.
-Their findings indicate that the Karachi Stock Exchange, in the long run, was significantly positively affected by the money supply, and significantly negatively affected by the exchange rate, and interest rate; in the short run, all of the variables were found to be insignificant except for the exchange rate, which was again negatively cointegrated with the stock price.
-In the case of the London Stock Exchange, Asprem (1989) examined the relationship between various European stock indices, asset portfolios and macroeconomic variables in ten European countries, the author’s results showing that employment, inflation, imports and interest rate are inversely related to stock prices.
-The relationships between stock prices and macroeconomic variables were strongest in Germany, Switzerland and the UK, suggesting that these variables have a significant influence on stock indices in the UK.
-Additionally, Olomu (2015) examined monthly inflation, industrial production index, M1 money supply, effective exchange rate, and interest rate effect on the FTSE 100 and found that inflation and exchange rate showed a positive relationship with the FTSE100 over the long run, whereas the industrial production index, money supply and interest rate showed negative long-run relationships.
-Olawale et.
-al (2014) found that significant long-run relationships existed between FTSE100 returns and the industrial production index, interest rate, and inflation in the UK, and in the short-run, industrial production index, short-term interest rates, and unemployment rates have no significant causal link with FTSE100 returns; they also used M3 money supply as a proxy for Quantitative Easing, finding a significant positive impact on stock returns in the US, but a negative one in the UK.
-Finally, Marshall and Vasilev (2022) investigated the London Stock Exchange, and found that the FTSE 100 is sensitive to interest rate and exchange rate fluctuations but not sensitive to the employment rate, real GDP, gold prices or public sector budget deficit.
-
-The accelerating development of Machine learning (ML) presents a challenge to the EMH and traditional economic theories through its ability to field advanced computational tools capable of uncovering patterns and inefficiencies in the market that were previously thought to be obscured by the ‘noise’ and randomness posited by the EMH, and beyond the capability of standard econometric models.
-ML models are able to continually improve over time as they learn from new data, something which regular econometric models are incapable of doing, in particular Neural Networks (NN) have been used extensively for forecasting time-series data, including asset volatility, volume, and prices and future macroeconomic variable movements.
-
-Recurrent Neural Networks (RNN) are a type of neural network especially suited for sequential data due to their feedback loops which enable them to use both current and past inputs, therefore allowing information to persist.
-This allows RNNs to learn and take into account recent trends when training and making predictions.
-However, there is an intricacy in how RNNs perform optimisation during training which causes RNNs to lose memory in the long run; of course in many cases we desire the model to retain long-run dependencies, so the Long Short-Term Memory (LSTM) network was developed by Hochreiter and Schmidhuber and introduced in 1997.
-LSTMs have been used to great effect in many-time series forecasting problems, including stock price prediction, with some studies such as insert and insert achieving insert metric of insert however they are not suitable for small, noisy datasets - such as those involving economic data - and tend to overfit the training data (Foster et.
-al.
-1991).
+The accelerating development of Machine learning presents a challenge to the 
+EMH and traditional economic theories through its ability to leverage advanced 
+computational tools capable of uncovering patterns and inefficiencies in the
+market that were previously thought to be obscured by the ‘noise’ and randomness 
+posited by the EMH, and beyond the capability of standard econometric models. 
+ML models are able to continually improve over time as they learn from new data, 
+something which regular econometric models are incapable of doing, in particular 
+neural networks, mainly Recurrent Neural Networks (RNN) and Long-Short Term 
+Memory (LSTM) Networks, have been used extensively for forecasting time-series data, 
+including asset volatility, volume, and prices. However they are not suitable for smaller datasets
+and tend to overfit the training data \citep{foster1992}.
 
@@
 neural network that use radial basis functions at each node in the 
-activation layer of the network. They were developed in 1988 by 
-Broomhead and Lowe at the Royal Signals and Radar Establishment. 
-They have since been employed in a variety of contexts such as 
-hydrological forecasting (Chang et. al 2001), biomedical engineering 
-(Acikkar et. al 2001) and time-series analysis (insert), with success. 
+activation layer of the network. They have since been employed in a variety of contexts such as 
+hydrological forecasting \citep{chang2001}, biomedical engineering 
+\citep{ibrikci2002} and time-series analysis \citep{xiong2015} with success. 
 Whilst neural networks trained on small datasets often display unstable 
-behaviour during performance (LeBaron and Weigend 1998), The advantage of 
-RBF-NNs lies in their strong tolerance for input noise, being able to 
+behaviour during performance \citep{lebaron1998}, The advantage of 
+RBF-NNs lies in their ability to 
 operate effectively in extremely volatile financial time-series 
-environments (Cafferate et. al 2019), and their strong ability to 
-generalise on such data (Sharkawy 2020) as well as their efficacy on 
-relatively small datasets (Kosarac et. al 2022). RBF-NNs have been used 
-considerably for stock price forecasting. Esfandyari et. al (2001) 
+environments \citep{cafferata2019}, their strong ability to 
+generalise on such data \citep{sharkawy2020} as well as their efficacy on small datasets \citep{kosarac2022}. 
+\citeboth{esfandyari2016}
 supports this, using a sample size of 70 and 29 to forecast daily NASDAQ 
-index values using an Artificial Neural Network; they achieved test $R^2$
-statistics as small as 83$\%$ and large as 98$\%$. Cao et. al (2005) 
+index values using an Artificial Neural Network; they achieved a minimum
+test $R^2$ of 83$\%$. RBF-NNs have been used 
+considerably for stock price forecasting. \citeboth{cao2004} 
 used an RBF-NN based on the Optimal Partition Algorithm (OPA) to 
-forecast S$\&$P 500 prices, using the index values from January 1993 to 
-December 1995. Kumar et al. (2019) use an RBFN, with the Back-propagation 
+forecast S$\&$P 500 prices. \citeboth{dass2019} use an RBF-NN, with the Back-propagation 
 Algorithm (BPA) used instead to tune parameters; their results demonstrate 
-that the RBF-NN outperforms a standard feed-forward neural network with a 
-single hidden layer. Che et. al (2014) also found that the RBF-NN, on a 
-small training dataset, was able to predict 45 days in advance the value 
-of the Shanghai Composite Index, with a mean prediction error of 
+that the RBF-NN outperformed a standard feed-forward neural network. 
+\citeboth{ji2014} also found that the RBF-NN, on a 
+small training dataset, was able to predict the value 
+of the Shanghai Composite Index on a 45 day time-horizon, with a mean prediction error of 
 1.4387$\%$, suggesting that the RBF-NN is competent in time-series 
-forecasting using small training datasets. Aunjum et. al (2021) were 
-able to predict crude oil prices using macroeconomic and political news 
-using an RBF-NN, with an $R^2$ of 86$\%$. Recently, Abotaleb et.al (2024)
+forecasting using small training datasets. Recently, \citeboth{abotaleb2024}
 used quarterly data from 2001 to 2023 to model the influence of economic 
 policies on the quarterly prices of various stock markets including the 
-FTSE100 and Dow Jones price using an RBF-NN, finding that the model 
+FTSE100 using an RBF-NN, finding that the model 
 achieved relative forecast error rates as low as 6.8$\%$ and an error of 
@@
 
-For the scope of this research, monthly data spanning from October 1992—marking the inception of the FTSE 250 index—through to March 2024 has been utilised, giving us 
-altogether a dataset of 379 observations.
+FTSE-250 monthly price data between Oct 1992 to March 2024 has been utilised for this study giving 379 observations.
 
-During the data collection process, it was necessary to deal with an issue raised by the fact that the Bank of England’s Monetary Policy Committee (MPC) does not convene on a monthly basis, and are not obligated to announce a rate on a regular basis.
-Thus, it was imperative to address the resulting irregularities in the interest rate data; to ensure a more accurate representation of interest rate fluctuations over time, the data was adjusted using the following algorithm:
-\begin{enumerate}
-    \item If the interest rate was unchanged through the month, the previous set rate was used.
-    \item If the rate changed from $i_1,\ldots,i_n$ on day $d_j$, for $1\leq j\leq n$ of a month with $x$ days, then a weighted average $W_{int}$ was calculated using (1) - insert hyperlink - to more accurately account for the effects of the interest rate over the month.
-    \item The value of INT would accordingly be $W_{int}$ for those months in which the rate changed. 
-\end{enumerate}
+During the data collection process, it was necessary to deal with an issue raised by the fact that the Bank of England’s Monetary Policy Committee (MPC) 
+does not convene on a monthly basis, and are not obligated to 
+announce a rate on a regular basis. To address the resulting irregularities the data was adjusted using the following algorithm:
 
-\begin{align}
-    W_{int} = i_1 \frac{d_{j_1}}{x} + \cdots +i_n\frac{d_{j_2}}{x}
-\end{align}
+\begin{algorithm}[H]
+    \caption{Calculate monthly interest rate}
+    \label{alg:interest_rate_adjustment}
+    \KwData{Interest rate data adjustments by the Bank of England’s MPC}
+    \KwResult{Monthly interest rate data}
+
+    \For{each month $m$}{
+        \If{$\forall d \in D_m \; \text{(interest rate unchanged)}$}{
+            \textbf{Set:} $\text{Rate}_m \leftarrow \text{Rate}_{m-1}$\;
+        }
+        \Else{
+            \For{each day $d_j$ with a rate change}{
+                \textbf{Compute:} 
+                $W_{int} = \sum_{i_k \in I_m} i_k \frac{d_{j_k}}{x}$\;
+                where $I_m = \{i_1, i_2, \ldots, i_n\}$ is the set of interest rates on days with changes, 
+                and $d_{j_k}$ is the number of days for each rate $i_k$.
+            }
+            \textbf{Set:} $\text{Rate}_m \leftarrow W_{int}$\;
+        }
+    }
+\end{algorithm}
 
@@
 
-Table 1 presents the sources of the data; the original datasets are available on request. 
-The precedent for the selection of these variables as some of the most important 
-factors in the fluctuation of the stock market has already been set in the literature, however, 
-a brief explanation of why the M3 measure of the money supply was chosen is 
-required. The M3 measure includes a broader range of liquid assets 
-compared to M2, encompassing all the components of M2 plus large time 
-deposits, institutional money market funds, and other larger liquid assets
-making it a more comprehensive indicator of the money supply's impact on economic activity and financial markets.
+Over the period covered by our dataset, the FTSE 250 index has 
+shown substantial growth, reflecting the structural transition of the UK economy.
+It has also experienced significant volatility, highlighting significant market shifts driven by 
+various crises such as the Dotcom bubble burst, and geopolitical shocks.
 
@@
 \end{table}
-
-Over the period covered by our dataset, the FTSE 250 index has demonstrated substantial volatility with an average value of 11,013.27, fluctuating between 2,519.9 and 24,102.2. 
-This volatility reflects significant market shifts driven by events such as the 
-Dotcom bubble burst in 2000, the global financial crisis of 2008, and more recent geopolitical uncertainties such as Brexit, the COVID-19 pandemic, and the Russia-Ukraine conflict. However, 
-despite these fluctuations, the index has shown considerable growth, 
-indicative of a generally upward trend in the UK stock market over the 
-long term.
 
@@
 ADF Unit Root test is that of non-stationarity, and the alternate hypothesis 
-is that the variables are stationary\footnote{All of the variables are tested in logarithmic form. The asterisks *, **, ***, and **** denote statistical significance
-at the 1$\%$, 2.5$\%$, 5$\%$ and 10$\%$ significance levels.}.
+is that the variables are stationary.
 
@@
     \centering
-    \caption{Augmented Dickey-Fuller (ADF) Unit Root Test}
+    \caption{Augmented Dickey-Fuller (ADF) Unit Root Test\tablefootnote{All of the variables are tested in logarithmic form. The asterisks *, **, ***, and **** denote statistical significance
+    at the 1$\%$, 2.5$\%$, 5$\%$ and 10$\%$ significance levels.}}
     \begin{tabular}{lccc}
@@
 spread of the basis function at each centre.
+
 \begin{figure}[h]
     \centering
-    \includegraphics[width=0.5\textwidth]{bishopnn.png}
-    \caption{Courtesy of C. Bishop (1995)}
-    \label{fig:bishopnn}
+    \hspace{-2.6cm} 
+    \begin{tikzpicture}[node distance=3cm,
+                        thick, 
+                        main node/.style={circle, draw, font=\sffamily\Large\bfseries}]
+
+      \node[main node] (1) {$y_{1}$};
+      \node[main node] (2) [right of=1] {$y_{c}$};
+      \node[main node] (3) [below left of=1] {$\phi_{1}$};
+      \node[main node] (4) [below right of=2] {$\phi_{M}$};
+      \node[main node] (5) [below right of=3] {$x_{1}$};
+      \node[main node] (6) [below left of=4] {$x_{d}$};
+      \node[main node] (7) [left of=3] {$\phi_{0}$};
+
+      \draw[thick, dashed] (1) -- (2);
+      \draw[thick, dashed] (3) -- (4);
+      \draw[thick, dashed] (5) -- (6);
+      \draw[thick] (1) -- (3);
+      \draw[thick] (1) -- (4);
+      \draw[thick] (2) -- (4);
+      \draw[thick] (7) -- (1);
+      \draw[thick] (7) -- (2);
+      \draw[thick] (3) -- (1);
+      \draw[thick] (3) -- (2);
+      \draw[thick] (3) -- (6);
+      \draw[thick] (3) -- (5);
+      \draw[thick] (4) -- (5);
+      \draw[thick] (4) -- (6);
+
+    \end{tikzpicture}
+    \caption{RBF-NN Architecture}
+    \label{fig:your_label}
 \end{figure}

### 2024-08-13 tex/paper.tex
intro, review, algorithm modified + references added

@@
 dangers that stock markets carry, where the French colonial administration 
-in America went bankrupt as a result of a speculative bubble. Wth the
+in America went bankrupt as a result of a speculative bubble. With the
 growing interconnectedness of markets, the stock market as an institution 
@@
 Depression and the Global Financial Crisis of 2008/9. Thus, economic 
-policymakers and administrators must take into careful consideration 
-both the impacts of their macroeconomic policies on the stock market.
+policymakers and administrators require forecasts for how policy changes or 
+unexpected economic trends may impact the economy through the stock market.
 
-Investors and traders are also affected by the 
-fluctuations of economic factors due to their effect on firms.
-The ability to forecast market trends allows 
-investors to optimise their trading strategies, allowing them to make
+Investors and traders are also intimately affected by the 
+fluctuations of economic factors due to their destabilising effect on firms.
+The ability to forecast stock index price changes allows 
+investors to optimise their trading strategies, enabling them to make
 informed decisions about when to buy or sell assets. 
@@
 making it even more critical for investors to stay ahead of potential market movements.
-The traditional methods that have been leveraged in this context have been econometric, 
-however machine learning methods have emerged as powerful alternatives capable
-of capturing complex relationships in noisy financial data, and thereby providing more
-accurate forecasts. Although econometric methods are still valuable in informing us
-of the relevance of various macroeconomic variables.
 
-We utilise the FTSE 250 as our measure of the UK stock market, ecause unlike the 
+The traditional methods that have been leveraged in both of these contexts have been econometric.
+Machine learning methods, particularly neural networks, have emerged as powerful alternatives capable
+of capturing complex relationships in financial data, and thereby providing more
+accurate forecasts than econometric models.
+
+In this study we first use an Autoregressive Distributed Lag (ARDL) model in conjunction with an Error Correction Model (ECM), to investigate 
+the different effects of selected macroeconomic variables on the FTSE250 Index over the January 1993 - March 2024 period.
+Second, we test the efficacy of a Radial Basis Function Neural Network (RBF-NN) in forecasting the price movements of the FTSE250 based on the selected macroeconomic indicators, over the short-term 
+(5 months) and long-term (57 months).
+
+We utilise the FTSE 250 as our measure of the UK stock market, because unlike the 
 FTSE100 its overseas share of earnings is 
@@
 
-Thus, the study has a twofold aim:
-\begin{enumerate}
-    \item To investigate which macroeconomic variables influence the stock market price significantly, and quantify the effects of the following selection of relevant macroeconomic indicators on the FTSE 250, chosen as a more representative measure of the UK stock market.
-An Auto-regressive Distributed Lag (ARDL) model is utilised, in conjunction with an Error Correction Model (ARDL-ECM) over the January 1993 - March 2024 period: interest rate, inflation, USD exchange rate, and M3 money supply.
-    \item To investigate the efficacy of machine learning methods, in particular a radial basis function neural network, in predicting stock price movements based on the same macroeconomic indicators.
-\end{enumerate}
+The macroeconomic variables we consider are inflation, indicated by the Consumer Price
+Index (CPI); interest rate, represented by the Bank of England policy rate; the exchange rate, 
+represented by the USD/GBP exchange rate; and the money supply, represented by the 
+M3 measure of money supply, preferred over other 
+measures of the money supply because it considers more assets than M2 money supply 
+but does not include extremely illiquid assets as the M4 measure does. The selection 
+of these variables was based on the literature and the structure of the UK economy since 
+the establishment of the FTSE250 Index.
+
+The remainder of the paper is organised as follows. 
+Section 2 presents a review of the literature in this problem space, and explains the rationale behind our selection of macroeconomic variables.
+Section 3 covers the data utilised in the study. Section 4 outlines the ARDL-ECM model, our methodology, and our results describing
+the effect of the selected macroeconomic variables on the FTSE250. Section 5 constructs the RBF-NN model, and presents the short-term and long-term
+forecasts for the FTSE250 Index price movements, based on our selected macroeconomic variables. Section 6 concludes the study.
 
@@
 
-The USD/GBP exchange rate has exhibited moderate variability, 
-ranging from 0.4831 to 0.8834, suggesting periods of both 
-appreciation and depreciation of the Sterling. The little variability 
-there is can be
-attributed to major events such as the 2008 financial crisis and Brexit, 
-which both greatly affected the value of the Pound.
+The USD/GBP exchange rate has exhibited moderate variability, exhibiting the stable relations between 
+the two countries. Nevertheless, geopolitical shocks such as the 2008 financial crisis and Brexit both greatly affected the value of the Pound, 
+which accounts for the large range.
 
-The Consumer Price Index (CPI), with an average of 88.36, has 
-experienced significant inflationary fluctuations, ranging from 
-62.8 to 133.5. These fluctuations highlight varying price levels and 
-reflect periods of both high inflation, particularly in the late 2000s 
-near the global financial crisis, and also during 2021-22 as the Russia-Ukraine
-conflict destabilised energy and food markets, immediately after the devastation 
-of supply chains by the COVID-19 pandemic; and also low inflation in the 
-years of austerity in the UK and the global recession that onset during 
-the COVID19 pandemic.
+The CPI, with an average of 88.36, has 
+experienced significant fluctuations, highlight periods of high inflationary pressure such as in the late 2000s 
+near the global financial crisis, during 2021-22 as the Russia-Ukraine, and periods of stable, low
+inflation in the 2010s.
 
-Interest rates, averaging $3.21\%$, have varied widely from $0.1\%$ to $8.4\%$, with a standard deviation of $2.52\%$, reflecting the 
+Interest rates, have varied widely from $0.1\%$ to $8.4\%$, with a standard deviation of $2.52\%$, reflecting the 
 increasing role of monetary policy in the UK. The Bank of England was forced to adjust interest rates 
-many times in response to worsening economic conditions, examples being the steep rate cuts during the 2008 crisis to stimulate the economy and the gradual rate hikes in response to subsequent inflationary pressures.
+several times in response to worsening economic conditions, examples being the steep rate cuts during the 2008 financial crisis to stimulate the economy and the gradual rate hikes in response to 
+subsequent post-COVID inflationary pressures.
 
-The money supply (M3), averaging 1,861,958.47, has shown substantial variability, ranging from 494,181 to 3,704,873, 
+The M3 money supply has shown substantial variability, ranging from 494,181 to 3,704,873, 
 in large part influenced by the various Quantitative Easing (QE) programs implemented by the Bank of England 
-in response to the aforementioned crises. These programs aimed to inject liquidity into the economy, thus inflating the overall money supply.
+in response to the aforementioned crises. These programs aimed to inject liquidity into the economy to stimulate economic activity, thus inflating the overall money supply.
 
@@
         CPI & $3.157179$ & CPI $(-4)$ & $-0.921235$ \\
-        INT & $-0.1341097^{**}$ & INT $(-7)$ & $-0.034572^{****}$\\
-        EXCHG &  $-0.890146^{***}$ & EXCHG $(-7)$ & $0.122318$ \\
+        INT & $-0.1341097^{***}$ & INT $(-7)$ & $-0.034572^{****}$\\
+        EXCHG &  $-0.890146^{****}$ & EXCHG $(-7)$ & $0.122318$ \\
         M3 & $-0.1141114$ & M3 $(-4)$ & $-0.527798^{***}$ \\
@@
 
-The ARDL Bounds Test confirmed that our variable selection was sensible by virtue
-of its establishment of cointegration. The RBF-NN, it is hoped, will be able
-to capture intricate non-linear patterns in the data that the econometric model
-was unable to, and accordingly be able to predict with accuracy the future price
-movements. The RBF-NN was utilised for both short-term and long-term 
+The RBF-NN was utilised for both short-term and long-term 
 predictions. For short-term forecasting, 
-the model was tested over a 5-month horizon, based on the recommendation of the
-ARDL model, while the long-term predictions spanned 57 months. 
+the model was tested over a 5-month horizon, based on the optimum lag calculated earlier, while the long-term predictions spanned 57 months. 
 We differenced the data based on the recommendations of the 

### 2024-08-13 tex/paper.tex
updated references

@@
     on the FTSE250. The efficacy of Radial Basis Function 
-    Neural Networks (RBF-NN) in this problem domain is then investigated. We find that 
+    Neural Networks (RBF-NN) in this problem domain is then investigated. It is established 
     the RBF-NN is a competent short-term forecasting tool, and is able to 
@@
 
-In this study we first use an Autoregressive Distributed Lag (ARDL) model in conjunction with an Error Correction Model (ECM), to investigate 
+In this study, first an Autoregressive Distributed Lag (ARDL) model in conjunction with an Error Correction Model (ECM), is employed to investigate 
 the different effects of selected macroeconomic variables on the FTSE250 Index over the January 1993 - March 2024 period.
-Second, we test the efficacy of a Radial Basis Function Neural Network (RBF-NN) in forecasting the price movements of the FTSE250 based on the selected macroeconomic indicators, over the short-term 
-(5 months) and long-term (57 months).
+Second, the efficacy of a Radial Basis Function Neural Network (RBF-NN) is tested in forecasting the price movements of the FTSE250 based on the selected macroeconomic indicators, over the short-term 
+(5 months) and long-term (57 months), the former intended to inform investors and the latter to provide
+economic policymakers with a tool to measure the influence of macroeconomic variables on the price of the stock market.
 
-We utilise the FTSE 250 as our measure of the UK stock market, because unlike the 
+The FTSE 250 is utilised as the measure of the UK stock market, because unlike the 
 FTSE100 its overseas share of earnings is 
@@
 
-The macroeconomic variables we consider are inflation, indicated by the Consumer Price
+The macroeconomic variables considered are inflation, indicated by the Consumer Price
 Index (CPI); interest rate, represented by the Bank of England policy rate; the exchange rate, 
@@
 
-There are a number of works which investigate the impact of macroeconomic variables
+There are a number of works which investigate the impact of 
+the selected macroeconomic variables
 on the stock market.
@@
 achieved relative forecast error rates as low as 6.8$\%$ and an error of 
-27.2$\%$ in the case of the FTSE100. Thus, we find that the RBF-NN is a 
+27.2$\%$ in the case of the FTSE100. Thus, it is concluded that the RBF-NN is a 
 competent tool in the sphere of financial time-series forecasting, and 
@@
 
-Based on the empirical research and characteristics of the UK economy, we 
-define the stock market index function as:
+Based on the empirical research and characteristics of the UK economy,
+the stock market function is defined as:
 
-\begin{equation}
-    FTSE_{250} = f\biggl(CPI, EXCHG, INT, M3\biggr)
-\end{equation}
+\begin{align}
+    FTSE_{250} = f\biggl(CPI, EXCHG, INT, M3\biggr) \label{eq:implicit}
+\end{align}
 
@@
 
-Before constructing the ARDL model we first apply a natural logarithmic transform
+Before constructing the ARDL model a natural logarithmic transform is applied
 on the variables, this is done because in a log-transformed model,
@@
 which furnishes us with easily interpretable results. Having done this,
-we must then check at which level the variables are stationary (sidenote) at.
-Stationarity refers to the statistical property of a time-series where 
-its mean, variance, and autocorrelation structure do not change over time; 
-either a variable is stationary at level ($I(0)$), requiring no alteration,
+the level at which the variables are stationary\footnote{Stationarity refers to the statistical property of a time-series where 
+its mean, variance, and autocorrelation structure do not change over time. Either a variable is stationary at level ($I(0)$), requiring no alteration,
 or it is stationary after taking the first difference ($I(1)$). 
-Non-stationarity leads to spurious regression, where relationships that 
-are not actually present between two time-series are produced. Various
-techniques exist for stationarity testing, however in this study the 
-Augmented Dickey-Fuller (ADF) unit root test, developed by (link), was used.
+Non-stationarity leads to spurious regression where relationships that 
+are not actually present between two time-series are produced.} is checked.
+Various techniques exist for stationarity testing, however in this study the 
+Augmented Dickey-Fuller (ADF) unit root test, was used.
 
 As the ARDL model involves lagged values of the dependent and independent 
-variables, we are required to determine the optimum lag for the variables 
+variables, it is necessary to determine the optimum lag for the variables 
 that minimises error; this was done using the Schwarz
@@
 
-The implicit form (2) is represented in the ARDL model as:
+The implicit form \eqref{eq:implicit} is represented in the ARDL model as:
 
@@
 significance of the lagged variables and compared against the critical values
-given by Pesaran et. al. (2001) to determine whether cointegration exists, 
-and at what significance level. If cointegration exists, then we estimate 
-the long-run coefficients $\kappa_i$ by estimating the cointegration equation:
+given by \citeboth{pesaran2001} to determine whether cointegration exists, 
+and at what significance level. If cointegration exists, then 
+the long-run coefficients $\kappa_i$ are estimating using the cointegration equation:
 
 \begin{align}
-    FTSE_{250} = \kappa_0 + \kappa_1 CPI + \kappa_2 INT + \kappa_3 EXCHG + \kappa_4 M3 + \varepsilon 
+    FTSE_{250} = \kappa_0 + \kappa_1 CPI + \kappa_2 INT + \kappa_3 EXCHG + \kappa_4 M3 + \varepsilon \label{eq:two}
 \end{align}
@@
 where the error correction term ECT is found by calculating the residual 
-produced by (3):
+produced by \eqref{eq:two}:
 
@@
 to the long-run equilibrium. To estimate the short-run coefficients, $\alpha_i$ and $\beta_k$, as 
-well as the short-run correction speed $\gamma$ we use OLS regression.
+well as the short-run correction speed $\gamma$ OLS regression is used.
 
@@
 
-We began by applying a natural logarithmic transform on the variables, and 
+First, a natural logarithmic transform was applied on the variables, and 
 then conducting the ADF unit root test on the data. The null hypothesis of the 
@@
 of ARDL(5, 4, 7, 7, 4). The resulting F-statistic of $4.0798$ was compared against the critical values
-given by Pesaran et. al. (2001), providing evidence of cointegration 
+given by \citeboth{pesaran2001}, providing evidence of cointegration 
 at the $10\%$, $5\%$ and $2.5\%$ significance levels, but not at the $1\%$ 
@@
         M3 & $-0.1141114$ & M3 $(-4)$ & $-0.527798^{***}$ \\
-        Intercept & $-3.447767$ & ECT $(-1)$ & $-0.16^{**}$ \\
+        Intercept & $-3.447767$ & ECT $(-1)$ & $-0.1001^{**}$ \\
         \bottomrule
@@
 interest rate have significant long term effects on the stock market price
-while money supply and inflation are statistically insignificant. Neifar (2023) 
+while money supply and inflation are statistically insignificant. \citeboth{neifar2023}
 proposes that this is because deflation, that is the negative movements in the 
@@
 The insignificant long-term effect of the money supply 
-was noted by Picha (2017) and appears to be confirmed here. There are a few explanations
+was noted by \citeboth{pícha2017} and appears to be confirmed here. There are a few explanations
 for why this may be.
@@
 against it by $1\%$ has an impact of $-0.89\%$ on the FTSE 250. The findings 
-are in line with Neifar (2023), Liang and Willett (2015), and Khan and Khan (2018).
+are in line with \citeboth{neifar2023} and \citeboth{khan2018}.
 
@@
 
-Turning to the short-run dynamics, we find that the error correction term 
+Turning to the short-run dynamics, the error correction term 
 is negative and statistically significant ensuring that long-run equilibrium 
-can be achieved. The adjustment speed of $0.16$ indicates that 
-roughly $16\%$ of the fluctuation in the previous month's price, caused by the 
+can be achieved. The adjustment speed of $0.1001$ indicates that 
+roughly $10\%$ of the fluctuation in the previous month's price caused by the 
 macroeconomic variable movements, is adjusted back to the long-run equilibrium in the current month.
@@
 By introducing several modifications to the exact interpolation 
-procedure, we obtain the RBF-NN
-model first constructed in (Broomhead and Lowe, 1988). 
+procedure, the RBF-NN
+model is obtained, as outlined originally in \citeboth{broomhead1988}. 
 The RBF-NN mapping takes the following form:
@@
 the model was tested over a 5-month horizon, based on the optimum lag calculated earlier, while the long-term predictions spanned 57 months. 
-We differenced the data based on the recommendations of the 
+The data was differenced based on the recommendations of the 
 unit root tests conducted earlier to achieve stationarity.
 
-We began by normalising the data, using a z-score normalisation. This is because
+First, the data was normalised using a z-score normalisation. This is because
 neural networks work best with data
@@
 (Mendelsohn, 1993, Preprocessing Data for Neural Networks). Following this 
-we had to program the RBF-NN into Python. As no specific RBF package exists
-in Python, we used the kernel regression approach outlined by 
-Bishop, using the GaussianProcessRegressor (GPR) package in Python.
+it was necessary to program the RBF-NN into Python. As no specific RBF package exists
+in Python, the kernel regression approach outlined by 
+\citeboth{bishop1995}, using the GaussianProcessRegressor (GPR) package in Python.
 
@@
 
-As this is a Bayesian model we need to specify a \textit{prior distribution} over the parameters, 
+As this is a Bayesian model a \textit{prior distribution} over the parameters must be specified, 
 expressing our beliefs about the prior
-parameters before we examine the observations. A covariance function or kernel
+parameters before the observations are used. A covariance function or kernel
 is specified according to the characteristics of the data; in our model this 
@@
 \begin{align}
-    p(w|y, X) \sim \mathcal{N}\bigl(\frac{A^{-1}Xy}{\sigma_{n}^{2}}, A^{-1}\bigr)
+    p(w|y, X) \sim \mathcal{N}\bigl(\frac{A^{-1}Xy}{\sigma_{n}^{2}}, A^{-1}\bigr) \label{eq:nine}
 \end{align}
@@
 values, weighted by their posterior probability, are averaged to obtain the
-predictive distribution, upon which predicted outputs are computed.
+predictive distribution, upon which predicted outputs are computed. \citep{rasmussen2006}
 
@@
 the posterior predictive distribution, and with it the predicted output, after 
-averaging over all possible $w$, weighted by their likelihood according to (10).
+averaging over all possible $w$, weighted by their likelihood according to \eqref{eq:nine}.
 
@@
 
-The interpretation of this is that, historically, the variations in the 
+The interpretation of this is that, over the long-run (4-5 years), the variations in the 
 Consumer Price Index, interest rate, USD/GBP exchange rate, and M3 money 
-supply have explained or accounted for $9\%$ of the fluctuations of the
-FTSE250 value; in the context of a phenomenon as complex as the stock market,
-which is influenced by an inumerable quantity of factors such as investor sentiment
-- inherently irrational, 
-political instability, geopolitical shocks, corporate governance, and financial regulation, this is 
-unsurprising. Furthermore, the selection of economic variables in this study
-is by no means exhaustive; commodity prices, international trade dynamics,
-economic growth, and many other variables also exert a non-insignificant 
-effect on the stock market.
-
-The short-term prediction yielded MAE = $0.07$ and 
-$R^2 = 61\%$, for a 5-month time horizon.
-The significantly better performance of the model in the short-run
-indicates that the selected macroeconomic variables possesses
-a stronger predictive power over shorter periods. This was highlighted 
-earlier in the ARDL model, with the interest rate and money supply significantly
-impacting the FTSE250 price.
-
-In the short-term, which we define as 5 months, 
-as per the optimum lag for the FTSE250 in the ARDL model, the market movements
-are more closely tied to recent deviations in the economy, because theoretically
-the short-run is not exposed to the various unforeseen external shocks and structural changes
-that dilute the relationship between these variables and the stock market in the 
-long-run. Essentially, as we stretch our prediction window, we 
-introduce more uncertainty and unknowns, which translates into a 
-lower predictive quality.
+supply explain or account for $9\%$ of the fluctuations of the
+FTSE250 value. This relates closely to the findings of \citeboth{conover1999}, 
+who used an econometric model to establish that $4\%$ of the variation in monthly UK stock returns can be 
+explained by variations in the money supply.
 
@@
 
-Furthermore, as the time horizon extends into the long-run, fundamental factors 
-begin to matter a lot more. The compound effect of earnings, cash flows and
-managerial efficiencies begins to outweigh the 
-influence of fluctuating macroeconomic sentiment. Thus, in the long-term
-we see a different picture altogether.
+In the context of a phenomenon as complex as the stock market,
+which is influenced by an inumerable quantity of factors such as investor sentiment
+- inherently irrational, 
+political instability, geopolitical shocks, corporate governance, and financial regulation, this result 
+unsurprising. Furthermore, the selection of economic variables in this study
+is by no means exhaustive; commodity prices, international trade dynamics,
+economic growth, and many other variables also exert a non-insignificant 
+effect on the stock market.
 
@@
 
+The short-term prediction yielded MAE = $0.07$ and 
+$R^2 = 61\%$, for a 5-month time horizon.
+The significantly better performance of the model in the short-run
+indicates that the selected macroeconomic variables possesses
+a stronger predictive power over shorter periods. This was highlighted 
+earlier in the ARDL model, with the interest rate and money supply significantly
+impacting the FTSE250 price in the short-run.
+
 \begin{figure}[h]
@@
 \end{figure}
+
+In the short-term, which defined as 5 months, 
+as per the optimum lag for the FTSE250 in the ARDL model, the market movements
+are more closely tied to recent deviations in the economy, because theoretically
+the short-run is not exposed to the various unforeseen external shocks and structural changes
+that dilute the relationship between these variables and the stock market in the 
+long-run. Essentially, as the prediction window expands, 
+more uncertainty and unknowns are introduced, which translates into 
+lower predictive quality. Furthermore, as the time horizon extends into the long-run, fundamental factors 
+begin to matter a lot more. The compound effect of earnings, cash flows and
+managerial efficiencies begins to outweigh the 
+influence of fluctuating macroeconomic sentiment. Thus, in the short-run 
+a different phenomenon is observed.
 
@@
 The the study also investigated the efficacy of machine learning methods in 
-the field of stock-econometric forecasting, and returned good results, providing investors
-with an accurate short-term forecast tool, and also providing economists
-with a quantitative non-econometric measure of the 
-long-term influence of the selected macroeconomic 
-variables on the FTSE250 that outperforms a standard econometric model.
+the field of stock-econometric forecasting. The RBF-NN performed well in
+short-term forecasting of the FTSE250, and offers investors with an accurate forecasting tool, 
+to leverage in their trading strategies. The long-term forecast was able to improve on the ARDL-ECM model, and provides 
+economists with a tool to compute non-econometric, quantitative measures of the 
+long-term influence of macroeconomic variables on the stock market.
 
@@
 subsequently testing the RBF-NN against other neural networks 
-such as auto-regressive LSTMs and GRUs, as well as other machine learning methods suited for 
+such as auto-regressive LSTMs, as well as other machine learning methods suited for 
 forecasting such as Support Vector Machines (SVM).

### 2024-08-13 tex/paper.tex
update

@@
 
-The FTSE 250 is utilised as the measure of the UK stock market, because unlike the 
+The FTSE250 is utilised as the measure of the UK stock market, because unlike the 
 FTSE100 its overseas share of earnings is 
-roughly $50\%$, similar to the S\&P 500. In contrast, nearly $80\%$ of the FTSE $100$'s earnings 
+roughly $50\%$, similar to the S\&P 500. In contrast, nearly $80\%$ of the FTSE$100$'s earnings 
 are from overseas. This is reflected in the constitution of the respective indices, with 
-the FTSE 250 historically resembling the structure 
+the FTSE250 historically resembling the structure 
 of the UK economy more closely than the FTSE100.
@@
 \citeboth{aydemir2009} confirmed this, in the case of the Turkish Stock Exchange. 
-\citeboth{khan2018} also found that the exchange rate showed a positive relationship with the Karachi Stock Exchange over the long run, and a negative relationship in the short-run. 
+\citeboth{khan2018} also found that the exchange rate showed a positive relationship with the Karachi Stock Exchange over the long-run, and a negative relationship in the short-run. 
 \citeboth{wong2022} demonstrated cointegration between exchange rates and stock prices across multiple countries, including the UK.
@@
 \subsection{Forecasting}
+
+There is a large body of literature regarding time-series forecasting. First, 
+a discussion on the feasibility of such forecasting is presented: the Efficient 
+Market Hypothesis. Following this, machine learning methods for forecasting are 
+presented,
 
@@
 
-The accelerating development of Machine learning presents a challenge to the 
-EMH and traditional economic theories through its ability to leverage advanced 
-computational tools capable of uncovering patterns and inefficiencies in the
-market that were previously thought to be obscured by the ‘noise’ and randomness 
-posited by the EMH, and beyond the capability of standard econometric models. 
-ML models are able to continually improve over time as they learn from new data, 
-something which regular econometric models are incapable of doing, in particular 
-neural networks, mainly Recurrent Neural Networks (RNN) and Long-Short Term 
-Memory (LSTM) Networks, have been used extensively for forecasting time-series data, 
-including asset volatility, volume, and prices. However they are not suitable for smaller datasets
-and tend to overfit the training data \citep{foster1992}.
+The accelerating development of machine learning presents a challenge to the EMH
+and traditional economic theories. 
+Machine learning leverages advanced computational 
+tools capable of uncovering patterns and inefficiencies 
+in the market that were previously thought to be obscured by the 
+‘noise’ and randomness posited by the EMH and beyond the capability 
+of standard econometric models. Machine learning models continually 
+improve over time as they learn from new data—something regular 
+econometric models are incapable of doing. Notably, g
+radient boosting machines and neural networks, including 
+Recurrent Neural Networks (RNNs) and Long Short-Term Memory (LSTM) 
+Networks, have been extensively used for forecasting time series data, 
+such as asset volatility, volume, and prices. However they 
+are not suitable for smaller datasets and tend to overfit on the training data
+in such datasets \citep{foster1992}.
 
@@
 
-Over the period covered by our dataset, the FTSE 250 index has 
+Over the period covered by our dataset, the FTSE250 index has 
 shown substantial growth, reflecting the structural transition of the UK economy.
@@
         \midrule
-        FTSE$_{250}$ & N/A & FTSE$_{250}$ $(-5)$  & $-0.033089$ \\
+        FTSE$_{250}$ & $-$ & FTSE$_{250}$ $(-5)$  & $-0.033089$ \\
         CPI & $3.157179$ & CPI $(-4)$ & $-0.921235$ \\
@@
         Normality & Shapiro-Wilk Test & $3.889$e-$08$ & Normality \\
-        CUSUM & CUSUM Test & N/A & Stable \\
-        CUSUM$^2$ & CUSUM$^2$ Test & N/A & Stable \\
+        CUSUM & CUSUM Test & $-$ & Stable \\
+        CUSUM$^2$ & CUSUM$^2$ Test & $-$ & Stable \\
         \bottomrule
@@
 
-The long term coefficients show that the interest rate and exchange rate
-interest rate have significant long term effects on the stock market price
+The long-term coefficients show that the interest rate and exchange rate
+interest rate have significant long-term effects on the stock market price
 while money supply and inflation are statistically insignificant. \citeboth{neifar2023}
@@
 
-Unsurprisingly, the USD/GBP exchange rate is significantly related to the FTSE 250's movements. It 
+Unsurprisingly, the USD/GBP exchange rate is significantly related to the FTSE250's movements. It 
 should not be forgotten that, although it is considered to be a more domestic indicator of the LSE, 
-at times $50\%$ of the FTSE 250's earnings have come from abroad, indicating that 
+at times $50\%$ of the FTSE250's earnings have come from abroad, indicating that 
 the value of the Pound has an important impact on this stock market. As the USD has been 
 the global reserve currency for many decades, the depreciation of the Pound 
-against it by $1\%$ has an impact of $-0.89\%$ on the FTSE 250. The findings 
+against it by $1\%$ has an impact of $-0.89\%$ on the FTSE250. The findings 
 are in line with \citeboth{neifar2023} and \citeboth{khan2018}.
@@
 The model returned a mean absolute error of MAE $=0.16$ and 
-$R^2 = 9\%$ for the long term prediction, outperforming the ARDL-Error Correction model by 
+$R^2 = 9\%$ for the long-term prediction, outperforming the ARDL-ECM model by 
 $2.4\%$.
@@
 supply explain or account for $9\%$ of the fluctuations of the
-FTSE250 value. This relates closely to the findings of \citeboth{conover1999}, 
+monthly FTSE250 value. This relates closely to the findings of \citeboth{conover1999}, 
 who used an econometric model to establish that $4\%$ of the variation in monthly UK stock returns can be 
-explained by variations in the money supply.
+explained by variations in the money supply. It also suggests a higher level of influence 
+than the ARDL-ECM model provided, that only $R^2=3.4\%$ of the monthly FTSE250 variations 
+are affected by the selected macroeconomic variables, demonstrating the ability of machine learning
+methods to capture patterns in the data unavailable to econometric models.
 

### 2024-08-13 tex/paper.tex
update

@@
 \usepackage[
-    colorlinks=true,      % Enable colored links
-    linkcolor=green,       % Color for internal links (e.g., \ref, \pageref)
-    citecolor=blue,       % Color for citations
-    urlcolor=blue,        % Optional: Set PDF metadata
+    colorlinks=true,      
+    linkcolor=green,       
+    citecolor=blue,       
+    urlcolor=blue,        
 ]{hyperref}
@@
 \author{Mateen Khan}
-\date{}
+\date{\today}
 
@@
 are affected by the selected macroeconomic variables, demonstrating the ability of machine learning
-methods to capture patterns in the data unavailable to econometric models.
+methods to capture subtle patterns in the data unavailable to econometric models.
 
@@
 
-This study first tested the findings of previous econometric research on the 
-UK stock market, relating the impact of some especially
-relevant macroeconomic factors on the FTSE250.
-
 Stock markets provide important information on the economy, 
@@
 
-The the study also investigated the efficacy of machine learning methods in 
+The study also investigated the efficacy of machine learning methods in 
 the field of stock-econometric forecasting. The RBF-NN performed well in
@@
 foreign direct and portfolio investment, commodity prices and economic growth.
-Falling outside the immediate scope of the current study, 
-an autoregresive model could also be tried, including the past values of the 
+An autoregresive model, falling outside the immediate scope of the current study, could also be tried, including the past values of the 
 FTSE250 in the input space to provide a more accurate model. In addition, 
@@
 such as auto-regressive LSTMs, as well as other machine learning methods suited for 
-forecasting such as Support Vector Machines (SVM).
+forecasting such as Gradient Boosting Machines and Support Vector Machines.
 

### 2024-08-13 tex/paper.tex
update

@@
 positively impacts stock prices. This was disputed by \citeboth{pesando1974}, 
-who critiqued the methodology used in their study. This was disputed by \citeboth{pícha2017},
-who investigated the effect of the post-2008 quantitative easing programs on the US and UK stock markets, finding 
-that the former was significantly affected by the M2 money supply. However, he found that in this period 
-the UK stock market was not significantly affected.
+who critiqued the methodology used in their study. 
 However, later research utilising more advanced models and different data than \citeboth{palmer1970}, such as \citeboth{bahloul2017} and 
 \citeboth{synek2024} found that the money supply does significantly influences developed stock indices, including the UK. 
+\citeboth{pícha2017}
+investigated the effect of the post-2008 quantitative easing programs on the US stock markets, finding 
+that it was significantly affected by the M2 money supply. 
+\citeboth{olawale2014} also found a significant positive impact on stock returns in the US, but a negative one in the UK stock market.
 
@@
 and at what significance level. If cointegration exists, then 
-the long-run coefficients $\kappa_i$ are estimating using the cointegration equation:
+the long-run coefficients $\kappa_i$ are estimated to find the cointegration equation:
 
@@
 
-The insignificant long-term effect of the money supply 
-was noted by \citeboth{pícha2017} and appears to be confirmed here. There are a few explanations
-for why this may be.
+The insignificant long-term effect of the money supply can be explained by several reasons.
+Firstly, as discussed in \citeboth{sellin2001}, it is possible that the FTSE250 is efficient in processing changes in the 
+money supply, at least over certain intervals if not always. This would mean anticipated changes in the money supply would not affect the stock market price, only the
+unanticipated component of a change in money supply would, as established by \citeboth{sorensen1982} and \citeboth{bernanke2005}. 
+It is also possible that the M3 measure is too broad of a measure and includes overly illiquid components that are not as impactful 
+on the stock market dynamics. In the short-run we see that the money supply does have a 
+significantly negative effect on the stock market, as outlined by \citeboth{olawale2014}. According to \citeboth{sellin2001}, 
+this supports the Keynesian argument that a positive money supply shock will lead people to
+expect a contraction in future monetary policy. 
+This leads to an increase in the demand for funds, which drives up the current rate of interest, in turn
+reducing the present value of earnings in the Discounted Cash Flow (DCF) stock valuation model. The effect is a decline in the stock market prices.
 
@@
 at times $50\%$ of the FTSE250's earnings have come from abroad, indicating that 
-the value of the Pound has an important impact on this stock market. As the USD has been 
+the value of the Pound has a significant impact on this stock market. 
+In the long run, the appreciation of the Pound would increase the relative 
+cost of UK exports increases, thereby reducing the 
+competitiveness of FTSE250 companies in international markets. This in turn would lower their earnings
+and depress the FTSE250. Moreover, foreign investors may view the stronger Pound as a signal of reduced returns on their investments in the UK, which can lead to decreased investment in the FTSE250. 
+In the short-term, however, before structural shifts can materialise we see that the
+weakening of the Pound has a positive effect. As the USD has been 
 the global reserve currency for many decades, the depreciation of the Pound 
-against it by $1\%$ has an impact of $-0.89\%$ on the FTSE250. The findings 
+against it will lead to increased revenue in Pound terms for FTSE250 companies with substantial overseas operations.
+This is because their foreign earnings, when converted back into Pounds, will be worth more, potentially increasing profits.
+The findings
 are in line with \citeboth{neifar2023} and \citeboth{khan2018}.
 
-The interest rate was also found to have significant negative long-run effects on the 
-stock market, affirming the findings of the extant literature ().
-The mechanisms by which this occurs were discussed earlier in the literature review.
+The interest rate was also found to have significant negative long-term and short-term effects on the stock market, confirming the findings of the extant literature \citep{alam2009,demir2019,neifar2023, talla2013}.
+This can be explained by the fact that higher interest 
+rates increase the cost of borrowing for companies, leading to reduced investment 
+and lower future earnings potential, in turn depressing stock prices. 
+Furthermore, as per the flight-to-quality phenomenon, persistently higher interest rates make bonds and other fixed-income 
+investments more attractive relative to stocks, leading investors to shift 
+their portfolios away from equities and toward safer assets over time \citep{asgharian2016}. 
+In the short term, interest rate hikes immediately decrease the present value of earnings, as per the 
+DCF valuation model, and depress prices.
 
-Turning to the short-run dynamics, the error correction term 
+The error correction term 
 is negative and statistically significant ensuring that long-run equilibrium 
@@
 roughly $10\%$ of the fluctuation in the previous month's price caused by the 
-macroeconomic variable movements, is adjusted back to the long-run equilibrium in the current month.
+macroeconomic variable movements is adjusted back to the long-run equilibrium in the current month.
 However, only increases in the interest rate and money supply are able to 
 significantly contribute towards this correction, with $1\%$ increases in the former 
-leading to a $-0.03\%$ change in the FTSE250 price 7-months later, and in the case of the latter
+leading to a $-0.03\%$ change in the FTSE250 price 7 months later, and in the case of the latter
 a $-0.53\%$ change after a 4-month lag.
@@
 \subsection{Model Architecture}
-
-The exact interpolation problem, which forms the motivation for RBF-NNs, 
-involves mapping a $d$-dimensional input space $\boldsymbol{\mathrm{x}}$ t
-o a one-dimensional target space $t$. The dataset consists of $N$ input
- vectors $\boldsymbol{\mathrm{x_n}}$, along with corresponding target 
- values $t_n$. The objective is to find a function
-  $h(\boldsymbol{\mathrm{x}})$ such that:
-
-\begin{align}
-  h(\boldsymbol{\mathrm{x_n}}) = t_n, \quad \text{for } n = 1, \ldots, N
-\end{align}
-
-The radial basis function approach (Powell, 1987) 
-introduces a set of $N$ basis functions, one for each 
-data point, which are functions of the distance
-$||\boldsymbol{\mathrm{x}} - \boldsymbol{\mathrm{x_n}}||$, where
-the basis function $\phi(\cdot)$ is some non-linear function. The output of the mapping is then expressed as a linear combination of these basis functions:
-
-\begin{align}
-    h(\boldsymbol{\mathrm{x}}) = \sum_{n=1}^{N} w_n \phi(||\boldsymbol{\mathrm{x}} - \boldsymbol{\mathrm{x_n}}||)
-\end{align}
-
-By introducing several modifications to the exact interpolation 
-procedure, the RBF-NN
-model is obtained, as outlined originally in \citeboth{broomhead1988}. 
-The RBF-NN mapping takes the following form:
-
-\begin{align}
-    y(\boldsymbol{\mathrm{x}}) =  \sum_{j=1}^{M} w_{j} \phi_j(\boldsymbol{\mathrm{x}}) + w_{0}
-\end{align}
-
-where the radial basis function $\phi$ is the Inverse Multi-Quadric:
-
-\[
-\phi(r) = \left(r^2 + \sigma^2 \right)^{-\beta}
-\]
-
-where $r = \|\boldsymbol{\mathrm{x}} - \boldsymbol{\mathrm{c}}\|$ is the Euclidean distance between the input vector 
-\(\boldsymbol{\mathrm{x}}\) and the center of the basis function 
-\(\boldsymbol{\mathrm{c}}\). The constant $\sigma$ represents the
-spread of the basis function at each centre.
 
@@
 
+The exact interpolation problem, which forms the motivation for RBF-NNs, 
+involves mapping a $d$-dimensional input space $\boldsymbol{\mathrm{X}}$ to
+a $c$-dimensional target space $\boldsymbol{\mathrm{y}}$. The dataset consists of $N$ input
+vectors $\boldsymbol{\mathrm{x_n}}$, along with corresponding target 
+values $y_n$. The objective is to find a function
+  $h(\boldsymbol{\mathrm{x}})$ such that:
+
+\begin{align}
+  h(\boldsymbol{\mathrm{x_n}}) = y_n, \quad \text{for } n = 1, \ldots, N
+\end{align}
+
+The radial basis function approach 
+introduces a set of $N$ basis functions, one for each 
+data point. The radial basis function, $\phi(\cdot)$ is some non-linear function of the Euclidean distance
+$||\boldsymbol{\mathrm{x}} - \boldsymbol{\mathrm{x_n}}||$. The output of the mapping is then expressed as a linear combination of these basis functions:
+
+\begin{align}
+    h(\boldsymbol{\mathrm{x}}) = \sum_{n=1}^{N} w_n \phi(||\boldsymbol{\mathrm{x}} - \boldsymbol{\mathrm{x_n}}||)
+\end{align}
+
+By introducing several modifications to the exact interpolation 
+procedure, the RBF-NN
+model is obtained, as outlined originally in \citeboth{broomhead1988}. 
+The RBF-NN mapping takes the following form:
+
+\begin{align}
+    y(\boldsymbol{\mathrm{x}}) =  \sum_{j=1}^{M} w_{j} \phi_j(\boldsymbol{\mathrm{x}}) + w_{0}
+\end{align}
+
+\begin{align}
+    \phi_j(\boldsymbol{\mathrm{x}})  &= f(||\boldsymbol{\mathrm{x}-c_j}||), \quad{j=1\ldots,M}
+\end{align}
+
+where $\boldsymbol{c_j}$ is a RBF centre and $f$ is the radial basis function.
+
 \subsection{Methodology}
@@
 
-First, the data was normalised using a z-score normalisation. This is because
+The data was normalised using a z-score normalisation. This is because
 neural networks work best with data
@@
 in Python, the kernel regression approach outlined by 
-\citeboth{bishop1995}, using the GaussianProcessRegressor (GPR) package in Python.
+\citeboth{bishop1995}, using the GaussianProcessRegressor package in Python.
 
@@
 values, weighted by their posterior probability, are averaged to obtain the
-predictive distribution, upon which predicted outputs are computed. \citep{rasmussen2006}
+predictive distribution, upon which predicted outputs are computed \citep{rasmussen2006}.
 

### 2024-08-13 ardl/ardl.R

@@ -116,10 +116,7 @@ print(shapiro_test_result)
 
 
 
-recursive_residuals <- rollapply(residuals, width = 1, by = 1, FUN = function(x) x, align = "right", fill = NA)
-cusum <- cumsum(residuals)
-plot(cusum, type = "l", col = "blue", main = "CUSUM of Recursive Residuals", xlab = "Time", ylab = "CUSUM")
-abline(h = 0, col = "red", lty = 2)
+library(memochange)
 
-efp_result <- efp(residuals ~ 1, type = "OLS-CUSUM")
-plot(efp_result)
+cusum_test(residuals_vector, trend = c("none", "linear"), tau = 0.2, type = c("LKT","SK"), m = 0, simu = 0, M = 10000)
+

### 2024-08-14 tex/paper.tex
abstract, intro, lit review

@@
 \usepackage[linesnumbered,ruled,vlined]{algorithm2e}
+\providecommand{\keywordname}{\textbf{Keywords:}} % Customize how "Keywords:" appears
+\newcommand{\keywords}[1]{%
+  \par\addvspace{\baselineskip}% Adds some space above the keywords
+  \noindent\keywordname\enspace\ignorespaces#1 % Displays "Keywords:" followed by the actual keywords
+}
 \newcommand{\citeboth}[1]{\citeauthor{#1} \citep{#1}}
@@
 \begin{abstract}
-    Investment decisions and economic policy are highly influenced by 
-    macroeconomic variables and the stock market. Econometric models
-    have been the predominant tools in this context. The aim of this study is to analyse the impacts of some especially important macroeconomic
-    indicators on the FTSE250 
-    (Financial Times Stock Exchange 250 Index), using monthly data 
-    from  October 1993 to April 2024 using an ARDL model, obtaining 
-    quantifiable estimates for the long-term and short-term impacts of the variables
-    on the FTSE250. The efficacy of Radial Basis Function 
-    Neural Networks (RBF-NN) in this problem domain is then investigated. It is established 
-    the RBF-NN is a competent short-term forecasting tool, and is able to 
-    provide an alternative approach to standard econometric models in measuring the long-term influence of
-    macroeconomic variables on the stock market.
+    This study investigates the impact of the Consumer Price Index, Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate and the M3 money supply measure on the FTSE 250 index, and assesses the effectiveness of radial basis function neural networks as a tool to forecast the price of the FTSE 250 index based on the same set of macroeconomic indicators.
+This study shows quantitative estimates for the long-term and short-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 index, and demonstrates the effectiveness of radial basis function neural networks as a short-term forecasting tool for the price of the FTSE 250 index, thus providing an alternative to econometric forecasting models.
+
+    \keywords{Macroeconomic indicators; Stock market; FTSE 250 Index; Autoregressive Distributed Lag
+    (ARDL) model, RBF Neural Network; Time-series forecasting}
 \end{abstract}
@@
 
-The stock market has played a central role in the economy and 
-political ascent of many states ever since its inception. 
-Accordingly, any instabilities that have arisen in the stock market 
-throughout history have had far-reaching destabilising effects on the 
-economy; the Mississippi Bubble (1717-1720) was an early example of the 
-dangers that stock markets carry, where the French colonial administration 
-in America went bankrupt as a result of a speculative bubble. With the
-growing interconnectedness of markets, the stock market as an institution 
-has also contributed significantly to global disasters such as the Great 
-Depression and the Global Financial Crisis of 2008/9. Thus, economic 
-policymakers and administrators require forecasts for how policy changes or 
-unexpected economic trends may impact the economy through the stock market.
+The stock market has played a central role in the economic and political ascent of many nations ever since its inception [1].
+However, whenever stock market instabilities have arisen in history they have had far-reaching destabilising effects.
+The Mississippi Bubble was an early example of the inherent dangers within stock markets in the form of speculative bubbles [2].
+In this particular case, the stock market crashed as result of an unsustainable rise in share prices fueled by speculation, leading to panic selling, a collapse in confidence, and a massive financial meltdown that rippled through the entire French economy..
+Another example is the bursting of the dot-com bubble which had a profound impact on the stock market, leading to a significant crash, particularly in technology stocks.
+It triggered a severe stock market crash by leading to a rapid sell-off of overvalued tech stocks, which then spread to the broader market.
+The collapse of key companies, along with forced selling due to margin calls, and the widespread loss of investor confidence created a downward spiral that devastated the stock market.
 
-Investors and traders are also intimately affected by the 
-fluctuations of economic factors due to their destabilising effect on firms.
-The ability to forecast stock index price changes allows 
-investors to optimise their trading strategies, enabling them to make
-informed decisions about when to buy or sell assets. 
-Additionally, in a globalised economy, the interconnectedness of 
-markets means that events in one region may have significant effects worldwide, 
-making it even more critical for investors to stay ahead of potential market movements.
+Nevertheless, stock markets provide essential benefits by allowing companies to raise capital, offering investment opportunities for individuals, and supporting economic growth through efficient capital allocation.
+They enhance corporate governance through transparency and accountability, facilitate price discovery and market liquidity, and contribute to government financing through taxes.(Shahbaz et al., 2016).
 
-The traditional methods that have been leveraged in both of these contexts have been econometric.
-Machine learning methods, particularly neural networks, have emerged as powerful alternatives capable
-of capturing complex relationships in financial data, and thereby providing more
-accurate forecasts than econometric models.
+The stock market is highly sensitive to macroeconomic indicators because these indicators provide insights into the overall health and direction of the economy.
+Stock markets react to macroeconomic indicators because these signals help investors anticipate future economic conditions, corporate profits, and investment returns.
+Positive indicators tend to boost stock prices, while negative indicators can lead to declines in stock prices.
 
-In this study, first an Autoregressive Distributed Lag (ARDL) model in conjunction with an Error Correction Model (ECM), is employed to investigate 
-the different effects of selected macroeconomic variables on the FTSE250 Index over the January 1993 - March 2024 period.
-Second, the efficacy of a Radial Basis Function Neural Network (RBF-NN) is tested in forecasting the price movements of the FTSE250 based on the selected macroeconomic indicators, over the short-term 
-(5 months) and long-term (57 months), the former intended to inform investors and the latter to provide
-economic policymakers with a tool to measure the influence of macroeconomic variables on the price of the stock market.
+The significance of this study is to quantify the effects of macroeconomic indicators on the stock market as well as assessing the use of these indicators to forecast stock prices over the long-term and short-term, both useful to economic policy makers and stock market investors respectively.
 
-The FTSE250 is utilised as the measure of the UK stock market, because unlike the 
-FTSE100 its overseas share of earnings is 
-roughly $50\%$, similar to the S\&P 500. In contrast, nearly $80\%$ of the FTSE$100$'s earnings 
-are from overseas. This is reflected in the constitution of the respective indices, with 
-the FTSE250 historically resembling the structure 
-of the UK economy more closely than the FTSE100.
+A key task which economic policy makers and investors undertake is to understand and predict how economic changes influence stock market behaviour, which is critical for making informed policy and investment decisions.
+This modelling of the impact of macroeconomic indicators on stock markets is typically achieved by a variety of methods: statistical analysis, econometric modelling, scenario planning, and advanced machine learning techniques.
+However, it is fraught with challenges that require careful consideration, advanced techniques, and a deep understanding of both economics and financial markets.
+In particular, there is a challenge to find a subset of macroeconomic indicators which can most accurately capture stock market dynamics.
 
-The macroeconomic variables considered are inflation, indicated by the Consumer Price
-Index (CPI); interest rate, represented by the Bank of England policy rate; the exchange rate, 
-represented by the USD/GBP exchange rate; and the money supply, represented by the 
-M3 measure of money supply, preferred over other 
-measures of the money supply because it considers more assets than M2 money supply 
-but does not include extremely illiquid assets as the M4 measure does. The selection 
-of these variables was based on the literature and the structure of the UK economy since 
-the establishment of the FTSE250 Index.
+This study aims to explore the relationship between macroeconomic indicators and stock markets 
+through modelling. Its primary objective is to address questions such as how specific 
+macroeconomic indicators influence stock market prices in major economies and whether machine 
+learning models can substitute traditional econometric models in predicting stock market prices. 
+Ultimately, it aims to investigate three aspects of modelling the impact of macroeconomic 
+indicators on stock markets: 
+\begin{enumerate}
+    \item To investigate how macroeconomic indicators influence the FTSE 250 index
+    \item To investigate the short-term and long-term effects of macroeconomic variables on the price of the FTSE 250 index
+    \item To investigate whether machine learning techniques can provide accurate short-term and long-term forecasts of the price of the FTSE 250 index when using a low-granularity dataset.
+
+\end{enumerate}
+
+In order to analyse the dynamic relationship between variables in time series data, this study will employ the Autoregressive Distributed Lag (ARDL) model in conjunction with an Error Correction Model (ECM).
+By integrating the ARDL model with the ECM, a comprehensive understanding of both the immediate and longer-term relationships between variables is obtained, which is valuable for robust economic analysis and policy-making.
+This research also employs radial basis function neural networks to predict the price of the FTSE 250 index.
+This artificial deep learning network is particularly useful for tasks like function approximation, pattern recognition, and regression when the dataset is of low-granularity.
 
 The remainder of the paper is organised as follows.
-Section 2 presents a review of the literature in this problem space, and explains the rationale behind our selection of macroeconomic variables.
-Section 3 covers the data utilised in the study. Section 4 outlines the ARDL-ECM model, our methodology, and our results describing
-the effect of the selected macroeconomic variables on the FTSE250. Section 5 constructs the RBF-NN model, and presents the short-term and long-term
-forecasts for the FTSE250 Index price movements, based on our selected macroeconomic variables. Section 6 concludes the study.
+Section 2 presents a brief review of the literature in this problem space, and explains the rationale behind our selection of macroeconomic variables.
+Section 3 covers the data utilised in this study.
+Section 4 outlines the ARDL-ECM model, the methodology, and the results achieved describing the effect of the selected macroeconomic variables on the FTSE 250 index.
+Section 5 constructs the radial basis function neural network (RBF-NN) model, and presents the results of forecasting the short-term and long-term forecasts for the price of the FTSE 250 Index price movements based on the selected macroeconomic variables.
+Section 6 concludes the study.
 
@@
 
-\subsection{Stock Market-Macroeconomy Relationship}
+The traditional methods that have been leveraged in stock market forecasting, 
+besides fundamental and technical analysis, have been econometric models such as 
+linear regression \citep{shakhla2018stock,roy2015stock}; ARMA and ARCH models, both extensively used in numerous stock market forecasting problems \citep{herwartz2017stock,hu2020stock,oberholzer2015garch,rounaghi}; and cointegration models such as the
+ARDL and ECM model have been used significantly in forecasting problems involving macroeconomic factors \citep{demir2019,khan2018,neifar2023}.
+
+In addition to econometric models, in recent years, machine learning methods, 
+particularly artificial and deep neural networks \citep{chen2015lstm,kara2011ann,long2019deep,nelson2017lstm}, have emerged as powerful alternatives capable of overcoming the limitations of econometric models to capture complex relationships in financial data \citep{rossi2018ml}, and thereby provide more accurate forecasts than ordinary statistical models \citep{lapitskaya2021armax}.
+
+\subsection{Stock Market-Macroeconomic Relationships}
 
@@
 the selected macroeconomic variables
-on the stock market.
+on the stock market. However, as of yet the effect of the entirety of the selected 
+macroeconomic variables has not been studied on the FTSE250.
+
+\subsubsection{FTSE 250 Index}
+
+The FTSE250 has historically been considered a more closer representation of the domestic economy than the FTSE100.
+Roughly $50\%$ of the earnings of the FTSE250 are from overseas, similar to the S\&P 500.
+In contrast, nearly $80\%$ of the FTSE$100$'s earnings are from overseas.
+This is reflected in the constitution of the respective indices, with the 
+FTSE250 historically resembling the structure of the UK economy more closely than the FTSE100 \citep{ftse250history}.
+In spite of this, the FTSE250 has seldom been utilised as the preferred indicator of the UK stock market in the literature.
 
@@
 improve over time as they learn from new data—something regular 
-econometric models are incapable of doing. Notably, g
-radient boosting machines and neural networks, including 
+econometric models are incapable of doing. Notably, gradient boosting machines and neural networks, including 
 Recurrent Neural Networks (RNNs) and Long Short-Term Memory (LSTM) 
 Networks, have been extensively used for forecasting time series data, 
-such as asset volatility, volume, and prices. However they 
+such as stock market volatility \citep{cho2022forecasting,praveenraj2023} and prices \citep{zhang2022lstm,song2023forecasting,dutta2024hybrid}. However they 
 are not suitable for smaller datasets and tend to overfit on the training data
@@
 index values using an Artificial Neural Network; they achieved a minimum
-test $R^2$ of 83$\%$. RBF-NNs have been used 
-considerably for stock price forecasting. \citeboth{cao2004} 
+test $R^2$ of 83$\%$.
+
+RBF-NNs have been used 
+considerably for stock market forecasting. \citeboth{cao2004} 
 used an RBF-NN based on the Optimal Partition Algorithm (OPA) to 
@@
 that the RBF-NN outperformed a standard feed-forward neural network. 
-\citeboth{ji2014} also found that the RBF-NN, on a 
-small training dataset, was able to predict the value 
+\citeboth{ji2014} also found that the RBF-NN was able to predict the value 
 of the Shanghai Composite Index on a 45 day time-horizon, with a mean prediction error of 
-1.4387$\%$, suggesting that the RBF-NN is competent in time-series 
-forecasting using small training datasets. Recently, \citeboth{abotaleb2024}
+1.4387$\%$. Recently, \citeboth{abotaleb2024}
 used quarterly data from 2001 to 2023 to model the influence of economic 
@@
 is able to operate effectively within the constraints placed by our 
-problem domain and data.
+problem domain and data. However, with the exception of \citeboth{abotaleb2024}, who used 
+quarterly data, the effectiveness of RBF-NNs in stock market forecasting using 
+low-granularity economic time-series has not been investigated and presents an area requiring further 
+research.
 
@@
 indicators, and if so to quantify the long-run and short-run relationships
-between the former and the latter (Panopoulou and Pittis 2004).
+between the former and the latter
 
@@
 
-$$ k(x,x') = \biggl( ||x-x'|| + \sigma^2\biggr)^{-\beta} $$
+\begin{align}
+    k(x,x') = \biggl( ||x-x'|| + \sigma^2\biggr)^{-\beta} \label{eq: gimq}
+\end{align}
 and used to construct the covariance matrix $K$, which along with 
 a mean of $0$ forms the distribution of the the weights 
-$$ w \sim \mathcal{N}\bigl(0, K\bigr)$$
-By Bayes' Theorem,
+$$ w \sim \mathcal{N}\bigl(m(x), K\bigr)$$
+
+The mean function $m(x)$ acts as a bias, providing a baseline expectation of the function values. Then, by Bayes' Theorem,
 
@@
 \end{align}
-where the covariance matrix $A^{-1} = K^{-1} + \frac{XX^T}{\sigma_{n}^2}$.
-To make predictions on a test case all possible $w$
-values, weighted by their posterior probability, are averaged to obtain the
-predictive distribution, upon which predicted outputs are computed \citep{rasmussen2006}.
+Tthe covariance matrix $A^{-1} = K^{-1} + \sigma_{n}^{-2}(XX^T$).
+To make predictions on a test case all possible $w$, weighted by their probability, are averaged to obtain the
+predicted output \citep{rasmussen2006}.
 
@@
 
-A $4$-dimensional input space $X$
-passes inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
-and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \), through the first
-layer, in which the covariance function $(x,x')$
-is applied, the goal being to determine the distribution of the weights. Prior to this,
-the optimal hyperparameters in the covariance function, which plays an analogous
-role to the RBF basis function, are calculated using GridSearch Cross Validation; these hyperparameters are:
+First, the optimal hyperparameters in the covariance function \eqref{eq: gimq} 
+are computed using GridSearch Cross Validation; these hyperparameters are:
 \begin{enumerate}
@@
 \end{enumerate}
-following this the model uses the training observation $(y,X)$ to find
-the posterior predictive distribution, and with it the predicted output, after 
-averaging over all possible $w$, weighted by their likelihood according to \eqref{eq:nine}.
+
+Then, training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
+and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) are successively passed through the 
+covariance function $(x,x')$, to obtain the matrix of the pairwise covariance of 
+all training observations, $K$.
+
+A $4$-dimensional input space $X$ takes in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
+calculates the posterior weights distribution using \eqref{eq:nine}. 
+Each possible weight vector $w$ represents a different function, explicitly 
+based on the covariance function, that could fit the data. The expectation is then taken, which
+effectively combines all weights and finds the mean, weighted by the 
+respective probabilities. This represents the best functional estimate of the price movement. Based on this estimate 
+the predictive output will be generated.
+
+The predictive distribution for the test observations are derived similarly. The 
+ultimate training process posterior weight distribution is used to derive the 
+first test weight disttribution, which then averages over all $w$ again to find the
+best functional estimate. It only requires past x and y.
+
+\begin{align}
+    p(y_{*}|x_{*}, X, y) \, dw = \mathcal{N}(\sigma_{n}^{2} x_{*}^{T}A^{-1}Xy, )
+\end{align}
 

### 2024-08-14 rbf-nn/results-plots.R

@@ -1,0 +1,37 @@
+library(ggplot2)
+updated_dataframe$daily.date <- as.Date(updated_dataframe$daily.date)
+
+ggplot(updated_dataframe, aes(x = daily.date)) +
+  geom_line(aes(y = DNFTSE, color = "Real"), linewidth = 1) +    # Plot DNFTSE first
+  geom_line(aes(y = daily.predicted.price, color = "Predicted"), linewidth = 0.5) +    # Plot Predicted second (on top)
+  scale_color_manual(values = c("Predicted" = "blue", "Real" = "green")) +
+  labs(title = "",
+       x = "",
+       y = "Price changes",
+       color = "") +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    legend.position = "bottom",
+    panel.grid.major = element_line(color = "grey80"),
+    panel.grid.minor = element_line(color = "grey90")
+  )
+
+long_data$Date <- as.Date(long_data$Date)
+
+# Create the plot
+ggplot(long_data, aes(x = Date)) +
+  geom_line(aes(y = Real, color = "Real"), linewidth = 1) +    # Plot Real first
+  geom_line(aes(y = Predicted, color = "Predicted"), linewidth = 0.5) +    # Plot Predicted second (on top)
+  scale_color_manual(values = c("Predicted" = "blue", "Real" = "green")) +
+  labs(title = "",
+       x = "Date",
+       y = "Price Changes",
+       color = "") +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    legend.position = "bottom",
+    panel.grid.major = element_line(color = "grey80"),
+    panel.grid.minor = element_line(color = "grey90")
+  )

### 2024-08-14 tex/paper.tex
update
lit review

@@
 In particular, there is a challenge to find a subset of macroeconomic indicators which can most accurately capture stock market dynamics.
+In addition, macroeconomic indicator data is of such low-granularity that it precludes the use of certain deep learning neural network methods for prediction purposes.
 
@@
 the selected macroeconomic variables
-on the stock market. However, as of yet the effect of the entirety of the selected 
-macroeconomic variables has not been studied on the FTSE250.
-
-\subsubsection{FTSE 250 Index}
-
-The FTSE250 has historically been considered a more closer representation of the domestic economy than the FTSE100.
-Roughly $50\%$ of the earnings of the FTSE250 are from overseas, similar to the S\&P 500.
-In contrast, nearly $80\%$ of the FTSE$100$'s earnings are from overseas.
-This is reflected in the constitution of the respective indices, with the 
-FTSE250 historically resembling the structure of the UK economy more closely than the FTSE100 \citep{ftse250history}.
-In spite of this, the FTSE250 has seldom been utilised as the preferred indicator of the UK stock market in the literature.
+on the stock market.
 
@@
 
-\subsection{Forecasting}
+\subsubsection{FTSE 250 Index}
+
+The FTSE250 has historically been considered a more closer representation of the domestic economy than the FTSE100.
+Roughly $50\%$ of the earnings of the FTSE250 are from overseas, similar to the S\&P 500.
+In contrast, nearly $80\%$ of the FTSE$100$'s earnings are from overseas.
+This is reflected in the constitution of the respective indices, with the 
+FTSE250 historically resembling the structure of the UK economy more closely than the FTSE100 \citep{ftse250history}.
+
+In spite of this, the FTSE250 has seldom been utilised as the preferred indicator of the UK stock market in the literature.
+With the exception of \citeboth{smit2023longrun}, \citeboth{khatri2024impact} and \citeboth{engstrom2006impact} who investigate the effect of variable changes on the FTSE All Share which aggregates the
+FTSE100, FTSE250 and FTSE SmallCap indices, as of yet, the impact of the four selected variables on the FTSE250 has not been empirically studied and 
+warrants further investigation.
+
+\subsection{Stock Market Forecasting}
 
@@
 is able to operate effectively within the constraints placed by our 
-problem domain and data. However, with the exception of \citeboth{abotaleb2024}, who used 
-quarterly data, the effectiveness of RBF-NNs in stock market forecasting using 
-low-granularity economic time-series has not been investigated and presents an area requiring further 
+problem domain and data.
+
+However, with the exception of \citeboth{abotaleb2024}, who used 
+quarterly data and the economic uncertainty index, the effectiveness of RBF-NNs in stock market forecasting using 
+major macroeconomic indicators and low-granularity time-series has not been investigated and presents an area for further 
 research.
@@
                 \textbf{Compute:} 
-                $W_{int} = \sum_{i_k \in I_m} i_k \frac{d_{j_k}}{x}$\;
+                $W_{int} = \sum_{i_k \in I_m} \frac{i_k  d_{j_k}}{x}$\;
                 where $I_m = \{i_1, i_2, \ldots, i_n\}$ is the set of interest rates on days with changes, 
-                and $d_{j_k}$ is the number of days for each rate $i_k$.
+                and $d_{j_k} \leq x$ is the number of days into the month each rate $i_k$ occurs on.
             }

### 2024-08-14 tex/paper.tex
lit review

@@
 The Fisher hypothesis states that nominal interest rates adjust to expected 
-inflation, implying market efficiency as firms correctly anticipate inflation 
+inflation, implying that markets are efficient as firms correctly anticipate inflation 
 and adjust prices. This was tested this in the US stock market by \citeboth{jaffe1976}, and
@@
 \citeboth{gultekin1983} and \citeboth{firth1979}, however, found that the relationship between nominal stock returns and inflation in the UK is positive, and that real returns on UK stocks have remained stable even as inflation varied.
-Later, \citeboth{hasan2008} provided evidence against this, challenging the Fisher hypothesis.
+Later, \citeboth{hasan2008} provided evidence against this, challenging the Fisher hypothesis and enabling 
+the use of inflation to
 
@@
 \citeboth{asgharian2016} provided evidence for this, particularly in developed markets like the UK.
-This typically results in a negative correlation between stock prices and interest rates.
+This suggests a typically negative correlation between stock prices and interest rates.
 
@@
 There is a large body of literature regarding time-series forecasting. First, 
-a discussion on the feasibility of such forecasting is presented: the Efficient 
-Market Hypothesis. Following this, machine learning methods for forecasting are 
-presented,
+a discussion on the feasibility of such forecasting is presented. Following this, machine learning methods for forecasting are 
+discussed.
 
@@
 its central argument, something acknowledged in the revised formulation of the EMH 
-in \citeboth{fama1991}. Furthermore, the empirical research is not in a
+by \citeboth{fama1991}. Furthermore, the empirical research is not in a
 greement on whether the London Stock Exchange is efficient at any of the three levels. 
@@
 efficiency of the FTSE100 and FTSE250 varied over time. The empirical findings then suggest 
-that it is not impossible to determine movements in the 
+that it is not impossible to forecast movements in the 
 price of stock indices on the FTSE250 using macroeconomic variables.
@@
 which furnishes us with easily interpretable results. Having done this,
-the level at which the variables are stationary\footnote{Stationarity refers to the statistical property of a time-series where 
-its mean, variance, and autocorrelation structure do not change over time. Either a variable is stationary at level ($I(0)$), requiring no alteration,
-or it is stationary after taking the first difference ($I(1)$). 
-Non-stationarity leads to spurious regression where relationships that 
-are not actually present between two time-series are produced.} is checked.
+the level at which the variables are stationary is checked.
 Various techniques exist for stationarity testing, however in this study the 

### 2024-08-14 tex/paper.tex
lit review intro update

@@
 The traditional methods that have been leveraged in stock market forecasting, 
-besides fundamental and technical analysis, have been econometric models such as 
-linear regression \citep{shakhla2018stock,roy2015stock}; ARMA and ARCH models, both extensively used in numerous stock market forecasting problems \citep{herwartz2017stock,hu2020stock,oberholzer2015garch,rounaghi}; and cointegration models such as the
-ARDL and ECM model have been used significantly in forecasting problems involving macroeconomic factors \citep{demir2019,khan2018,neifar2023}.
+besides fundamental and technical analysis, have been econometric models. Linear regression \citep{shakhla2018stock,roy2015stock} and ARMA and ARCH models have
+both extensively used in numerous stock market forecasting problems \citep{herwartz2017stock,hu2020stock,oberholzer2015garch,rounaghi}. Cointegration methods such as the 
+ARDL and ECM models have been used significantly in forecasting problems involving macroeconomic factors \citep{demir2019,khan2018,neifar2023}.
 
-In addition to econometric models, in recent years, machine learning methods, 
-particularly artificial and deep neural networks \citep{chen2015lstm,kara2011ann,long2019deep,nelson2017lstm}, have emerged as powerful alternatives capable of overcoming the limitations of econometric models to capture complex relationships in financial data \citep{rossi2018ml}, and thereby provide more accurate forecasts than ordinary statistical models \citep{lapitskaya2021armax}.
+In addition to econometric models, machine learning methods, 
+particularly artificial and deep neural networks \citep{chen2015lstm,kara2011ann,long2019deep,nelson2017lstm}, have recently emerged as powerful alternatives capable of overcoming the limitations of econometric models.
+They are able to capture complex relationships in financial data \citep{rossi2018ml}, and are thereby able to provide more accurate forecasts than ordinary statistical models \citep{lapitskaya2021armax}.
 
@@
 Interest rates, being the cost of borrowing, directly impact firms by increasing financing costs, reducing profitability, and lowering growth prospects.
-For consumers, higher rates raise borrowing costs, leading to reduced spending and demand, which further lowers corporate profitability.
+For consumers, higher rates raise borrowing costs, leading to reduced spending and demand, which further lowers corporate profitability \citep{boe199section}.
 The empirical research supports this theory: \citeboth{alam2009} found a significant inverse relationship between interest rates and stock prices across multiple countries from 1988 to 2003; 
@@
 The USD/GBP exchange rate has exhibited moderate variability, exhibiting the stable relations between 
-the two countries. Nevertheless, geopolitical shocks such as the 2008 financial crisis and Brexit both greatly affected the value of the Pound, 
+the two countries. Nonetheless, geopolitical shocks such as the 2008 financial crisis and Brexit both greatly affected the value of the Pound, 
 which accounts for the large range.
@@
 The CPI, with an average of 88.36, has 
-experienced significant fluctuations, highlight periods of high inflationary pressure such as in the late 2000s 
-near the global financial crisis, during 2021-22 as the Russia-Ukraine, and periods of stable, low
-inflation in the 2010s.
+experienced significant fluctuations, highlighting periods of high inflationary pressure such as in the late 2000s 
+near the global financial crisis, and during the onset of the Russia-Ukraine conflict in 2022, and periods of stable, low
+inflation such as the early to late 2010s.
 
 Interest rates, have varied widely from $0.1\%$ to $8.4\%$, with a standard deviation of $2.52\%$, reflecting the 
-increasing role of monetary policy in the UK. The Bank of England was forced to adjust interest rates 
+increasing role played by monetary policy in the UK since May 1997 when the Bank of England was given operational responsibility for setting interest rates \citep{king1997changes}.
+The Bank of England was forced to adjust interest rates
 several times in response to worsening economic conditions, examples being the steep rate cuts during the 2008 financial crisis to stimulate the economy and the gradual rate hikes in response to 
@@
 in large part influenced by the various Quantitative Easing (QE) programs implemented by the Bank of England 
-in response to the aforementioned crises. These programs aimed to inject liquidity into the economy to stimulate economic activity, thus inflating the overall money supply.
+in response to the recessions onset by the 2008 financial crisis and the COVID-19 pandemic.
+These programs aimed to inject liquidity into the economy to stimulate economic activity, thus inflating the overall money supply.
 
@@
 
-The study uses an autoregressive distributed lag (ARDL) model to determine 
+The study uses an autoregressive distributed lag (ARDL) model in conjunction with 
+an error correction model (ECM) to determine 
 whether a long-term dependence exists for the FTSE250 and our macroeconomic 
 indicators, and if so to quantify the long-run and short-run relationships
-between the former and the latter
+between the former and the latter.
 
@@
 on the variables, this is done because in a log-transformed model,
-the coefficients can be interpreted as elasticities. Put simply, 
-a coefficient in a model, where both the dependent and independent variables
-are in logarithmic form, represents the percentage change in the 
-dependent variable for a one percent change in an independent variable, 
-which furnishes us with easily interpretable results. Having done this,
+the coefficients can be interpreted as elasticities (percentage changes). Having done this,
 the level at which the variables are stationary is checked.
 Various techniques exist for stationarity testing, however in this study the 
-Augmented Dickey-Fuller (ADF) unit root test, was used.
-
-As the ARDL model involves lagged values of the dependent and independent 
-variables, it is necessary to determine the optimum lag for the variables 
-that minimises error; this was done using the Schwarz
+Augmented Dickey-Fuller (ADF) unit root test was used. The optimum lag for the variables in the ARDL model was then found using the Schwarz
 Information Criterion.
@@
 
-Here, the intercept term, denoting so and so, is given by $\alpha_0$; the 
+Here, the intercept term, represents the baseline level of \(\Delta \text{FTSE}_{250_t}\), is given by $\alpha_0$; the 
 remaining $\alpha_i$ terms represent the short-run relationships 

### 2024-08-14 tex/paper.tex
LR update

@@
 
-The traditional methods that have been leveraged in stock market forecasting, 
-besides fundamental and technical analysis, have been econometric models. Linear regression \citep{shakhla2018stock,roy2015stock} and ARMA and ARCH models have
-both extensively used in numerous stock market forecasting problems \citep{herwartz2017stock,hu2020stock,oberholzer2015garch,rounaghi}. Cointegration methods such as the 
-ARDL and ECM models have been used significantly in forecasting problems involving macroeconomic factors \citep{demir2019,khan2018,neifar2023}.
-
-In addition to econometric models, machine learning methods, 
-particularly artificial and deep neural networks \citep{chen2015lstm,kara2011ann,long2019deep,nelson2017lstm}, have recently emerged as powerful alternatives capable of overcoming the limitations of econometric models.
-They are able to capture complex relationships in financial data \citep{rossi2018ml}, and are thereby able to provide more accurate forecasts than ordinary statistical models \citep{lapitskaya2021armax}.
-
 \subsection{Stock Market-Macroeconomic Relationships}
 
-There are a number of works which investigate the impact of 
-the selected macroeconomic variables
-on the stock market.
+The findings of \citeboth{ChenRollRoss1986} state that a long-term equilibrium relationship 
+exists between stock market prices and relevant macroeconomic indicators. Based on this, 
+\citeboth{EngleGranger1987} proposed an econometric approach utilising a causality test and an 
+error correction model (ECM) to investigate this relationship.
+This approach has been used extensively \citep{QuadriMasih, Plíhal2016,olomu2015} to assess the impact of macroeconomic variables on the stock market.
+Other prominent techniques that have been developed include the Johansen cointegration model \citep{YadavKheraMishra2021,Ozcan2012,ChistiShakeelGanai2020} and the Autoregressive Distributed Lag (ARDL) model \citep{khan2018,demir2019,neifar2023}.
+
+The literature investigating the impacts of the selected variables on the stock market is now presented.
+A discussion on the FTSE250 as the most accurate representation of the UK stock market is also included.
 
@@
 The Fisher hypothesis states that nominal interest rates adjust to expected 
-inflation, implying that markets are efficient as firms correctly anticipate inflation 
-and adjust prices. This was tested this in the US stock market by \citeboth{jaffe1976}, and
-\citeboth{bodie1976} who both found a negative relationship between inflation and stock returns.
+inflation as firms correctly anticipate inflation 
+and adjust prices, implying that markets are efficient and that inflation has no impact on the stock market. This was tested this in the US stock market by \citeboth{jaffe1976}, and
+\citeboth{bodie1976} who both found a negative relationship between inflation and stock returns, providing evidence against the Fisher hypothesis.
 To account for this,
@@
 \citeboth{gultekin1983} and \citeboth{firth1979}, however, found that the relationship between nominal stock returns and inflation in the UK is positive, and that real returns on UK stocks have remained stable even as inflation varied.
-Later, \citeboth{hasan2008} provided evidence against this, challenging the Fisher hypothesis and enabling 
-the use of inflation to
+Later, \citeboth{hasan2008} provided evidence against this, challenging the Fisher hypothesis and establishing a 
+significant causal relationship between the stock market
 
@@
 For consumers, higher rates raise borrowing costs, leading to reduced spending and demand, which further lowers corporate profitability \citep{boe199section}.
-The empirical research supports this theory: \citeboth{alam2009} found a significant inverse relationship between interest rates and stock prices across multiple countries from 1988 to 2003; 
-\citeboth{talla2013} observed a negative relationship between interest rates and stock prices on the Stockholm Stock Exchange; 
+The empirical research supports this theory. \citeboth{alam2009} found a significant inverse relationship between interest rates and stock prices across multiple countries from 1988 to 2003.
+\citeboth{talla2013} observed a negative relationship between interest rates and stock prices on the Stockholm Stock Exchange.
 \citeboth{demir2019} reported similar findings for the Turkish Stock Exchange.
-In the case of the UK, \citeboth{neifar2023} used a Non-Linear ARDL model to find a significantly negative long-run relationship between interest rates and the UK stock market.
-Higher interest rates also encourage a shift from stocks to safer
-assets like bonds, this is known as the flight-to-quality phenomenon, where rising bond yields make bonds more attractive compared to stocks.
-\citeboth{asgharian2016} provided evidence for this, particularly in developed markets like the UK.
-This suggests a typically negative correlation between stock prices and interest rates.
+In the case of the UK, \citeboth{neifar2023} used a Non-Linear ARDL model to find a significantly negative long-run relationship between interest rates and the FTSE100.
+In addition, the flight-to-quality phenomenon that igher interest rates also encourage a shift from stocks to fixed-income 
+assets such as bonds. \citeboth{asgharian2016} provided evidence for this in developed stock markets, including the UK. This suggests interest rates 
+negative impact the stock market.
 
@@
 
-Classical economic theory explains the relationship between exchange rates and stock prices through flow and portfolio balance models.
-The flow model suggests that currency depreciations negatively impact the cash flow of international firms, reducing their stock prices.
+Classical economic theory explains the impact exchange rates have on stock prices through the portfolio balance model.
 The portfolio balance model indicates that expected currency depreciations lead investors to shift from domestic to foreign assets, decreasing demand for domestic stocks and lowering their prices.
-Moreover, \citeboth{branson1977} established a bidirectional relation between the exchange rate and stock prices: the demand for money in each country depends on the performance of assets in that country’s currency, and conversely the demand for money is determined by real wealth, itself partly influenced by stock market returns.
-\citeboth{aydemir2009} confirmed this, in the case of the Turkish Stock Exchange. 
-\citeboth{khan2018} also found that the exchange rate showed a positive relationship with the Karachi Stock Exchange over the long-run, and a negative relationship in the short-run. 
-\citeboth{wong2022} demonstrated cointegration between exchange rates and stock prices across multiple countries, including the UK.
+\citeboth{branson1977} established a bidirectional relation between the exchange rate and stock prices. This was confirmed
+in the case of the Turkish Stock Exchange by \citeboth{aydemir2009}. 
+Moreover, stock markets often contain multinational firms involved in international transactions.
+Their profits are strongly influenced by the exchange rate, and consequently so is the stock market price \citep{Wong2018}.
+In support of this, \citeboth{khan2018} found that the exchange rate showed a positive relationship with the Karachi Stock Exchange over the long-run, and a negative impact in the short-run. 
+In the case of the UK, \citeboth{wong2022} also demonstrated a positive long-run impact by the exchange rate on the stock market.
 
@@
 
-One of the foundational studies on the relationship between money supply and 
+One of the foundational studies on the relationship between the money supply and 
 stock market prices was \citeboth{palmer1970}, who analysed the impact of M1 money supply on the S$\&$P 425 Industrial Stock Index. 
@@
 predictable movements in stock prices. \citeboth{homa1971} further argued that 
-stock prices are significantly influenced by the risk-free interest rate, 
-which is, according to the monetarist liquidity preference theory, a 
-function of money supply. They found that an increase in money supply 
-positively impacts stock prices. This was disputed by \citeboth{pesando1974}, 
-who critiqued the methodology used in their study. 
-However, later research utilising more advanced models and different data than \citeboth{palmer1970}, such as \citeboth{bahloul2017} and 
-\citeboth{synek2024} found that the money supply does significantly influences developed stock indices, including the UK. 
-\citeboth{pícha2017}
-investigated the effect of the post-2008 quantitative easing programs on the US stock markets, finding 
-that it was significantly affected by the M2 money supply. 
-\citeboth{olawale2014} also found a significant positive impact on stock returns in the US, but a negative one in the UK stock market.
+stock prices are significantly impacted by the risk-free interest rate, 
+which is a function of money supply. This relationship was disputed by \citeboth{pesando1974}, 
+who critiqued the theoretical and empirical methodology used in their study. In addition, \citeboth{sorensen1982} and 
+\citeboth{bernanke2005} argue that only the unanticipated component of money supply changes 
+affects the stock market price. Later research, however, such as \citeboth{bahloul2017}, 
+\citeboth{synek2024}, and \citeboth{pícha2017} has found that the price of developed stock markets, such as the US and UK's, is in fact significantly influenced by the money supply. 
+The nature of this relationship is disputed and varies across countries. \citeboth{homa1971} established a positive relationship between the 
+stock market and the money supply.
+This was disputed by \citeboth{sellin2001}, who presents a negative theoretical relationship betwen the stock market and money supply.
+\citeboth{olawale2014} empirically found a significant positive impact by the money supply on stock returns in the US, but a negative one in the case of the UK.
 
@@
 
-In spite of this, the FTSE250 has seldom been utilised as the preferred indicator of the UK stock market in the literature.
-With the exception of \citeboth{smit2023longrun}, \citeboth{khatri2024impact} and \citeboth{engstrom2006impact} who investigate the effect of variable changes on the FTSE All Share which aggregates the
-FTSE100, FTSE250 and FTSE SmallCap indices, as of yet, the impact of the four selected variables on the FTSE250 has not been empirically studied and 
+In spite of this, the FTSE250 has seldom been utilised as the preferred indicator of the UK stock market in the literature,
+with the exception of \citeboth{smit2023longrun}, \citeboth{khatri2024impact} and \citeboth{engstrom2006impact} who investigate the effect of variable changes on the FTSE All Share, an index that aggregates the
+FTSE100, FTSE250 and FTSE SmallCap indices. The impact of the four selected variables on the FTSE250 has not yet been empirically studied and 
 warrants further investigation.
@@
 
-There is a large body of literature regarding time-series forecasting. First, 
+There is a large body of literature regarding stock market price forecasting. First, 
 a discussion on the feasibility of such forecasting is presented. Following this, machine learning methods for forecasting are 
@@
 
-The Efficient Market Hypothesis proposed in \citeboth{fama1970} argues that asset prices fully 
-reflect all available information; Fama outlined three forms of the EMH - 
-weak-form, semi-strong form and strong form EMH:
+The Efficient Market Hypothesis proposed in \citeboth{fama1970} argues that stock market prices fully 
+reflect all available information. Fama outlined three forms of the EMH:
+weak-form, semi-strong form and strong form EMH.
 \begin{enumerate}
@@
 The consistent success of various high-profile investors in the decades 
-after the development of the EMH have challenged 
+after the development of the EMH has challenged 
 its central argument, something acknowledged in the revised formulation of the EMH 
-by \citeboth{fama1991}. Furthermore, the empirical research is not in a
-greement on whether the London Stock Exchange is efficient at any of the three levels. 
+by \citeboth{fama1991}. Furthermore, the empirical research is not in 
+agreement on whether the UK stock market is efficient at any of the three levels. 
 \citeboth{libberton2010} and \citeboth{rounaghi} give evidence of weak-form efficiency,
@@
 Some studies such as \citeboth{rosini2020} even found that the 
-efficiency of the FTSE100 and FTSE250 varied over time. The empirical findings then suggest 
+efficiency of the FTSE100 and FTSE250 varied over time. Thus, the empirical findings suggest 
 that it is not impossible to forecast movements in the 
@@
 
-The accelerating development of machine learning presents a challenge to the EMH
-and traditional economic theories. 
-Machine learning leverages advanced computational 
-tools capable of uncovering patterns and inefficiencies 
-in the market that were previously thought to be obscured by the 
-‘noise’ and randomness posited by the EMH and beyond the capability 
-of standard econometric models. Machine learning models continually 
-improve over time as they learn from new data—something regular 
-econometric models are incapable of doing. Notably, gradient boosting machines and neural networks, including 
-Recurrent Neural Networks (RNNs) and Long Short-Term Memory (LSTM) 
-Networks, have been extensively used for forecasting time series data, 
-such as stock market volatility \citep{cho2022forecasting,praveenraj2023} and prices \citep{zhang2022lstm,song2023forecasting,dutta2024hybrid}. However they 
-are not suitable for smaller datasets and tend to overfit on the training data
-in such datasets \citep{foster1992}.
+In addition to econometric models, machine learning methods, 
+particularly gradient boosting machines \citep{gumelar,Liu2024,Chen2023} and artificial and deep neural networks \citep{chen2015lstm,kara2011ann,long2019deep,nelson2017lstm}, have recently emerged as powerful alternatives capable of overcoming the limitations of econometric models.
+They are able to capture complex relationships in financial data \citep{rossi2018ml}, and 
+are thereby able to provide more accurate forecasts than ordinary statistical models 
+\citep{lapitskaya2021armax}.
+
+Deep learning methods such as Recurrent Neural Networks (RNNs) and Long Short-Term Memory (LSTM) Neural Networks
+have been extensively used for forecasting time series data, 
+such as stock market volatility \citep{cho2022forecasting,praveenraj2023} and prices \citep{zhang2022lstm,song2023forecasting,dutta2024hybrid}. The drawback of such 
+methods is that they are not suitable for smaller datasets and tend to overfit on the training data
+in such datasets \citep{foster1992} or display unstable behaviour during performance \citeboth{lebaron1998}
 
@@
 neural network that use radial basis functions at each node in the 
-activation layer of the network. They have since been employed in a variety of contexts such as 
-hydrological forecasting \citep{chang2001}, biomedical engineering 
-\citep{ibrikci2002} and time-series analysis \citep{xiong2015} with success. 
-Whilst neural networks trained on small datasets often display unstable 
-behaviour during performance \citep{lebaron1998}, The advantage of 
-RBF-NNs lies in their ability to 
+activation layer of the network.
+The advantage of RBF-NNs lies in their ability to 
 operate effectively in extremely volatile financial time-series 
@@
 
-RBF-NNs have been used 
-considerably for stock market forecasting. \citeboth{cao2004} 
+RBF-NNs have been used considerably for stock market forecasting. \citeboth{cao2004} 
 used an RBF-NN based on the Optimal Partition Algorithm (OPA) to 
@@
 \citeboth{ji2014} also found that the RBF-NN was able to predict the value 
-of the Shanghai Composite Index on a 45 day time-horizon, with a mean prediction error of 
+of the Shanghai Composite Index on a 45 day time-horizon using the closing price, with a mean prediction error of 
 1.4387$\%$. Recently, \citeboth{abotaleb2024}
-used quarterly data from 2001 to 2023 to model the influence of economic 
-policies on the quarterly prices of various stock markets including the 
-FTSE100 using an RBF-NN, finding that the model 
-achieved relative forecast error rates as low as 6.8$\%$ and an error of 
-27.2$\%$ in the case of the FTSE100. Thus, it is concluded that the RBF-NN is a 
-competent tool in the sphere of financial time-series forecasting, and 
-is able to operate effectively within the constraints placed by our 
-problem domain and data.
+used quarterly data from 2001 to 2023 to model the influence of the economic 
+uncertainty index on the quarterly prices of various stock markets including the 
+FTSE100 using an RBF-NN. They found that the model 
+achieved a relative error of 27.2$\%$ in the case of the FTSE100.
 
-However, with the exception of \citeboth{abotaleb2024}, who used 
-quarterly data and the economic uncertainty index, the effectiveness of RBF-NNs in stock market forecasting using 
-major macroeconomic indicators and low-granularity time-series has not been investigated and presents an area for further 
+Whilst \citeboth{abotaleb2024}, used quarterly data and a relatively minor economic indicator, the effectiveness of RBF-NNs in stock market forecasting using 
+major macroeconomic indicators and coarse-grain time-series has not yet been investigated and presents an area for further 
 research.

### 2024-08-15 tex/long_term_monthly.tex
draft

% Created by tikzDevice version 0.12.6 on 2024-08-15 14:44:41
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,289.08);
\begin{scope}
\path[clip] ( 41.49, 77.31) rectangle (500.39,266.42);
\definecolor{drawColor}{gray}{0.90}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.49, 96.28) --
	(500.39, 96.28);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.49,137.92) --
	(500.39,137.92);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.49,179.57) --
	(500.39,179.57);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.49,221.21) --
	(500.39,221.21);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.49,262.86) --
	(500.39,262.86);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 55.49, 77.31) --
	( 55.49,266.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (143.38, 77.31) --
	(143.38,266.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (231.26, 77.31) --
	(231.26,266.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (319.03, 77.31) --
	(319.03,266.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (406.79, 77.31) --
	(406.79,266.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (494.68, 77.31) --
	(494.68,266.42);
\definecolor{drawColor}{gray}{0.80}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.49,117.10) --
	(500.39,117.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.49,158.75) --
	(500.39,158.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.49,200.39) --
	(500.39,200.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.49,242.04) --
	(500.39,242.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 99.38, 77.31) --
	( 99.38,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (187.38, 77.31) --
	(187.38,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.15, 77.31) --
	(275.15,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (362.91, 77.31) --
	(362.91,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (450.68, 77.31) --
	(450.68,266.42);
\definecolor{drawColor}{RGB}{0,255,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 62.35,205.92) --
	( 69.80,193.01) --
	( 77.01,215.09) --
	( 84.47,202.69) --
	( 91.68,221.81) --
	( 99.14,229.38) --
	(106.59,180.36) --
	(113.56,151.33) --
	(121.02, 85.90) --
	(128.23,237.03) --
	(135.68,216.32) --
	(142.90,202.45) --
	(150.35,195.34) --
	(157.81,223.55) --
	(165.02,187.59) --
	(172.47,197.66) --
	(179.69,257.83) --
	(187.14,231.57) --
	(194.60,193.36) --
	(201.33,218.85) --
	(208.78,216.86) --
	(216.00,226.88) --
	(223.45,205.44) --
	(230.66,192.06) --
	(238.12,215.90) --
	(245.57,231.61) --
	(252.78,171.41) --
	(260.24,202.43) --
	(267.45,184.51) --
	(274.91,226.41) --
	(282.36,158.32) --
	(289.09,177.51) --
	(296.55,202.53) --
	(303.76,188.17) --
	(311.21,192.52) --
	(318.43,152.99) --
	(325.88,240.78) --
	(333.34,170.75) --
	(340.55,149.09) --
	(348.00,219.92) --
	(355.22,234.86) --
	(362.67,191.99) --
	(370.13,227.47) --
	(376.86,201.74) --
	(384.31,174.00) --
	(391.53,213.84) --
	(398.98,181.39) --
	(406.19,192.11) --
	(413.65,220.07) --
	(421.10,185.83) --
	(428.31,191.56) --
	(435.77,168.01) --
	(442.98,231.53) --
	(450.44,239.81) --
	(457.89,191.42) --
	(464.86,192.19) --
	(472.32,222.85) --
	(479.53,202.58);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 62.35,207.04) --
	( 69.80,202.58) --
	( 77.01,193.41) --
	( 84.47,207.21) --
	( 91.68,200.32) --
	( 99.14,200.01) --
	(106.59,204.48) --
	(113.56,176.23) --
	(121.02,201.09) --
	(128.23,203.61) --
	(135.68,197.91) --
	(142.90,205.17) --
	(150.35,201.19) --
	(157.81,201.04) --
	(165.02,213.74) --
	(172.47,213.80) --
	(179.69,214.20) --
	(187.14,203.90) --
	(194.60,209.41) --
	(201.33,206.21) --
	(208.78,206.94) --
	(216.00,208.57) --
	(223.45,198.53) --
	(230.66,202.09) --
	(238.12,215.10) --
	(245.57,200.30) --
	(252.78,204.65) --
	(260.24,205.57) --
	(267.45,205.89) --
	(274.91,203.29) --
	(282.36,201.54) --
	(289.09,201.01) --
	(296.55,203.94) --
	(303.76,194.77) --
	(311.21,203.74) --
	(318.43,200.22) --
	(325.88,206.35) --
	(333.34,190.58) --
	(340.55,203.50) --
	(348.00,200.70) --
	(355.22,199.32) --
	(362.67,200.88) --
	(370.13,198.74) --
	(376.86,201.52) --
	(384.31,201.55) --
	(391.53,204.83) --
	(398.98,201.30) --
	(406.19,198.34) --
	(413.65,199.50) --
	(421.10,200.25) --
	(428.31,207.28) --
	(435.77,202.72) --
	(442.98,202.61) --
	(450.44,229.12) --
	(457.89,198.15) --
	(464.86,203.79) --
	(472.32,204.65) --
	(479.53,215.21);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round,line cap=round] ( 41.49, 77.31) rectangle (500.39,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.54,114.07) {-0.50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.54,155.72) {-0.25};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.54,197.36) {0.00};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.54,239.01) {0.25};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (103.66, 68.07) {2020};

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (191.67, 68.07) {2021};

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (279.43, 68.07) {2022};

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (367.20, 68.07) {2023};

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (454.96, 68.07) {2024};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (270.94, 44.09) {$t$ = time};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,171.86) {$\delta$ = FTSE250 price change};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.6pt,line join=round] (225.41, 18.23) -- (236.98, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,255,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (287.67, 18.23) -- (299.24, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (243.92, 15.20) {Predicted};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (306.18, 15.20) {Real};
\end{scope}
\end{tikzpicture}

### 2024-08-15 tex/paper.tex
draft

@@
 
+% Define a command to create a hyperlink with only the figure number
+\newcommand{\figref}[1]{%
+    \hyperref[fig:#1]{Figure \ref{fig:#1}}%
+}
+
+% Define a command to create a hyperlink with only the algorithm number
+\newcommand{\algoref}[1]{%
+    \hyperref[alg:#1]{Algorithm \ref{alg:#1}}%
+}
+
 \usepackage[
@@
 \begin{abstract}
-    This study investigates the impact of the Consumer Price Index, Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate and the M3 money supply measure on the FTSE 250 index, and assesses the effectiveness of radial basis function neural networks as a tool to forecast the price of the FTSE 250 index based on the same set of macroeconomic indicators.
+    This study investigates the impact of the Consumer Price Index, Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate and the money supply (M3) measure on the FTSE 250 index, and assesses the effectiveness of radial basis function neural networks as a tool to forecast the price of the FTSE 250 index based on the same set of macroeconomic indicators.
 This study shows quantitative estimates for the long-term and short-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 index, and demonstrates the effectiveness of radial basis function neural networks as a short-term forecasting tool for the price of the FTSE 250 index, thus providing an alternative to econometric forecasting models.
@@
 
-The stock market has played a central role in the economic and political ascent of many nations ever since its inception [1].
-However, whenever stock market instabilities have arisen in history they have had far-reaching destabilising effects.
-The Mississippi Bubble was an early example of the inherent dangers within stock markets in the form of speculative bubbles [2].
-In this particular case, the stock market crashed as result of an unsustainable rise in share prices fueled by speculation, leading to panic selling, a collapse in confidence, and a massive financial meltdown that rippled through the entire French economy..
-Another example is the bursting of the dot-com bubble which had a profound impact on the stock market, leading to a significant crash, particularly in technology stocks.
-It triggered a severe stock market crash by leading to a rapid sell-off of overvalued tech stocks, which then spread to the broader market.
-The collapse of key companies, along with forced selling due to margin calls, and the widespread loss of investor confidence created a downward spiral that devastated the stock market.
+The stock market has played a central role in the economic and 
+political ascent of many nations ever since its inception \citep{levine2002}.  
+However, whenever stock market instabilities have arisen in history they 
+have had far-reaching destabilising effects. The Mississippi Bubble was 
+an early example of the inherent dangers within stock markets in the form of 
+speculative bubbles \citep{mississippibubble}. The role of
+the macroeconomy in this crash was central. The French central bank, the 
+\textit{Banque Royale}, aimed to convert its vast national debt into shares through the issuance of large 
+amounts of government bonds and paper notes. This considerably increased the money supply, in turn 
+greatly facilitating the ease of obtaining credit by individuals, many of whom invested in the Mississippi Company.
+A mass sell-off began after the \textit{Banque Royale} announced the number of paper notes exceeded the 
+value of metal coinage it possessed. This large-scale expansion of the money supply led to 
+inflation burgeoning, reaching $23\%$ at its peak, and devastating France \citep{moen2001john}.  Nonetheless, stock markets also provide essential benefits by allowing companies to 
+raise capital, offering investment opportunities for individuals, and 
+supporting economic growth through efficient capital allocation. 
+They enhance corporate governance through transparency and accountability, 
+facilitate price discovery and market liquidity, and contribute to government 
+financing through taxes \citep{shahbaz2016}. Effective and sustainable
+macroeconomic policy can greatly aid in this beneficial process. Therefore, it is essential 
+for economic policymakers to understand the impact of macroeconomic changes 
+on the stock market.
 
-Nevertheless, stock markets provide essential benefits by allowing companies to raise capital, offering investment opportunities for individuals, and supporting economic growth through efficient capital allocation.
-They enhance corporate governance through transparency and accountability, facilitate price discovery and market liquidity, and contribute to government financing through taxes.(Shahbaz et al., 2016).
+The sensitivity of the stock market to macroeconomic indicators is also of great 
+relevance to investors. Macroeconomic indicators provide 
+insights into the overall health and direction of the 
+economy, which affect investor sentiment and in turn influence market trends.
+Changes in the macroeconomy have an effect on both consumers and firms 
+which impacts their profitability, because these signals 
+help investors anticipate future economic conditions and corporate profits
+and consequently investment returns. Positive indicators tend to boost stock prices, while 
+negative indicators can lead to declines in stock prices.
 
-The stock market is highly sensitive to macroeconomic indicators because these indicators provide insights into the overall health and direction of the economy.
-Stock markets react to macroeconomic indicators because these signals help investors anticipate future economic conditions, corporate profits, and investment returns.
-Positive indicators tend to boost stock prices, while negative indicators can lead to declines in stock prices.
+The significance of this study is to quantify the effects of macroeconomic 
+indicators on the stock market as well as assessing the use of these indicators 
+to forecast stock prices over the long-term and short-term, both useful to 
+economic policy makers and stock market investors respectively.
 
-The significance of this study is to quantify the effects of macroeconomic indicators on the stock market as well as assessing the use of these indicators to forecast stock prices over the long-term and short-term, both useful to economic policy makers and stock market investors respectively.
-
-A key task which economic policy makers and investors undertake is to understand and predict how economic changes influence stock market behaviour, which is critical for making informed policy and investment decisions.
+A key task which economic policy makers and investors undertake is to 
+understand and predict how economic changes influence stock market behaviour, 
+which is critical for making informed policy and investment decisions. 
 This modelling of the impact of macroeconomic indicators on stock markets is typically achieved by a variety of methods: statistical analysis, econometric modelling, scenario planning, and advanced machine learning techniques.
@@
 In the case of the UK, \citeboth{neifar2023} used a Non-Linear ARDL model to find a significantly negative long-run relationship between interest rates and the FTSE100.
-In addition, the flight-to-quality phenomenon that igher interest rates also encourage a shift from stocks to fixed-income 
+In addition, the flight-to-quality phenomenon that higher interest rates also encourage a shift from stocks to fixed-income 
 assets such as bonds. \citeboth{asgharian2016} provided evidence for this in developed stock markets, including the UK. This suggests interest rates 
-negative impact the stock market.
+have a negative impact the stock market.
 
@@
 stock market prices was \citeboth{palmer1970}, who analysed the impact of M1 money supply on the S$\&$P 425 Industrial Stock Index. 
-He posited that trends in the nation's money stock could lead the 
+Palmer posited that trends in the nation's money stock could lead the 
 private sector to adjust wealth portfolios, resulting in 
@@
 
+\begin{table}[h!]
+    \centering
+    \caption{Data sources}
+    \begin{tabular}{lll}
+        \toprule
+        \textbf{Variable} & \textbf{Notation} & \textbf{Source} \\
+        \midrule
+        FTSE 250 Index & FTSE$_{250}$ & Yahoo Finance \\
+        Consumer Price Index & CPI & Office for National Statistics (ONS) \\
+        USD/GBP Exchange Rate (£) & EXCHG & Federal Reserve of St Louis \\
+        Interest Rate ($\%$) & INT & Bank of England \\
+        Money Supply (M3) (millions £) & M3 & Bank of England \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
 During the data collection process, it was necessary to deal with an issue raised by the fact that the Bank of England’s Monetary Policy Committee (MPC) 
 does not convene on a monthly basis, and are not obligated to 
-announce a rate on a regular basis. To address the resulting irregularities the data was adjusted using the following algorithm:
+announce a rate on a regular basis. To address the resulting irregularities the data was adjusted using \algoref{interest_rate_adjustment}.
 
@@
 \end{algorithm}
-
-\begin{table}[h!]
-    \centering
-    \caption{Data sources}
-    \begin{tabular}{lll}
-        \toprule
-        \textbf{Variable} & \textbf{Notation} & \textbf{Source} \\
-        \midrule
-        FTSE 250 Index & FTSE$_{250}$ & Yahoo Finance \\
-        Consumer Price Index & CPI & Office for National Statistics (ONS) \\
-        USD/GBP Exchange Rate (£) & EXCHG & Federal Reserve of St Louis \\
-        Interest Rate ($\%$) & INT & Bank of England \\
-        Money Supply (M3) (millions £) & M3 & Bank of England \\
-        \bottomrule
-    \end{tabular}
-\end{table}
 
@@
 The CPI, with an average of 88.36, has 
-experienced significant fluctuations, highlighting periods of high inflationary pressure such as in the late 2000s 
-near the global financial crisis, and during the onset of the Russia-Ukraine conflict in 2022, and periods of stable, low
+experienced significant fluctuations, highlighting periods of high inflationary pressure such as the onset of the Russia-Ukraine conflict in 2022, and periods of stable, low
 inflation such as the early to late 2010s.
@@
 
-Before constructing the ARDL model a natural logarithmic transform is applied
-on the variables, this is done because in a log-transformed model,
-the coefficients can be interpreted as elasticities (percentage changes). Having done this,
-the level at which the variables are stationary is checked.
-Various techniques exist for stationarity testing, however in this study the 
-Augmented Dickey-Fuller (ADF) unit root test was used. The optimum lag for the variables in the ARDL model was then found using the Schwarz
+Prior to constructing the ARDL model a natural logarithmic transform was applied
+on the variables. This was done so that the coefficients appear as elasticities, 
+which we can interpret easily as percentage changes. Subsequently,
+the level at which the variables are stationary was ascertained using the 
+Augmented Dickey-Fuller (ADF) unit root test. The optimum lag for the variables in the ARDL model was then determined using the Schwarz
 Information Criterion.
 
-The implicit form \eqref{eq:implicit} is represented in the ARDL model as:
+The implicit form \eqref{eq:implicit} was represented in the ARDL model as:
 
@@
 
-Here, the intercept term, represents the baseline level of \(\Delta \text{FTSE}_{250_t}\), is given by $\alpha_0$; the 
+Here, the intercept term was given by $\alpha_0$. The 
 remaining $\alpha_i$ terms represent the short-run relationships 
@@
 
-for, $i,j = 1,2,3,4,5$, with $i\neq j$. The F-statistic is computed for the joint
+for, $i,j = 1,2,3,4,5$, with $i\neq j$. The F-statistic was computed for the joint
 significance of the lagged variables and compared against the critical values
@@
 
-Then, an Error Correction Model (ECM) is constructed:
+Then, an Error Correction Model (ECM) was constructed:
 
@@
 
-where the error correction term ECT is found by calculating the residual 
+where the error correction term, ECT, was found by calculating the residual 
 produced by \eqref{eq:two}:
@@
 to the long-run equilibrium. To estimate the short-run coefficients, $\alpha_i$ and $\beta_k$, as 
-well as the short-run correction speed $\gamma$ OLS regression is used.
+well as the short-run correction speed $\gamma$ Ordinary Least Squares (OLS) regression was used.
+
+Finally, diagnostic tests are conducted to confirm that the model does not
+display autocorrelation, heteroskedascity, non-normality and that the cumulative 
+sum of residuals and squared residuals are both stable.
 
@@
 
-First, a natural logarithmic transform was applied on the variables, and 
-then conducting the ADF unit root test on the data. The null hypothesis of the 
-ADF Unit Root test is that of non-stationarity, and the alternate hypothesis 
+The null hypothesis of the ADF Unit Root test is that of non-stationarity, and the alternate hypothesis 
 is that the variables are stationary.
@@
 Schwarz Information Criterion, to obtain the optimum model 
-of ARDL(5, 4, 7, 7, 4). The resulting F-statistic of $4.0798$ was compared against the critical values
+of ARDL(5, 4, 7, 7, 4). The resulting F-statistic of $4.1$ was compared against the critical values
 given by \citeboth{pesaran2001}, providing evidence of cointegration 
@@
 display autocorrelation, heteroskedascity, non-normality and the cumulative 
-sum of residuals and squared residuals are both stable, indicating that the 
-relationship between the variables has not changed over time. The error
-correction term is also given, and it is negative, as expected.
+sum of residuals and squared residuals are both stable. The error
+correction term is also given, and it was negative, as expected.
 
@@
         M3 & $-0.1141114$ & M3 $(-4)$ & $-0.527798^{***}$ \\
-        Intercept & $-3.447767$ & ECT $(-1)$ & $-0.1001^{**}$ \\
+        Intercept & $-3.447767$ & ECT $(-1)$ & $-0.1794^{**}$ \\
         \bottomrule
@@
 
-Unsurprisingly, the USD/GBP exchange rate is significantly related to the FTSE250's movements. It 
+Unsurprisingly, the USD/GBP exchange rate was significantly related to the FTSE250's movements. It 
 should not be forgotten that, although it is considered to be a more domestic indicator of the LSE, 
@@
 The error correction term 
-is negative and statistically significant ensuring that long-run equilibrium 
-can be achieved. The adjustment speed of $0.1001$ indicates that 
-roughly $10\%$ of the fluctuation in the previous month's price caused by the 
-macroeconomic variable movements is adjusted back to the long-run equilibrium in the current month.
-However, only increases in the interest rate and money supply are able to 
-significantly contribute towards this correction, with $1\%$ increases in the former 
-leading to a $-0.03\%$ change in the FTSE250 price 7 months later, and in the case of the latter
-a $-0.53\%$ change after a 4-month lag.
+was negative and statistically significant ensuring that long-run equilibrium 
+can be achieved. The adjustment speed of $0.1794$ indicates that 
+roughly $18\%$ of the fluctuation in the previous month's price caused by the 
+macroeconomic variable movements is adjusted back to the long-run equilibrium in the subsequent month. This implies 
+that it takes approximately 6 months for the FTSE250 to return to its long-run equilibrium following 
+changes in the selected macroeconomic variables.
 
@@
     \centering
-    \hspace{-2.6cm} 
-    \begin{tikzpicture}[node distance=3cm,
-                        thick, 
-                        main node/.style={circle, draw, font=\sffamily\Large\bfseries}]
+    \begin{tikzpicture}[shorten >=1pt,->,draw=black!50, node distance=2cm']
 
-      \node[main node] (1) {$y_{1}$};
-      \node[main node] (2) [right of=1] {$y_{c}$};
-      \node[main node] (3) [below left of=1] {$\phi_{1}$};
-      \node[main node] (4) [below right of=2] {$\phi_{M}$};
-      \node[main node] (5) [below right of=3] {$x_{1}$};
-      \node[main node] (6) [below left of=4] {$x_{d}$};
-      \node[main node] (7) [left of=3] {$\phi_{0}$};
+        % Labels for the layers
+        \node[align=center] (inputlabel) at (-2,-2.8) {Input Layer};
+        \node[align=center] (hiddenlabel) at (3,-7) {Hidden Layer};
+        \node[align=center] (hiddenlabel) at (4,-0.3) {Bias};
+        \node[align=center] (outputlabel) at (8,-2.8) {Output Layer};
 
-      \draw[thick, dashed] (1) -- (2);
-      \draw[thick, dashed] (3) -- (4);
-      \draw[thick, dashed] (5) -- (6);
-      \draw[thick] (1) -- (3);
-      \draw[thick] (1) -- (4);
-      \draw[thick] (2) -- (4);
-      \draw[thick] (7) -- (1);
-      \draw[thick] (7) -- (2);
-      \draw[thick] (3) -- (1);
-      \draw[thick] (3) -- (2);
-      \draw[thick] (3) -- (6);
-      \draw[thick] (3) -- (5);
-      \draw[thick] (4) -- (5);
-      \draw[thick] (4) -- (6);
+        % Input Layer Nodes
+        \node[circle, draw, minimum size=1cm] (I-1) at (0,-1) {$x_1$};
+        \node[circle, draw, minimum size=1cm] (I-2) at (0,-2.5) {$x_2$};
+        \node[] (I-dots) at (0,-4) {$\vdots$};
+        \node[circle, draw, minimum size=1cm] (I-d) at (0,-5.5) {$x_d$};
+
+        % Hidden Layer Nodes
+        \node[circle, draw, minimum size=1cm] (H-0) at (3,-0.5) {$\phi_0$};
+        \node[circle, draw, minimum size=1cm] (H-1) at (3,-2) {$\phi_1$};
+        \node[] (H-dots) at (3,-3.5) {$\vdots$};
+        \node[circle, draw, minimum size=1cm] (H-M) at (3,-6) {$\phi_M$};
+
+        % Output Layer Node
+        \node[circle, draw, minimum size=1cm] (O) at (6,-2.75) {$y$};
+
+        % Connect input layer to hidden layer (excluding phi_0)
+        \foreach \i in {1,2,d}
+            \foreach \j in {1,M}
+                \draw[->] (I-\i) -- (H-\j);
+
+        % Connect phi_0 directly to output layer
+        \draw[->] (H-0) -- (O);
+
+        % Connect other hidden layer nodes to output layer
+        \foreach \j in {1,M}
+            \draw[->] (H-\j) -- (O);
 
@@
     \caption{RBF-NN Architecture}
-    \label{fig:your_label}
+    \label{fig:rbf_nn_architecture}
 \end{figure}
@@
 introduces a set of $N$ basis functions, one for each 
-data point. The radial basis function, $\phi(\cdot)$ is some non-linear function of the Euclidean distance
+data point. The radial basis function, $\phi(\cdot)$, is some non-linear function of the Euclidean distance
 $||\boldsymbol{\mathrm{x}} - \boldsymbol{\mathrm{x_n}}||$. The output of the mapping is then expressed as a linear combination of these basis functions:
@@
 procedure, the RBF-NN
-model is obtained, as outlined originally in \citeboth{broomhead1988}. 
+model is obtained, as presented in \figref{rbf_nn_architecture}. 
 The RBF-NN mapping takes the following form:
@@
 
-where $\boldsymbol{c_j}$ is a RBF centre and $f$ is the radial basis function.
+where $\boldsymbol{c_j}$ is a RBF centre and $f$ is the radial basis function \citep{bishop1995}.
 
@@
 The RBF-NN was utilised for both short-term and long-term 
-predictions. For short-term forecasting, 
-the model was tested over a 5-month horizon, based on the optimum lag calculated earlier, while the long-term predictions spanned 57 months. 
+forecasting. For the short-term, 
+the model was tested over a 5-month horizon, based on the optimum lag calculated earlier. 
+The long-term forecst spanned 57 months. 
 The data was differenced based on the recommendations of the 
@@
 
-The data was normalised using a z-score normalisation. This is because
-neural networks work best with data
-that is between a specified range; it ensures that the data is roughly uniformly
-distributed between the network inputs and outputs 
-(Mendelsohn, 1993, Preprocessing Data for Neural Networks). Following this 
-it was necessary to program the RBF-NN into Python. As no specific RBF package exists
-in Python, the kernel regression approach outlined by 
-\citeboth{bishop1995}, using the GaussianProcessRegressor package in Python.
+The data was normalised using a z-score normalisation, because neural networks work best with data
+that is between a specified range \citep{mendelsohn1993}. Following this 
+it was necessary to program the RBF-NN into Python. The kernel regression approach outlined by 
+\citeboth{bishop1995}, using the GaussianProcessRegressor package in Python was employed to model the RBF-NN.
 
@@
 
-The Scikit GaussianProcessRegressor is based on the 
-standard linear regression 
-model, equipped with Gaussian noise,
+The GaussianProcessRegressor function in Python's Scikit library is based on the 
+standard function
 
@@
 where $y$ differs from the function values $f(x)$ by noise which is assumed
-to follow an independent and identically distributed Gaussian distribution
+to follow an iid Gaussian distribution
 $\mathcal{N}_1$
@@
 probability density of the observations given the inputs, 
-which is factored over cases in the training set, 
-permissible because of the independence assumption,
+which is factored over cases in the training set. This is
+permissible because of the independence assumption \citep{rasmussen2006}.
 
@@
 
-As this is a Bayesian model a \textit{prior distribution} over the parameters must be specified, 
-expressing our beliefs about the prior
-parameters before the observations are used. A covariance function or kernel
-is specified according to the characteristics of the data; in our model this 
+The weights \( w \) represent the parameters that define the relationship between the inputs and outputs. 
+They are treated as random variables defined by the Gaussian distribution \eqref{eq: weight}.
+
+A covariance function or kernel
+is specified according to the characteristics of the data. In our model this 
 is the generalised inverse multiquadric
@@
 and used to construct the covariance matrix $K$, which along with 
-a mean of $0$ forms the distribution of the the weights 
-$$ w \sim \mathcal{N}\bigl(m(x), K\bigr)$$
-
-The mean function $m(x)$ acts as a bias, providing a baseline expectation of the function values. Then, by Bayes' Theorem,
+a mean function $m(x)$ forms the distribution of the the weights
 
 \begin{align}
-    p(w|y, X) = \frac{p(y|X,w) p(w)}{p(y|X)}
+    w \sim \mathcal{N}\bigl(m(x), K\bigr) \label{eq: weight}
 \end{align}
-Using this, the posterior distribution is then obtained
+
+The mean function $m(x)$ acts as a bias, 
+providing a baseline expectation of the function values. 
+This forms the prior distribution over the weights \eqref{eq: weight}.
+
+Then, by Bayes' Theorem,
+
+\begin{align}
+    p(w|y, X) = \frac{p(y|X,w) p(w)}{p(y|X)} \label{eq: post}
+\end{align}
+Using this, the posterior distribution \eqref{eq:nine} is obtained \citep{rasmussen2006}.
 
@@
 \end{align}
-Tthe covariance matrix $A^{-1} = K^{-1} + \sigma_{n}^{-2}(XX^T$).
-To make predictions on a test case all possible $w$, weighted by their probability, are averaged to obtain the
-predicted output \citep{rasmussen2006}.
+
+The covariance matrix $A^{-1} = K^{-1} + \sigma_{n}^{-2}(XX^T$).
+All possible $w$, weighted by their probability, are then averaged to obtain the
+most likely functional estimate of the output \citep{rasmussen2006}.
 
@@
 \begin{enumerate}
-    \item $\alpha$: a regularisation parameter which represents a small quantity
+    \item $\alpha$: a regularisation parameter representing a small quantity
     that is added to the diagonals of the covariance matrix to ensure it is 
@@
     \item $\beta$: the power in the covariance function, which controls the shape of the function.
-    \item $\sigma$: a parameter which controls how much of the noise in the data the kernel accounts for; naturally, 
-    it moderates the risk of overfitting and underfitting.
-    \item Length scale: a parameter that controls the influence - analogous to 
-    the width each RBF neuron possesses - that each kernel (analogous to node) is to have.
+    \item $\sigma$: a parameter which controls how much of the noise in the data the kernel accounts for. It moderates the risk of overfitting and underfitting.
+    \item Length scale: a parameter that controls the influence that each kernel is to have. Analogous to 
+    the width each RBF neuron possesses.
 \end{enumerate}
@@
 A $4$-dimensional input space $X$ takes in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
-calculates the posterior weights distribution using \eqref{eq:nine}. 
+calculates the posterior weights distribution. 
 Each possible weight vector $w$ represents a different function, explicitly 
-based on the covariance function, that could fit the data. The expectation is then taken, which
-effectively combines all weights and finds the mean, weighted by the 
+based on the RBF covariance function, that could fit the data. In the 
+hidden layer the expectation is then taken, which
+combines all weights and finds the mean, weighted by the 
 respective probabilities. This represents the best functional estimate of the price movement. Based on this estimate 
-the predictive output will be generated.
+the predicted output generated.
 
-The predictive distribution for the test observations are derived similarly. The 
-ultimate training process posterior weight distribution is used to derive the 
-first test weight disttribution, which then averages over all $w$ again to find the
-best functional estimate. It only requires past x and y.
+The testing process is conducted similarly. The 
+posterior weight distribution that contains the information of the training phase becomes the new 
+prior distribution. From this the posterior weight distributions for the test observations \eqref{eq: testw}
+are calculated, and then the predictive output is found \citep{rasmussen2006}.
 
 \begin{align}
-    p(y_{*}|x_{*}, X, y) \, dw = \mathcal{N}(\sigma_{n}^{2} x_{*}^{T}A^{-1}Xy, )
+    p(y_{*}|x_{*}, X, y) \, dw = \mathcal{N}(\sigma_{n}^{2} x_{*}^{T}A^{-1}Xy, x_{*}^{T}A^{-1}x_{*}) \label{eq: testw}
 \end{align}
@@
 The model returned a mean absolute error of MAE $=0.16$ and 
-$R^2 = 9\%$ for the long-term prediction, outperforming the ARDL-ECM model by 
-$2.4\%$.
+$R^2 = 9\%$ for the long-term prediction, providing a 
+$2.4\%$ increase in the $R^2$ obtained by the ARDL-ECM model.
 
-\begin{figure}[h]
+\begin{figure}[h!]
     \centering
-    \includegraphics[width=0.8\textwidth]{long-term-monthly.png}
-    \caption{Long-term predicted vs monthly real FTSE250 price changes}
-    \label{fig:lmonthly}
+    \input{long_term_monthly.tex}  
+    \caption{Long-term predicted vs real monthly FTSE250 changes}
+    \label{fig:longmonthly}
 \end{figure}
-
 The interpretation of this is that, over the long-run (4-5 years), the variations in the 
@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
     \centering
-    \includegraphics[width=0.8\textwidth]{long-term-daily.png}
-    \caption{Long-term predicted vs daily real FTSE250 price changes}
-    \label{fig:ldaily}
+    \input{long_term_daily.tex}  
+    \caption{Long-term predicted vs real daily FTSE250 changes}
+    \label{fig:longdaily}
 \end{figure}
@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
     \centering
-    \includegraphics[width=0.8\textwidth]{short-term-monthly.png}
-    \caption{Short-term predicted vs monthly real FTSE250 price changes}
-    \label{fig:smonthly}
+    \input{short_term_monthly.tex}  
+    \caption{Short-term predicted vs real monthly FTSE250 changes}
+    \label{fig:shortmonthly}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \input{short_term_daily.tex}  
+    \caption{Short-term predicted vs real daily FTSE250 changes}
+    \label{fig:shortdaily}
 \end{figure}
@@
 The short-term prediction yielded MAE = $0.07$ and 
-$R^2 = 61\%$, for a 5-month time horizon.
+$R^2 = 61\%$ for a 5-month time horizon.
 The significantly better performance of the model in the short-run
-indicates that the selected macroeconomic variables possesses
+indicates that the selected macroeconomic variables possess
 a stronger predictive power over shorter periods. This was highlighted 
@@
 impacting the FTSE250 price in the short-run.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{short-term-daily.png}
-    \caption{Short-term predicted vs daily real FTSE250 price changes}
-    \label{fig:sdaily}
-\end{figure}
 

### 2024-08-15 tex/short_term_daily.tex
draft

% Created by tikzDevice version 0.12.6 on 2024-08-15 14:32:08
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,289.08);
\begin{scope}
\path[clip] ( 37.09, 76.79) rectangle (500.39,266.42);
\definecolor{drawColor}{gray}{0.90}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.09, 85.39) --
	(500.39, 85.39);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.09,138.04) --
	(500.39,138.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.09,190.69) --
	(500.39,190.69);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 37.09,243.34) --
	(500.39,243.34);

\path[draw=drawColor,line width= 0.3pt,line join=round] (122.02, 76.79) --
	(122.02,266.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (225.59, 76.79) --
	(225.59,266.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (329.15, 76.79) --
	(329.15,266.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (434.45, 76.79) --
	(434.45,266.42);
\definecolor{drawColor}{gray}{0.80}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.09,111.72) --
	(500.39,111.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.09,164.37) --
	(500.39,164.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.09,217.02) --
	(500.39,217.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.50, 76.79) --
	( 68.50,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (175.53, 76.79) --
	(175.53,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.64, 76.79) --
	(275.64,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (382.67, 76.79) --
	(382.67,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (486.24, 76.79) --
	(486.24,266.42);
\definecolor{drawColor}{RGB}{0,255,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 58.15,172.17) --
	( 71.96,211.24) --
	( 75.41,213.25) --
	( 78.86,152.32) --
	( 82.31,207.00) --
	( 92.67,116.02) --
	( 96.12,190.68) --
	( 99.58,168.03) --
	(103.03,209.77) --
	(106.48,140.72) --
	(116.84,163.52) --
	(120.29,166.35) --
	(123.74,251.08) --
	(127.19,142.33) --
	(130.65,184.56) --
	(141.00,110.53) --
	(144.46,186.25) --
	(147.91,117.15) --
	(151.36,150.82) --
	(154.81,134.08) --
	(165.17,171.43) --
	(168.62,154.27) --
	(172.07,162.13) --
	(175.53,224.16) --
	(178.98,153.45) --
	(189.34,204.97) --
	(192.79,124.11) --
	(196.24,181.98) --
	(199.69,164.84) --
	(203.15,175.02) --
	(213.50,127.04) --
	(216.95,238.21) --
	(220.41,143.25) --
	(223.86,139.14) --
	(227.31,140.03) --
	(237.67,157.78) --
	(241.12,192.66) --
	(244.57,161.89) --
	(248.03,126.27) --
	(251.48,186.49) --
	(261.83,178.26) --
	(265.29,154.67) --
	(268.74,203.94) --
	(272.19,153.48) --
	(275.64, 85.41) --
	(286.00,192.13) --
	(289.45,158.59) --
	(292.91,111.06) --
	(296.36,135.16) --
	(299.81,159.67) --
	(310.17,183.27) --
	(313.62,155.11) --
	(317.07,164.71) --
	(320.52,184.91) --
	(323.98,157.28) --
	(334.33,171.33) --
	(337.79,178.53) --
	(341.24,150.76) --
	(344.69, 96.64) --
	(348.14,168.85) --
	(358.50,193.58) --
	(361.95,121.10) --
	(365.40,155.64) --
	(368.86,144.86) --
	(386.12,209.40) --
	(389.57,153.90) --
	(393.02,132.84) --
	(396.48,203.20) --
	(406.83,130.44) --
	(410.28,188.41) --
	(413.74,154.24) --
	(417.19,168.30) --
	(420.64,181.69) --
	(431.00,170.25) --
	(434.45,257.80) --
	(437.90,165.53) --
	(441.36,135.21) --
	(444.81,180.03) --
	(455.16,109.50) --
	(458.62,111.56) --
	(462.07,185.54) --
	(465.52,195.32) --
	(468.97,105.79) --
	(479.33, 95.66);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 58.15,255.18) --
	( 71.96,250.73) --
	( 75.41,246.28) --
	( 78.86,241.83) --
	( 82.31,237.38) --
	( 92.67,232.93) --
	( 96.12,228.48) --
	( 99.58,224.03) --
	(103.03,219.58) --
	(106.48,215.13) --
	(116.84,210.68) --
	(120.29,206.23) --
	(123.74,201.78) --
	(127.19,197.33) --
	(130.65,192.88) --
	(141.00,188.43) --
	(144.46,183.98) --
	(147.91,179.53) --
	(151.36,175.08) --
	(154.81,170.63) --
	(165.17,166.18) --
	(168.62,161.73) --
	(172.07,157.28) --
	(175.53,158.13) --
	(178.98,158.98) --
	(189.34,159.83) --
	(192.79,160.68) --
	(196.24,161.53) --
	(199.69,162.38) --
	(203.15,163.23) --
	(213.50,164.08) --
	(216.95,164.92) --
	(220.41,165.77) --
	(223.86,166.62) --
	(227.31,167.47) --
	(237.67,168.32) --
	(241.12,169.17) --
	(244.57,170.02) --
	(248.03,170.87) --
	(251.48,171.72) --
	(261.83,172.57) --
	(265.29,173.41) --
	(268.74,174.26) --
	(272.19,175.11) --
	(275.64,175.24) --
	(286.00,175.37) --
	(289.45,175.50) --
	(292.91,175.63) --
	(296.36,175.76) --
	(299.81,175.89) --
	(310.17,176.02) --
	(313.62,176.15) --
	(317.07,176.28) --
	(320.52,176.41) --
	(323.98,176.54) --
	(334.33,176.67) --
	(337.79,176.80) --
	(341.24,176.93) --
	(344.69,177.06) --
	(348.14,177.19) --
	(358.50,177.32) --
	(361.95,177.44) --
	(365.40,177.57) --
	(368.86,177.70) --
	(386.12,177.83) --
	(389.57,179.59) --
	(393.02,181.34) --
	(396.48,183.10) --
	(406.83,184.86) --
	(410.28,186.61) --
	(413.74,188.37) --
	(417.19,190.12) --
	(420.64,191.88) --
	(431.00,193.63) --
	(434.45,195.39) --
	(437.90,197.14) --
	(441.36,198.90) --
	(444.81,200.66) --
	(455.16,202.41) --
	(458.62,204.17) --
	(462.07,205.92) --
	(465.52,207.68) --
	(468.97,209.43) --
	(479.33,211.19);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round,line cap=round] ( 37.09, 76.79) rectangle (500.39,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 32.14,108.69) {-0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 32.14,161.34) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 32.14,213.99) {0.1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 72.79, 67.55) {Jan};

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (179.81, 67.55) {Feb};

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (279.93, 67.55) {Mar};

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (386.95, 67.55) {Apr};

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (490.52, 67.55) {May};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (268.74, 44.09) {$t$ = time};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,171.61) {$\delta$ = FTSE250 price change};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.21, 18.23) -- (234.78, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,255,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (285.47, 18.23) -- (297.04, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (241.72, 15.20) {Predicted};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (303.98, 15.20) {Real};
\end{scope}
\end{tikzpicture}

### 2024-08-15 tex/paper.tex
draft

@@
 In particular, there is a challenge to find a subset of macroeconomic indicators which can most accurately capture stock market dynamics.
-In addition, macroeconomic indicator data is of such low-granularity that it precludes the use of certain deep learning neural network methods for prediction purposes.
+In addition, macroeconomic indicator data is of such low granularity that it precludes the use of certain deep learning neural network methods for prediction purposes.
 
@@
     \item To investigate the short-term and long-term effects of macroeconomic variables on the price of the FTSE 250 index
-    \item To investigate whether machine learning techniques can provide accurate short-term and long-term forecasts of the price of the FTSE 250 index when using a low-granularity dataset.
+    \item To investigate whether machine learning techniques can provide accurate short-term and long-term forecasts of the price of the FTSE 250 index when using a coarse-grain dataset.
 
@@
 This research also employs radial basis function neural networks to predict the price of the FTSE 250 index.
-This artificial deep learning network is particularly useful for tasks like function approximation, pattern recognition, and regression when the dataset is of low-granularity.
+This artificial deep learning network is particularly useful for tasks like function approximation, pattern recognition, and regression when the dataset is of low granularity.
 
@@
 
-In the short-term, which defined as 5 months, 
-as per the optimum lag for the FTSE250 in the ARDL model, the market movements
-are more closely tied to recent deviations in the economy, because theoretically
+In the short-term the market movements
+are more closely tied to recent deviations in the economy. This is because,
+theoretically, 
 the short-run is not exposed to the various unforeseen external shocks and structural changes
 that dilute the relationship between these variables and the stock market in the 
-long-run. Essentially, as the prediction window expands, 
+long-run. As the prediction window expands, 
 more uncertainty and unknowns are introduced, which translates into 
 lower predictive quality. Furthermore, as the time horizon extends into the long-run, fundamental factors 
-begin to matter a lot more. The compound effect of earnings, cash flows and
-managerial efficiencies begins to outweigh the 
-influence of fluctuating macroeconomic sentiment. Thus, in the short-run 
+begin to matter a lot more. The compound effect of earnings, cash flows,
+innovations and managerial efficiencies begins to outweigh the 
+influence of fluctuating macroeconomic variables. Essentially, firms become accustomed 
+to the changed economic environment and over time. Thus, in the short-run 
 a different phenomenon is observed.
@@
 foreign direct and portfolio investment, commodity prices and economic growth.
-An autoregresive model, falling outside the immediate scope of the current study, could also be tried, including the past values of the 
+An autoregressive neural network could also be tried, including the past values of the 
 FTSE250 in the input space to provide a more accurate model. In addition, 
 a more granular data set could be utilised by applying a 
-Brownian bridge algorithm to generate daily macroeconomic data and 
-subsequently testing the RBF-NN against other neural networks 
-such as auto-regressive LSTMs, as well as other machine learning methods suited for 
-forecasting such as Gradient Boosting Machines and Support Vector Machines.
+Brownian bridge algorithm to generate daily macroeconomic data. This would allow 
+the testing of the RBF-NN against other neural networks 
+such as the LSTM. Other machine learning methods suited for 
+forecasting such as Gradient Boosting Machines and Support Vector Machines could 
+also be investigated in future work in this problem domain.
 

### 2024-08-15 tex/paper.tex
draft update

@@
 \begin{abstract}
-    This study investigates the impact of the Consumer Price Index, Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate and the money supply (M3) measure on the FTSE 250 index, and assesses the effectiveness of radial basis function neural networks as a tool to forecast the price of the FTSE 250 index based on the same set of macroeconomic indicators.
-This study shows quantitative estimates for the long-term and short-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 index, and demonstrates the effectiveness of radial basis function neural networks as a short-term forecasting tool for the price of the FTSE 250 index, thus providing an alternative to econometric forecasting models.
+    This study investigates the impact of the Consumer Price Index (CPI), Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate and the money supply (M3) measure on the price of the FTSE 250 index (FTSE250).
+    In addition, it assesses the effectiveness of radial basis function neural networks as a tool to forecast the price of the FTSE 250 Index based on the same set of macroeconomic variables.
+This study finds quantitative estimates for the long-term and short-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 Index, and demonstrates the effectiveness of radial basis function neural networks as a short-term forecasting tool for the price of the FTSE 250 Index, thus providing an alternative to econometric forecasting models.
 
-    \keywords{Macroeconomic indicators; Stock market; FTSE 250 Index; Autoregressive Distributed Lag
-    (ARDL) model, RBF Neural Network; Time-series forecasting}
+    \keywords{macroeconomic indicators, FTSE 250, autoregressive distributed lag model, RBF neural network, time-series forecasting}
 \end{abstract}
@@
 The stock market has played a central role in the economic and 
-political ascent of many nations ever since its inception \citep{levine2002}.  
+political ascent of many nations ever since its inception (\citeboth{levine2002}).  
 However, whenever stock market instabilities have arisen in history they 
@@
 an early example of the inherent dangers within stock markets in the form of 
-speculative bubbles \citep{mississippibubble}. The role of
+speculative bubbles (\citeboth{mississippibubble}). The role of
 the macroeconomy in this crash was central. The French central bank, the 
@@
 value of metal coinage it possessed. This large-scale expansion of the money supply led to 
-inflation burgeoning, reaching $23\%$ at its peak, and devastating France \citep{moen2001john}.  Nonetheless, stock markets also provide essential benefits by allowing companies to 
+inflation burgeoning, reaching $23\%$ at its peak, and devastating the French economy (\citeboth{moen2001john}).  Nonetheless, stock markets also provide essential benefits by allowing companies to 
 raise capital, offering investment opportunities for individuals, and 
@@
 facilitate price discovery and market liquidity, and contribute to government 
-financing through taxes \citep{shahbaz2016}. Effective and sustainable
+financing through taxes (\citeboth{shahbaz2016}). Effective and sustainable
 macroeconomic policy can greatly aid in this beneficial process. Therefore, it is essential 
@@
 In particular, there is a challenge to find a subset of macroeconomic indicators which can most accurately capture stock market dynamics.
-In addition, macroeconomic indicator data is of such low granularity that it precludes the use of certain deep learning neural network methods for prediction purposes.
+In addition, coarse-grain macroeconomic indicator data is often so small in size that it precludes the use of certain deep learning neural network methods for prediction purposes.
 
@@
 \begin{enumerate}
-    \item To investigate how macroeconomic indicators influence the FTSE 250 index
-    \item To investigate the short-term and long-term effects of macroeconomic variables on the price of the FTSE 250 index
-    \item To investigate whether machine learning techniques can provide accurate short-term and long-term forecasts of the price of the FTSE 250 index when using a coarse-grain dataset.
+    \item To investigate how macroeconomic indicators influence the FTSE250
+    \item To investigate the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250 
+    \item To investigate whether machine learning techniques can provide accurate short-term and long-term forecasts of the price of the FTSE250 when using a small, coarse-grain dataset.
 
@@
 By integrating the ARDL model with the ECM, a comprehensive understanding of both the immediate and longer-term relationships between variables is obtained, which is valuable for robust economic analysis and policy-making.
-This research also employs radial basis function neural networks to predict the price of the FTSE 250 index.
-This artificial deep learning network is particularly useful for tasks like function approximation, pattern recognition, and regression when the dataset is of low granularity.
+This research also employs radial basis function neural networks to predict the price of the FTSE250.
+This artificial deep learning network is particularly useful for tasks like function approximation, pattern recognition, and regression when the dataset is small and of low granularity.
 
 The remainder of the paper is organised as follows.
-Section 2 presents a brief review of the literature in this problem space, and explains the rationale behind our selection of macroeconomic variables.
+Section 2 presents a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables.
 Section 3 covers the data utilised in this study.
-Section 4 outlines the ARDL-ECM model, the methodology, and the results achieved describing the effect of the selected macroeconomic variables on the FTSE 250 index.
-Section 5 constructs the radial basis function neural network (RBF-NN) model, and presents the results of forecasting the short-term and long-term forecasts for the price of the FTSE 250 Index price movements based on the selected macroeconomic variables.
+Section 4 outlines the ARDL-ECM model, the methodology, and the results achieved.
+Section 5 constructs the radial basis function neural network (RBF-NN) model, and presents the results of forecasting the short-term and long-term forecasts for the price of the FTSE250 based on the selected macroeconomic variables.
 Section 6 concludes the study.
@@
 error correction model (ECM) to investigate this relationship.
-This approach has been used extensively \citep{QuadriMasih, Plíhal2016,olomu2015} to assess the impact of macroeconomic variables on the stock market.
-Other prominent techniques that have been developed include the Johansen cointegration model \citep{YadavKheraMishra2021,Ozcan2012,ChistiShakeelGanai2020} and the Autoregressive Distributed Lag (ARDL) model \citep{khan2018,demir2019,neifar2023}.
+This approach has been used extensively (\citeboth{QuadriMasih, Plíhal2016,olomu2015}) to assess the impact of macroeconomic variables on the stock market.
+Other prominent techniques that have been developed include the Johansen cointegration model (\citeboth{YadavKheraMishra2021,Ozcan2012,ChistiShakeelGanai2020}) and the Autoregressive Distributed Lag (ARDL) model (\citeboth{khan2018,demir2019,neifar2023}).
 
@@
 \citeboth{demir2019} reported similar findings for the Turkish Stock Exchange.
-In the case of the UK, \citeboth{neifar2023} used a Non-Linear ARDL model to find a significantly negative long-run relationship between interest rates and the FTSE100.
+In the case of the UK, \citeboth{neifar2023} used a Non-Linear ARDL model to find a significantly negative long-run relationship between interest rates and the FTSE 100 Index.
 In addition, the flight-to-quality phenomenon that higher interest rates also encourage a shift from stocks to fixed-income 
@@
 One of the foundational studies on the relationship between the money supply and 
-stock market prices was \citeboth{palmer1970}, who analysed the impact of M1 money supply on the S$\&$P 425 Industrial Stock Index. 
+stock market prices was \citeboth{palmer1970}, who analysed the impact of M1 money supply on the S$\&$P 425 Industrial Stock index. 
 Palmer posited that trends in the nation's money stock could lead the 
@@
 
-The FTSE250 has historically been considered a more closer representation of the domestic economy than the FTSE100.
+The FTSE250 has historically been considered a more closer representation of the domestic economy than the FTSE 100 Index.
 Roughly $50\%$ of the earnings of the FTSE250 are from overseas, similar to the S\&P 500.
-In contrast, nearly $80\%$ of the FTSE$100$'s earnings are from overseas.
+In contrast, nearly $80\%$ of the FTSE 100 Index's earnings are from overseas.
 This is reflected in the constitution of the respective indices, with the 
-FTSE250 historically resembling the structure of the UK economy more closely than the FTSE100 \citep{ftse250history}.
+FTSE250 historically resembling the structure of the UK economy more closely than the FTSE100 (\citeboth{ftse250history}).
 
@@
 with the exception of \citeboth{smit2023longrun}, \citeboth{khatri2024impact} and \citeboth{engstrom2006impact} who investigate the effect of variable changes on the FTSE All Share, an index that aggregates the
-FTSE100, FTSE250 and FTSE SmallCap indices. The impact of the four selected variables on the FTSE250 has not yet been empirically studied and 
+FTSE 100 Index, FTSE 250 Index and FTSE SmallCap Indices. The impact of the four selected variables on the FTSE250 has not yet been empirically studied and 
 warrants further investigation.
@@
 
-There is a large body of literature regarding stock market price forecasting. First, 
-a discussion on the feasibility of such forecasting is presented. Following this, machine learning methods for forecasting are 
+First, 
+a discussion on the feasibility of stock market price forecasting is presented. Following this, machine learning methods for forecasting are 
 discussed.
@@
 Some studies such as \citeboth{rosini2020} even found that the 
-efficiency of the FTSE100 and FTSE250 varied over time. Thus, the empirical findings suggest 
+efficiency of the FTSE 100 Index and FTSE250 varied over time. Thus, the empirical findings suggest 
 that it is not impossible to forecast movements in the 
@@
 In addition to econometric models, machine learning methods, 
-particularly gradient boosting machines \citep{gumelar,Liu2024,Chen2023} and artificial and deep neural networks \citep{chen2015lstm,kara2011ann,long2019deep,nelson2017lstm}, have recently emerged as powerful alternatives capable of overcoming the limitations of econometric models.
-They are able to capture complex relationships in financial data \citep{rossi2018ml}, and 
+particularly gradient boosting machines (\citeboth{gumelar,Liu2024,Chen2023}) and artificial and deep neural networks (\citeboth{chen2015lstm,kara2011ann,long2019deep,nelson2017lstm}), have recently emerged as powerful alternatives capable of overcoming the limitations of econometric models.
+They are able to capture complex relationships in financial data (\citeboth{rossi2018ml}), and 
 are thereby able to provide more accurate forecasts than ordinary statistical models 
-\citep{lapitskaya2021armax}.
+(\citeboth{lapitskaya2021armax}).
 
-Deep learning methods such as Recurrent Neural Networks (RNNs) and Long Short-Term Memory (LSTM) Neural Networks
+Deep learning methods such as Recurrent neural networks (RNN) and Long Short-Term Memory (LSTM) neural networks
 have been extensively used for forecasting time series data, 
-such as stock market volatility \citep{cho2022forecasting,praveenraj2023} and prices \citep{zhang2022lstm,song2023forecasting,dutta2024hybrid}. The drawback of such 
+such as stock market volatility (\citeboth{cho2022forecasting,praveenraj2023}) and prices (\citeboth{zhang2022lstm,song2023forecasting,dutta2024hybrid}). The drawback of such 
 methods is that they are not suitable for smaller datasets and tend to overfit on the training data
-in such datasets \citep{foster1992} or display unstable behaviour during performance \citeboth{lebaron1998}
+in such datasets (\citeboth{foster1992}) or display unstable behaviour during performance (\citeboth{lebaron1998})
 
-Radial Basis Function Neural Networks (RBF-NN) are a type of artificial 
+Radial basis function neural networks are a type of artificial 
 neural network that use radial basis functions at each node in the 
 activation layer of the network.
-The advantage of RBF-NNs lies in their ability to 
-operate effectively in extremely volatile financial time-series 
-environments \citep{cafferata2019}, their strong ability to 
-generalise on such data \citep{sharkawy2020} as well as their efficacy on small datasets \citep{kosarac2022}. 
+The advantage of radial basis function neural networks lies in their ability to 
+operate effectively in volatile financial time-series 
+environments (\citeboth{cafferata2019}), their strong ability to 
+generalise on such data (\citeboth{sharkawy2020}) as well as their efficacy on small datasets (\citeboth{kosarac2022}). 
 \citeboth{esfandyari2016}
-supports this, using a sample size of 70 and 29 to forecast daily NASDAQ 
+supports this, using a sample size of 70 to forecast daily NASDAQ 
 index values using an Artificial Neural Network; they achieved a minimum
@@
 
-RBF-NNs have been used considerably for stock market forecasting. \citeboth{cao2004} 
-used an RBF-NN based on the Optimal Partition Algorithm (OPA) to 
-forecast S$\&$P 500 prices. \citeboth{dass2019} use an RBF-NN, with the Back-propagation 
-Algorithm (BPA) used instead to tune parameters; their results demonstrate 
-that the RBF-NN outperformed a standard feed-forward neural network. 
-\citeboth{ji2014} also found that the RBF-NN was able to predict the value 
-of the Shanghai Composite Index on a 45 day time-horizon using the closing price, with a mean prediction error of 
+Radial basis function neural networks have been used considerably for stock market forecasting. \citeboth{cao2004} 
+used a radial basis function network based on the Optimal Partition Algorithm (OPA) to 
+forecast S$\&$P 500 prices. \citeboth{dass2019} use a radial basis function neural network, utilising the Back-propagation 
+Algorithm (BPA) instead to tune parameters; their results demonstrate 
+that the radial basis function network outperformed a standard feed-forward neural network. 
+\citeboth{ji2014} also found that the radial basis function network was able to predict the value 
+of the Shanghai Composite index on a 45 day time-horizon using the closing price, with a mean prediction error of 
 1.4387$\%$. Recently, \citeboth{abotaleb2024}
@@
 uncertainty index on the quarterly prices of various stock markets including the 
-FTSE100 using an RBF-NN. They found that the model 
-achieved a relative error of 27.2$\%$ in the case of the FTSE100.
+FTSE 100 Index using a radial basis function. They found that the model 
+achieved a relative error of 27.2$\%$ in the case of the FTSE 100 Index.
 
-Whilst \citeboth{abotaleb2024}, used quarterly data and a relatively minor economic indicator, the effectiveness of RBF-NNs in stock market forecasting using 
+Whilst \citeboth{abotaleb2024}, used quarterly data and a relatively minor economic indicator, the effectiveness of radial basis function neural networks in stock market forecasting using 
 major macroeconomic indicators and coarse-grain time-series has not yet been investigated and presents an area for further 
@@
 
-FTSE-250 monthly price data between Oct 1992 to March 2024 has been utilised for this study giving 379 observations.
+Monthly FTSE 250 Index price data between Oct 1992 to March 2024 has been utilised for this study giving 379 observations.
 
@@
 
-During the data collection process, it was necessary to deal with an issue raised by the fact that the Bank of England’s Monetary Policy Committee (MPC) 
+During the data collection process, it was necessary to address an issue raised by the fact that the Bank of England’s Monetary Policy Committee (MPC) 
 does not convene on a monthly basis, and are not obligated to 
@@
 
-Over the period covered by our dataset, the FTSE250 index has 
+Over the period covered by the dataset, the FTSE250 has 
 shown substantial growth, reflecting the structural transition of the UK economy.
@@
 an error correction model (ECM) to determine 
-whether a long-term dependence exists for the FTSE250 and our macroeconomic 
+whether a long-term dependence exists for the FTSE250 and the selected macroeconomic 
 indicators, and if so to quantify the long-run and short-run relationships
@@
 on the variables. This was done so that the coefficients appear as elasticities, 
-which we can interpret easily as percentage changes. Subsequently,
+which can be interpreted easily as percentage changes. Subsequently,
 the level at which the variables are stationary was ascertained using the 
@@
 
-In the ARDL bounds test, the null and alternate hypothesis are:
+In the ARDL bounds test, the null and alternate hypothesis were:
 
@@
 It is also possible that the M3 measure is too broad of a measure and includes overly illiquid components that are not as impactful 
-on the stock market dynamics. In the short-run we see that the money supply does have a 
+on the stock market dynamics. In the short-run it is observed that the money supply does have a 
 significantly negative effect on the stock market, as outlined by \citeboth{olawale2014}. According to \citeboth{sellin2001}, 
@@
 and depress the FTSE250. Moreover, foreign investors may view the stronger Pound as a signal of reduced returns on their investments in the UK, which can lead to decreased investment in the FTSE250. 
-In the short-term, however, before structural shifts can materialise we see that the
+In the short-term, however, before structural shifts can materialise the
 weakening of the Pound has a positive effect. As the USD has been 
@@
 
-The interest rate was also found to have significant negative long-term and short-term effects on the stock market, confirming the findings of the extant literature \citep{alam2009,demir2019,neifar2023, talla2013}.
+The interest rate was also found to have significant negative long-term and short-term effects on the stock market, confirming the findings of \citeboth{alam2009}, \citeboth{demir2019} and \citeboth{neifar2023}
 This can be explained by the fact that higher interest 
@@
 investments more attractive relative to stocks, leading investors to shift 
-their portfolios away from equities and toward safer assets over time \citep{asgharian2016}. 
+their portfolios away from equities and toward safer assets over time (\citeboth{asgharian2016}). 
 In the short term, interest rate hikes immediately decrease the present value of earnings, as per the 
@@
     \end{tikzpicture}
-    \caption{RBF-NN Architecture}
+    \caption{Radial basis function neural network}
     \label{fig:rbf_nn_architecture}
@@
 
-The exact interpolation problem, which forms the motivation for RBF-NNs, 
+The exact interpolation problem, which forms the motivation for radial basis function neural networks, 
 involves mapping a $d$-dimensional input space $\boldsymbol{\mathrm{X}}$ to
@@
 By introducing several modifications to the exact interpolation 
-procedure, the RBF-NN
+procedure, the radial basis function neural network
 model is obtained, as presented in \figref{rbf_nn_architecture}. 
-The RBF-NN mapping takes the following form:
+The network mapping takes the following form:
 
@@
 
-where $\boldsymbol{c_j}$ is a RBF centre and $f$ is the radial basis function \citep{bishop1995}.
+where $\boldsymbol{c_j}$ is a radial basis function centre and $f$ is the radial basis function (\citeboth{bishop1995}).
 
@@
 
-The RBF-NN was utilised for both short-term and long-term 
+The radial basis function neural network was utilised for both short-term and long-term 
 forecasting. For the short-term, 
@@
 The data was normalised using a z-score normalisation, because neural networks work best with data
-that is between a specified range \citep{mendelsohn1993}. Following this 
-it was necessary to program the RBF-NN into Python. The kernel regression approach outlined by 
-\citeboth{bishop1995}, using the GaussianProcessRegressor package in Python was employed to model the RBF-NN.
+that is between a specified range (\citeboth{mendelsohn1993}). Following this 
+it was necessary to program the radial basis function neural network into Python. The kernel regression approach outlined by 
+\citeboth{bishop1995}, using the GaussianProcessRegressor package in Python was employed to model the network.
 
@@
 where $y$ differs from the function values $f(x)$ by noise which is assumed
-to follow an iid Gaussian distribution
+to follow an iid Gaussian distribution (\citeboth{rasmussen2006}). 
 $\mathcal{N}_1$
@@
 which is factored over cases in the training set. This is
-permissible because of the independence assumption \citep{rasmussen2006}.
+permissible because of the independence assumption (\citeboth{rasmussen2006}).
 
@@
 A covariance function or kernel
-is specified according to the characteristics of the data. In our model this 
-is the generalised inverse multiquadric
+is specified according to the characteristics of the data. In this study the
+generalised inverse multiquadric is used.
 
@@
 \end{align}
-Using this, the posterior distribution \eqref{eq:nine} is obtained \citep{rasmussen2006}.
+Using this, the posterior distribution \eqref{eq:nine} is obtained (\citeboth{rasmussen2006}).
 
@@
 All possible $w$, weighted by their probability, are then averaged to obtain the
-most likely functional estimate of the output \citep{rasmussen2006}.
+most likely functional estimate of the output (\citeboth{rasmussen2006}).
 
@@
     \item Length scale: a parameter that controls the influence that each kernel is to have. Analogous to 
-    the width each RBF neuron possesses.
+    the width each radial basis function neuron possesses.
 \end{enumerate}
@@
 Each possible weight vector $w$ represents a different function, explicitly 
-based on the RBF covariance function, that could fit the data. In the 
+based on the radial basis function covariance function, that could fit the data. In the 
 hidden layer the expectation is then taken, which
@@
 prior distribution. From this the posterior weight distributions for the test observations \eqref{eq: testw}
-are calculated, and then the predictive output is found \citep{rasmussen2006}.
+are calculated, and then the predictive output is found (\citeboth{rasmussen2006}).
 
 \begin{align}
-    p(y_{*}|x_{*}, X, y) \, dw = \mathcal{N}(\sigma_{n}^{2} x_{*}^{T}A^{-1}Xy, x_{*}^{T}A^{-1}x_{*}) \label{eq: testw}
+    p(y_{*}|x_{*}, X, y) = \mathcal{N}(\sigma_{n}^{2} x_{*}^{T}A^{-1}Xy, x_{*}^{T}A^{-1}x_{*}) \label{eq: testw}
 \end{align}
@@
 \subsection{Results}
-
-The model returned a mean absolute error of MAE $=0.16$ and 
-$R^2 = 9\%$ for the long-term prediction, providing a 
-$2.4\%$ increase in the $R^2$ obtained by the ARDL-ECM model.
 
@@
 \end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \input{long_term_daily.tex}  
+    \caption{Long-term predicted vs real daily FTSE250 changes}
+    \label{fig:longdaily}
+\end{figure}
+
+The model returned a mean absolute error of MAE $=0.16$ and 
+$R^2 = 9\%$ for the long-term prediction, providing a 
+$2.4\%$ increase in the $R^2$ obtained by the ARDL-ECM model. The results are presented 
+in Figures 2 and 3, against monthly and daily FTSE250 changes respectively.
+
 The interpretation of this is that, over the long-run (4-5 years), the variations in the 
@@
 methods to capture subtle patterns in the data unavailable to econometric models.
-
-\begin{figure}[h!]
-    \centering
-    \input{long_term_daily.tex}  
-    \caption{Long-term predicted vs real daily FTSE250 changes}
-    \label{fig:longdaily}
-\end{figure}
 
@@
 
-\begin{figure}[h!]
-    \centering
-    \input{short_term_daily.tex}  
-    \caption{Short-term predicted vs real daily FTSE250 changes}
-    \label{fig:shortdaily}
-\end{figure}
-
 The short-term prediction yielded MAE = $0.07$ and 
-$R^2 = 61\%$ for a 5-month time horizon.
+$R^2 = 61\%$ for a 5-month time horizon. The results are presented in Figures 4 
+and 5, against monthly and daily FTSE250 changes respectively.
 The significantly better performance of the model in the short-run
@@
 
+\begin{figure}[h!]
+    \centering
+    \input{short_term_daily.tex}  
+    \caption{Short-term predicted vs real daily FTSE250 changes}
+    \label{fig:shortdaily}
+\end{figure}
+
 \section{Conclusion}
@@
 M3 money supply market activity on
-the FTSE250 index over the October 1993 – April 2024 period.
+the FTSE250 over the October 1993 – April 2024 period.
 The results suggest a cointegrated long-run
@@
 The study also investigated the efficacy of machine learning methods in 
-the field of stock-econometric forecasting. The RBF-NN performed well in
+the field of stock-econometric forecasting. The radial basis function network performed well in
 short-term forecasting of the FTSE250, and offers investors with an accurate forecasting tool, 
@@
 Brownian bridge algorithm to generate daily macroeconomic data. This would allow 
-the testing of the RBF-NN against other neural networks 
+the testing of the radial basis function neural network against other neural networks 
 such as the LSTM. Other machine learning methods suited for 

### 2024-08-15 tex/paper.tex
draft update

@@
 
-\title{Investigating the Impact of Macroeconomic Indicators 
+\title{Investigating the Impact of Macroeconomic Indicatorspassword
+
 on the FTSE 250 Index Using Machine Learning Methods}
@@
 \begin{abstract}
-    This study investigates the impact of the Consumer Price Index (CPI), Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate and the money supply (M3) measure on the price of the FTSE 250 index (FTSE250).
+    This study investigates the impact of the Consumer Price Index (CPI), Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate and the money supply (M3) measure on the price of the FTSE 250 Index (FTSE250).
     In addition, it assesses the effectiveness of radial basis function neural networks as a tool to forecast the price of the FTSE 250 Index based on the same set of macroeconomic variables.
@@
 
-    \keywords{macroeconomic indicators, FTSE 250, autoregressive distributed lag model, RBF neural network, time-series forecasting}
+    \keywords{macroeconomic indicators, FTSE 250, autoregressive distributed lag model, radial basis function neural network, time-series forecasting}
 \end{abstract}
@@
 In particular, there is a challenge to find a subset of macroeconomic indicators which can most accurately capture stock market dynamics.
-In addition, coarse-grain macroeconomic indicator data is often so small in size that it precludes the use of certain deep learning neural network methods for prediction purposes.
+In addition, certain macroeconomic indicator data is often so small in size that it precludes the use of certain deep learning neural network methods for prediction purposes.
 
@@
 This research also employs radial basis function neural networks to predict the price of the FTSE250.
-This artificial deep learning network is particularly useful for tasks like function approximation, pattern recognition, and regression when the dataset is small and of low granularity.
+This artificial deep learning network is particularly useful for tasks like function approximation, pattern recognition, and regression when the dataset is small.
 
@@
 This is reflected in the constitution of the respective indices, with the 
-FTSE250 historically resembling the structure of the UK economy more closely than the FTSE100 (\citeboth{ftse250history}).
+FTSE250 historically resembling the structure of the UK economy more closely than the FTSE 100 Index (\citeboth{ftse250history}).
 
@@
 with the exception of \citeboth{smit2023longrun}, \citeboth{khatri2024impact} and \citeboth{engstrom2006impact} who investigate the effect of variable changes on the FTSE All Share, an index that aggregates the
-FTSE 100 Index, FTSE 250 Index and FTSE SmallCap Indices. The impact of the four selected variables on the FTSE250 has not yet been empirically studied and 
+FTSE 100 Index, FTSE 250 Index and FTSE SmallCap indices. The impact of the four selected variables on the FTSE250 has not yet been empirically studied and 
 warrants further investigation.
@@
 
-First, 
-a discussion on the feasibility of stock market price forecasting is presented. Following this, machine learning methods for forecasting are 
+A discussion on the feasibility of stock market price forecasting is presented. Following this, machine learning methods for forecasting are 
 discussed.
@@
 Some studies such as \citeboth{rosini2020} even found that the 
-efficiency of the FTSE 100 Index and FTSE250 varied over time. Thus, the empirical findings suggest 
+efficiency of the FTSE 100 and FTSE 250 indices varied over time. Thus, the empirical findings suggest 
 that it is not impossible to forecast movements in the 

### 2024-08-15 tex/paper.tex
draft update 3

@@
 \usepackage{parskip}
+\usepackage{float}
 \usepackage{tablefootnote}
@@
     colorlinks=true,      
-    linkcolor=green,       
+    linkcolor=red,       
     citecolor=blue,       
@@
 
-\title{Investigating the Impact of Macroeconomic Indicatorspassword
+\title{Investigating the Impact of Macroeconomic Indicators
 
@@
 One of the foundational studies on the relationship between the money supply and 
-stock market prices was \citeboth{palmer1970}, who analysed the impact of M1 money supply on the S$\&$P 425 Industrial Stock index. 
+stock market prices was \citeboth{palmer1970}, who analysed the impact of M1 money supply on the S$\&$P 425 Industrial Stock Index. 
 Palmer posited that trends in the nation's money stock could lead the 
@@
 supports this, using a sample size of 70 to forecast daily NASDAQ 
-index values using an Artificial Neural Network; they achieved a minimum
+Index values using an Artificial Neural Network; they achieved a minimum
 test $R^2$ of 83$\%$.
@@
 \citeboth{ji2014} also found that the radial basis function network was able to predict the value 
-of the Shanghai Composite index on a 45 day time-horizon using the closing price, with a mean prediction error of 
+of the Shanghai Composite Index on a 45 day time-horizon using the closing price, with a mean prediction error of 
 1.4387$\%$. Recently, \citeboth{abotaleb2024}
@@
 
-\begin{figure}[h!]
-    \centering
-    \input{long_term_monthly.tex}  
-    \caption{Long-term predicted vs real monthly FTSE250 changes}
-    \label{fig:longmonthly}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \input{long_term_daily.tex}  
-    \caption{Long-term predicted vs real daily FTSE250 changes}
-    \label{fig:longdaily}
-\end{figure}
-
 The model returned a mean absolute error of MAE $=0.16$ and 
@@
 $2.4\%$ increase in the $R^2$ obtained by the ARDL-ECM model. The results are presented 
-in Figures 2 and 3, against monthly and daily FTSE250 changes respectively.
+in Figures \ref{fig:longmonthly} and \ref{fig:longdaily}, against monthly and daily FTSE250 changes respectively.
 
@@
 
-\begin{figure}[h!]
-    \centering
-    \input{short_term_monthly.tex}  
-    \caption{Short-term predicted vs real monthly FTSE250 changes}
-    \label{fig:shortmonthly}
-\end{figure}
-
 The short-term prediction yielded MAE = $0.07$ and 
-$R^2 = 61\%$ for a 5-month time horizon. The results are presented in Figures 4 
-and 5, against monthly and daily FTSE250 changes respectively.
+$R^2 = 61\%$ for a 5-month time horizon. The results are presented in Figures \ref{fig:shortmonthly} 
+and \ref{fig:shortdaily}, against monthly and daily FTSE250 changes respectively.
 The significantly better performance of the model in the short-run
@@
 
+\clearpage
+
+\begin{figure}[H]
+    \centering
+    \input{long_term_monthly.tex}  
+    \caption{Long-term predicted vs real monthly FTSE250 changes}
+    \label{fig:longmonthly}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \input{long_term_daily.tex}  
+    \caption{Long-term predicted vs real daily FTSE250 changes}
+    \label{fig:longdaily}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[H]
+    \centering
+    \input{short_term_monthly.tex}  
+    \caption{Short-term predicted vs real monthly FTSE250 changes}
+    \label{fig:shortmonthly}
+\end{figure}
+
 \begin{figure}[h!]
@@
 impacts of the interest rate, USD/GBP
-exchange rate, the consumer price index, and 
+exchange rate, the Consumer Price Index, and 
 M3 money supply market activity on

### 2024-08-15 tex/paper.tex
title change

@@
 
-on the FTSE 250 Index Using Machine Learning Methods}
+on the FTSE 250 Index and the Use of Machine Learning Methods to Forecast the Price of the FTSE 250 Index}
 \author{Mateen Khan}
@@
 procedure, the radial basis function neural network
-model is obtained, as presented in \figref{rbf_nn_architecture}. 
+model is obtained, as presented in Figure \ref{fig:rbf_nn_architecture}. 
 The network mapping takes the following form:

### 2024-08-22 tex/paper.tex
draft update 4
abstract revised

@@
     In addition, it assesses the effectiveness of radial basis function neural networks as a tool to forecast the price of the FTSE 250 Index based on the same set of macroeconomic variables.
-This study finds quantitative estimates for the long-term and short-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 Index, and demonstrates the effectiveness of radial basis function neural networks as a short-term forecasting tool for the price of the FTSE 250 Index, thus providing an alternative to econometric forecasting models.
-
+This study finds quantitative estimates for the long-term and short-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 Index.
+It also demonstrates the effectiveness of radial basis function neural networks as a short-term forecasting tool for the price of the FTSE 250 Index.
     \keywords{macroeconomic indicators, FTSE 250, autoregressive distributed lag model, radial basis function neural network, time-series forecasting}
@@
 
+\begin{figure}[H]
+    \centering
+    \input{long_term_monthly.tex}  
+    \caption{Long-term predicted vs real monthly FTSE250 changes}
+    \label{fig:longmonthly}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \input{long_term_daily.tex}  
+    \caption{Long-term predicted vs real daily FTSE250 changes}
+    \label{fig:longdaily}
+\end{figure}
+
 The interpretation of this is that, over the long-run (4-5 years), the variations in the 
@@
 
+\begin{figure}[H]
+    \centering
+    \input{short_term_monthly.tex}  
+    \caption{Short-term predicted vs real monthly FTSE250 changes}
+    \label{fig:shortmonthly}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \input{short_term_daily.tex}  
+    \caption{Short-term predicted vs real daily FTSE250 changes}
+    \label{fig:shortdaily}
+\end{figure}
+
 In the short-term the market movements
@@
 a different phenomenon is observed.
-
-\clearpage
-
-\begin{figure}[H]
-    \centering
-    \input{long_term_monthly.tex}  
-    \caption{Long-term predicted vs real monthly FTSE250 changes}
-    \label{fig:longmonthly}
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \input{long_term_daily.tex}  
-    \caption{Long-term predicted vs real daily FTSE250 changes}
-    \label{fig:longdaily}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[H]
-    \centering
-    \input{short_term_monthly.tex}  
-    \caption{Short-term predicted vs real monthly FTSE250 changes}
-    \label{fig:shortmonthly}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \input{short_term_daily.tex}  
-    \caption{Short-term predicted vs real daily FTSE250 changes}
-    \label{fig:shortdaily}
-\end{figure}
 

### 2024-08-22 tex/paper.tex
conclusion revised

@@
 should consider the factors in the study, and to raise the stock
-market's value, a lower interest rate regime should be adopted and 
-the Pound should be made more competitive.
+market's value, a lower interest rate regime should be adopted. This will 
+stimulate consumer activity, reduce operating costs for firms, and make 
+stocks more attractive investments, thereby increasing the stock market value. 
+In addition, the Pound should be increased in value to lower import costs 
+and increase the real value of firm earnings, both of which should drive up 
+the value of the stock market. The combined effect of these changes will 
+create a more stimulating business environment by increasing the capital 
+available to firms for research and development, and more efficient 
+capital and labour, in turn 
+driving economic growth and prosperity. However, 
+as with all economic policy shifts, these changes may have 
+unintended consequences on other areas of the economy, potentially 
+undermining any positive changes. Any policy changes must cautiously be made 
+with this at the forefront.
 
@@
 such as the LSTM. Other machine learning methods suited for 
-forecasting such as Gradient Boosting Machines and Support Vector Machines could 
+forecasting such as Gradient Boosting Machines could 
 also be investigated in future work in this problem domain.

### 2024-08-22 tex/paper.tex
data section revised

@@
 
-During the data collection process, it was necessary to address an issue raised by the fact that the Bank of England’s Monetary Policy Committee (MPC) 
+For the interest rate data, it was necessary to address an issue raised by the fact that the Bank of England’s Monetary Policy Committee (MPC) 
 does not convene on a monthly basis, and are not obligated to 

### 2024-08-22 tex/paper.tex
Introduction revised

@@
 
+\subsection{Context}
+
 The stock market has played a central role in the economic and 
@@
 speculative bubbles (\citeboth{mississippibubble}). The role of 
-the macroeconomy in this crash was central. The French central bank, the 
-\textit{Banque Royale}, aimed to convert its vast national debt into shares through the issuance of large 
-amounts of government bonds and paper notes. This considerably increased the money supply, in turn 
-greatly facilitating the ease of obtaining credit by individuals, many of whom invested in the Mississippi Company.
-A mass sell-off began after the \textit{Banque Royale} announced the number of paper notes exceeded the 
-value of metal coinage it possessed. This large-scale expansion of the money supply led to 
-inflation burgeoning, reaching $23\%$ at its peak, and devastating the French economy (\citeboth{moen2001john}).  Nonetheless, stock markets also provide essential benefits by allowing companies to 
+macroeconomic policy in this crash was central, and the effects of the crash 
+devastated the French economy through rampant inflation. Nonetheless, stock markets also provide essential benefits by allowing companies to 
 raise capital, offering investment opportunities for individuals, and 
@@
 help investors anticipate future economic conditions and corporate profits
-and consequently investment returns. Positive indicators tend to boost stock prices, while 
-negative indicators can lead to declines in stock prices.
+and consequently investment returns.
 
@@
 
-A key task which economic policy makers and investors undertake is to 
-understand and predict how economic changes influence stock market behaviour, 
-which is critical for making informed policy and investment decisions. 
-This modelling of the impact of macroeconomic indicators on stock markets is typically achieved by a variety of methods: statistical analysis, econometric modelling, scenario planning, and advanced machine learning techniques.
-However, it is fraught with challenges that require careful consideration, advanced techniques, and a deep understanding of both economics and financial markets.
-In particular, there is a challenge to find a subset of macroeconomic indicators which can most accurately capture stock market dynamics.
-In addition, certain macroeconomic indicator data is often so small in size that it precludes the use of certain deep learning neural network methods for prediction purposes.
+Modelling the impact of macroeconomic indicators on stock markets 
+typically involves a variety of methods: statistical analysis, 
+econometric modelling, scenario planning, and advanced machine learning 
+techniques. However, this process contains challenges that require careful 
+consideration. In particular, there is a challenge to 
+find a subset of macroeconomic indicators which are most relevant in 
+stock market dynamics. Additionally, there is the task of
+choosing the most appropriate model based on the 
+research objectives and data to ensure reliable
+and accurate results.
+
+\subsection{Research objectives}
 
@@
 
+\subsection{Organisation of paper}
 The remainder of the paper is organised as follows.

### 2024-08-23 tex/paper.tex
introduction revised

@@
 
-on the FTSE 250 Index and the Use of Machine Learning Methods to Forecast the Price of the FTSE 250 Index}
+on the FTSE 250 Index Using Machine Learning Methods}
 \author{Mateen Khan}
@@
 \begin{abstract}
-    This study investigates the impact of the Consumer Price Index (CPI), Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate and the money supply (M3) measure on the price of the FTSE 250 Index (FTSE250).
+    This study investigates the impact of the Consumer Price Index (CPI), Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate, and the money supply (M3) measure on the price of the FTSE 250 Index (FTSE250).
     In addition, it assesses the effectiveness of radial basis function neural networks as a tool to forecast the price of the FTSE 250 Index based on the same set of macroeconomic variables.
-This study finds quantitative estimates for the long-term and short-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 Index.
+This study finds quantitative estimates for the short-term and long-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 Index.
 It also demonstrates the effectiveness of radial basis function neural networks as a short-term forecasting tool for the price of the FTSE 250 Index.
@@
 \section{Introduction}
-
-\subsection{Context}
 
@@
 macroeconomic policy in this crash was central, and the effects of the crash 
-devastated the French economy through rampant inflation. Nonetheless, stock markets also provide essential benefits by allowing companies to 
+devastated the economy through rampant inflation. Nonetheless, stock markets also provide essential benefits by allowing companies to 
 raise capital, offering investment opportunities for individuals, and 
-supporting economic growth through efficient capital allocation. 
-They enhance corporate governance through transparency and accountability, 
-facilitate price discovery and market liquidity, and contribute to government 
-financing through taxes (\citeboth{shahbaz2016}). Effective and sustainable
+supporting economic growth through efficient capital allocation (\citeboth{shahbaz2016}). Effective, sustainable
 macroeconomic policy can greatly aid in this beneficial process. Therefore, it is essential 
@@
 The sensitivity of the stock market to macroeconomic indicators is also of great 
-relevance to investors. Macroeconomic indicators provide 
-insights into the overall health and direction of the 
-economy, which affect investor sentiment and in turn influence market trends.
-Changes in the macroeconomy have an effect on both consumers and firms 
-which impacts their profitability, because these signals 
-help investors anticipate future economic conditions and corporate profits
-and consequently investment returns.
+relevance to investors. Changes in macroeconomic indicators have an effect on both consumer preferences
+and business costs which impacts the profitability of firms. This 
+leads to discrepancies between the valuation of firms by investors and 
+stock prices, which in turn drives trends in the stock 
+market price.
 
@@
 indicators on the stock market as well as assessing the use of these indicators 
-to forecast stock prices over the long-term and short-term, both useful to 
+to forecast stock market prices over the long-term and short-term, both useful to 
 economic policy makers and stock market investors respectively. 
+Modelling of this kind typically involves choosing 
+the most appropriate method from a variety: statistical analysis, 
+econometric modelling, scenario planning, and machine learning 
+techniques. Also, careful economic analysis is required to select
+the macroeconomic indicators most relevant in 
+the dynamics of a nation's stock market.
 
-Modelling the impact of macroeconomic indicators on stock markets 
-typically involves a variety of methods: statistical analysis, 
-econometric modelling, scenario planning, and advanced machine learning 
-techniques. However, this process contains challenges that require careful 
-consideration. In particular, there is a challenge to 
-find a subset of macroeconomic indicators which are most relevant in 
-stock market dynamics. Additionally, there is the task of
-choosing the most appropriate model based on the 
-research objectives and data to ensure reliable
-and accurate results.
-
-\subsection{Research objectives}
-
-This study aims to explore the relationship between macroeconomic indicators and stock markets 
-through modelling. Its primary objective is to address questions such as how specific 
-macroeconomic indicators influence stock market prices in major economies and whether machine 
+The objective of this study is to address how specific 
+macroeconomic indicators influence stock market prices in a developed economy, and whether machine 
 learning models can substitute traditional econometric models in predicting stock market prices. 
-Ultimately, it aims to investigate three aspects of modelling the impact of macroeconomic 
+Ultimately, it aims to investigate two aspects of the influence of macroeconomic 
 indicators on stock markets: 
 \begin{enumerate}
-    \item To investigate how macroeconomic indicators influence the FTSE250
-    \item To investigate the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250 
+    \item To examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250 
     \item To investigate whether machine learning techniques can provide accurate short-term and long-term forecasts of the price of the FTSE250 when using a small, coarse-grain dataset.
@@
 
-In order to analyse the dynamic relationship between variables in time series data, this study will employ the Autoregressive Distributed Lag (ARDL) model in conjunction with an Error Correction Model (ECM).
-By integrating the ARDL model with the ECM, a comprehensive understanding of both the immediate and longer-term relationships between variables is obtained, which is valuable for robust economic analysis and policy-making.
-This research also employs radial basis function neural networks to predict the price of the FTSE250.
-This artificial deep learning network is particularly useful for tasks like function approximation, pattern recognition, and regression when the dataset is small.
+In order to analyse the dynamic relationship between variables in 
+time series data, this study will employ the Autoregressive Distributed 
+Lag (ARDL) model in conjunction with an Error Correction Model (ECM). 
+This research also employs a radial basis function neural network (RBFNN) to 
+predict the price movements of the FTSE250.
 
-\subsection{Organisation of paper}
 The remainder of the paper is organised as follows. 
-Section 2 presents a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables.
-Section 3 covers the data utilised in this study.
-Section 4 outlines the ARDL-ECM model, the methodology, and the results achieved.
-Section 5 constructs the radial basis function neural network (RBF-NN) model, and presents the results of forecasting the short-term and long-term forecasts for the price of the FTSE250 based on the selected macroeconomic variables.
-Section 6 concludes the study.
+Section 2 presents a brief review of the literature in this problem space, 
+and explains the rationale behind the selection of macroeconomic variables. 
+Section 3 covers the data utilised in this study. Section 4 outlines the 
+ARDL-ECM model and the results achieved. 
+Section 5 constructs the RBFNN 
+model, and presents the forecasting results. Section 6 concludes the study.
 

### 2024-08-24 tex/paper.tex
introduction updated

@@
 
-on the FTSE 250 Index Using Machine Learning Methods}
+on the FTSE 250 Index Using Statistical \& Machine Learning Methods}
 \author{Mateen Khan}
@@
 
-The stock market has played a central role in the economic and 
-political ascent of many nations ever since its inception (\citeboth{levine2002}).  
-However, whenever stock market instabilities have arisen in history they 
-have had far-reaching destabilising effects. The Mississippi Bubble was 
-an early example of the inherent dangers within stock markets in the form of 
-speculative bubbles (\citeboth{mississippibubble}). The role of 
-macroeconomic policy in this crash was central, and the effects of the crash 
-devastated the economy through rampant inflation. Nonetheless, stock markets also provide essential benefits by allowing companies to 
-raise capital, offering investment opportunities for individuals, and 
-supporting economic growth through efficient capital allocation (\citeboth{shahbaz2016}). Effective, sustainable
-macroeconomic policy can greatly aid in this beneficial process. Therefore, it is essential 
-for economic policymakers to understand the impact of macroeconomic changes 
-on the stock market.
+The stock market is a significant component of the financial system and 
+plays a crucial role in the broader economy. Consequently, economic policy 
+makers take the stock market into account when making fiscal policy 
+decisions using a combination of quantitative and qualitative methods. 
+These methods such as econometric modelling and forecasting, and empirical 
+analysis allow them to understand the short-term and long-term effects of 
+macroeconomic indicators such as taxation rates, interest rates and exchange rates,
+on the wider economy through the stock market.
 
-The sensitivity of the stock market to macroeconomic indicators is also of great 
-relevance to investors. Changes in macroeconomic indicators have an effect on both consumer preferences
-and business costs which impacts the profitability of firms. This 
-leads to discrepancies between the valuation of firms by investors and 
-stock prices, which in turn drives trends in the stock 
-market price.
+Macroeconomic indicators can be effectively used in machine learning 
+methods to forecast stock prices by serving as input features that capture 
+the economic environment’s influence on market behaviour. By carefully 
+selecting the most relevant indicators and preprocessing data, 
+as well as choosing appropriate models, machine learning can enhance the accuracy of stock price forecasts (\citeboth{Sarma2023,Prasad2023}).
 
-The significance of this study is to quantify the effects of macroeconomic 
-indicators on the stock market as well as assessing the use of these indicators 
-to forecast stock market prices over the long-term and short-term, both useful to 
-economic policy makers and stock market investors respectively. 
-Modelling of this kind typically involves choosing 
-the most appropriate method from a variety: statistical analysis, 
-econometric modelling, scenario planning, and machine learning 
-techniques. Also, careful economic analysis is required to select
-the macroeconomic indicators most relevant in 
-the dynamics of a nation's stock market.
+The significance of this study is to quantify the effects of macroeconomic indicators on the stock market as well as assessing the use of these indicators to forecast stock market prices over the short-term and long-term.
+In addition, the study seeks to understand which macroeconomic indicators are most relevant in capturing the dynamics of a stock market to aid forecasting.
 
-The objective of this study is to address how specific 
-macroeconomic indicators influence stock market prices in a developed economy, and whether machine 
-learning models can substitute traditional econometric models in predicting stock market prices. 
-Ultimately, it aims to investigate two aspects of the influence of macroeconomic 
-indicators on stock markets: 
+In particular, the objectives of this study are to address how specific macroeconomic indicators influence stock market prices in a developed economy, and whether machine learning models can substitute traditional econometric models in predicting stock market prices.
+It aims to:
 \begin{enumerate}
-    \item To examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250 
-    \item To investigate whether machine learning techniques can provide accurate short-term and long-term forecasts of the price of the FTSE250 when using a small, coarse-grain dataset.
-
+    \item examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250
+    \item investigate whether machine learning techniques can provide accurate short-term and long-term forecasts of the price of the FTSE250 when using a small, coarse-grain dataset.
 \end{enumerate}
 
-In order to analyse the dynamic relationship between variables in 
-time series data, this study will employ the Autoregressive Distributed 
-Lag (ARDL) model in conjunction with an Error Correction Model (ECM). 
-This research also employs a radial basis function neural network (RBFNN) to 
-predict the price movements of the FTSE250.
+In order to analyse the dynamic relationship between variables in time series data, this study
+will employ the Autoregressive Distributed Lag (ARDL) model in conjunction with an Error
+Correction Model (ECM). This research also employs a radial basis function neural network
+(RBFNN) to predict the price movements of the FTSE250.
 
-The remainder of the paper is organised as follows. 
-Section 2 presents a brief review of the literature in this problem space, 
-and explains the rationale behind the selection of macroeconomic variables. 
-Section 3 covers the data utilised in this study. Section 4 outlines the 
-ARDL-ECM model and the results achieved. 
-Section 5 constructs the RBFNN 
-model, and presents the forecasting results. Section 6 concludes the study.
+The remainder of this paper will be organised as follows. 
+Section 2 provides a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables.
+Section 3 presents the methodology of the paper. Section 4 discusses the results and Section 5 concludes
+the study.
 
 \section{Literature Review}
+\label{sec:lit}
 

### 2024-08-24 tex/paper.tex
intro revised

@@
 
-on the FTSE 250 Index Using Statistical \& Machine Learning Methods}
+on the FTSE 250 Index Using Statistical and Machine Learning Methods}
 \author{Mateen Khan}
@@
 analysis allow them to understand the short-term and long-term effects of 
-macroeconomic indicators such as taxation rates, interest rates and exchange rates,
-on the wider economy through the stock market.
+macroeconomic indicators
+on the wider economy through the stock market (\citeboth{DemirgucKunt1996}).
 
@@
 selecting the most relevant indicators and preprocessing data, 
-as well as choosing appropriate models, machine learning can enhance the accuracy of stock price forecasts (\citeboth{Sarma2023,Prasad2023}).
+as well as choosing appropriate models, machine learning can enhance the accuracy of stock price forecasts (\citeboth{Prasad2023}).
 
@@
     \item examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250
-    \item investigate whether machine learning techniques can provide accurate short-term and long-term forecasts of the price of the FTSE250 when using a small, coarse-grain dataset.
+    \item investigate whether machine learning techniques can provide accurate short-term and long-term forecasts of the price of the FTSE250 when using a small dataset.
 \end{enumerate}

### 2024-08-24 tex/paper.tex
literature review revised

@@
 The remainder of this paper will be organised as follows. 
-Section 2 provides a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables.
+Section \ref{sec:lit} provides a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables.
 Section 3 presents the methodology of the paper. Section 4 discusses the results and Section 5 concludes
@@
 
-The Fisher hypothesis states that nominal interest rates adjust to expected 
+The impact of inflation on stock market prices has been a subject of debate. The Fisher hypothesis states that nominal interest rates adjust to expected 
 inflation as firms correctly anticipate inflation 
@@
 
+The interest rate has been confirmed to have a definite effect on the 
+stock market price by the findings of \citeboth{ChenRollRoss1986} and later research. 
 Interest rates, being the cost of borrowing, directly impact firms by increasing financing costs, reducing profitability, and lowering growth prospects.
@@
 
-Classical economic theory explains the impact exchange rates have on stock prices through the portfolio balance model.
-The portfolio balance model indicates that expected currency depreciations lead investors to shift from domestic to foreign assets, decreasing demand for domestic stocks and lowering their prices.
-\citeboth{branson1977} established a bidirectional relation between the exchange rate and stock prices. This was confirmed
-in the case of the Turkish Stock Exchange by \citeboth{aydemir2009}. 
+Economic theory posits that the exchange rate, through the portfolio 
+balance model and flow model, has a definite impact on stock prices.
+The portfolio balance model indicates that expected currency depreciations lead investors to shift from domestic to foreign assets, decreasing demand for domestic stocks and lowering their prices (\citeboth{branson1977}).
+In addition, variations in the exchange rate affect future cash flows, which in turn affect the present value of a stock and thus stock prices (\citeboth{dornbusch1980}).
+This was confirmed
+in the case of the US and Turkish Stock Exchanges by \citeboth{aggarwal1981} and \citeboth{aydemir2009}, respectively. 
 Moreover, stock markets often contain multinational firms involved in international transactions.
@@
 
-One of the foundational studies on the relationship between the money supply and 
-stock market prices was \citeboth{palmer1970}, who analysed the impact of M1 money supply on the S$\&$P 425 Industrial Stock Index. 
-Palmer posited that trends in the nation's money stock could lead the 
+The nature of the impact of the money supply on the stock market price is also 
+in debate. \citeboth{palmer1970} posited that trends in the nation's money stock could lead the 
 private sector to adjust wealth portfolios, resulting in 
@@
 FTSE250 historically resembling the structure of the UK economy more closely than the FTSE 100 Index (\citeboth{ftse250history}).
+
+Initially, the FTSE 250's composition was broadly distributed across 
+various industrial and service sectors. Roughly $47\%$ of its constituents 
+were involved in the industrial and utilities sectors, reflecting the UK's 
+slower transition from an industrial-based to a service-based economy. 
+Over time, the FTSE250 has evolved to align more closely with the 
+structural transformation of the UK economy. 
+By 2022, 73.39$\%$ of the index was comprised of service firms, 
+with a significant 43.37$\%$ of these being financial services companies (\citeboth{ftse250history}) .
+
+As a result, the relevance of the selected macroeconomic indicators for the FTSE250 
+has increased. The M3 money supply and interest rates significantly 
+impact the ability of financial firms 
+to provide loans, facilitate investments, and manage assets, thereby affecting their profitability. 
+Financial services firms also play a crucial role in supporting the 
+operations of firms in all sectors of the economy, meaning 
+changes in the interest rate and money supply affects their profitability too. 
+The CPI is also of critical importance across all sectors, 
+as inflation, measured by the CPI, erodes the purchasing power of money. 
+This erosion diminishes consumer spending and negatively affects the 
+profitability of service-oriented companies that depend on consumer 
+expenditure. The exchange rate is especially relevant for transnational 
+firms within the FTSE250. The profits of these companies, which make up a substantial 
+portion of the index, are directly impacted by fluctuations in exchange 
+rates that affect the value of their earnings.
 

### 2024-08-25 tex/paper.tex
structure revised + final minor edits

@@
 Section \ref{sec:lit} provides a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables.
-Section 3 presents the methodology of the paper. Section 4 discusses the results and Section 5 concludes
+Section \ref{sec:meth} presents the methodology of the paper. Section 4 discusses the results and Section 5 concludes
 the study.
@@
 \subsubsection{Machine Learning Methods for Forecasting}
+\label{subsubsec: fore}
 
@@
 
-\section{Data}
+\section{Methodology}
+\label{sec:meth}
 
-Monthly FTSE 250 Index price data between Oct 1992 to March 2024 has been utilised for this study giving 379 observations.
+\subsection{Data}
+Monthly FTSE 250 Index price data between Oct 1992 to March 2024 has been utilised for this study giving us a dataset with 379 observations.
 
@@
 does not convene on a monthly basis, and are not obligated to 
-announce a rate on a regular basis. To address the resulting irregularities the data was adjusted using \algoref{interest_rate_adjustment}.
+announce a rate on a regular basis. To address the resulting irregularities the data was adjusted using \ref{alg:interest_rate_adjustment}.
 
@@
 Interest rates, have varied widely from $0.1\%$ to $8.4\%$, with a standard deviation of $2.52\%$, reflecting the 
-increasing role played by monetary policy in the UK since May 1997 when the Bank of England was given operational responsibility for setting interest rates \citep{king1997changes}.
+increasing role played by monetary policy in the UK since May 1997 when the Bank of England was given operational responsibility for setting interest rates (\citeboth{king1997changes}).
 The Bank of England was forced to adjust interest rates
 several times in response to worsening economic conditions, examples being the steep rate cuts during the 2008 financial crisis to stimulate the economy and the gradual rate hikes in response to 
-subsequent post-COVID inflationary pressures.
+post-COVID inflationary pressures.
 
@@
 
-\section{ARDL Model}
+\subsection{ARDL-ECM Model}
 
-Based on the empirical research and characteristics of the UK economy,
-the stock market function is defined as:
+The study uses an autoregressive distributed lag (ARDL) model in conjunction with 
+an error correction model (ECM) to examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250
+and to quantify these relationships. Prior to constructing the model a natural logarithmic transform was applied
+on the variables. This was done so that the coefficients can be interpreted easily as percentage changes.
+
+Based on the selected macroeconomic indicators,
+the stock market function was defined as:
 
@@
 
-The study uses an autoregressive distributed lag (ARDL) model in conjunction with 
-an error correction model (ECM) to determine 
-whether a long-term dependence exists for the FTSE250 and the selected macroeconomic 
-indicators, and if so to quantify the long-run and short-run relationships
-between the former and the latter.
-
-\subsection{Methodology}
-
-Prior to constructing the ARDL model a natural logarithmic transform was applied
-on the variables. This was done so that the coefficients appear as elasticities, 
-which can be interpreted easily as percentage changes. Subsequently,
-the level at which the variables are stationary was ascertained using the 
-Augmented Dickey-Fuller (ADF) unit root test. The optimum lag for the variables in the ARDL model was then determined using the Schwarz
-Information Criterion.
+The ARDL model specifies the dependent variable as a function of its own 
+lagged values and the lagged values of independent variables. 
+The variables are permitted to be of differing stationarity. The stationarity 
+of the variables was tested using the Augmented Dickey-Fuller 
+Unit Root Test found within the tseries package in R. The optimal lags for each of the 
+variables were calculated using the Schwarz Information Criterion, contained
+within the auto\_ardl function in the ARDL package in R.
 
@@
     \Delta \text{FTSE}_{250_t} &= \alpha_0 + \sum_{i=1}^{p} \alpha_i \Delta \text{FTSE}_{250_{t-i}} + \sum_{j_{1}=0}^{q_1} \beta_{1j_{1}} \Delta \text{CPI}_{t-j_{1}} + \sum_{j_{2}=0}^{q_2} \beta_{2j_{2}} \Delta \text{INT}_{t-j_{2}} \\
-                               & \sum_{j_{3}=0}^{q_3} \beta_{3j_{3}} \Delta \text{EXCHG}_{t-j_{3}} + \sum_{j_{4}=0}^{q_4} \beta_{4j_{4}} \Delta \text{M3}_{t-j_{4}} + \varepsilon_t + \phi_{1} FTSE_{250_{t-1}} \\
-                               & + \phi_{2} CPI_{t-1} + \phi_{3} INT_{t-1} +\phi_4 EXCHG_{t-1} + \phi_5 M3_{t-1}
+                               & \sum_{j_{3}=0}^{q_3} \beta_{3j_{3}} \Delta \text{EXCHG}_{t-j_{3}} + \sum_{j_{4}=0}^{q_4} \beta_{4j_{4}} \Delta \text{M3}_{t-j_{4}} + \varepsilon_t 
 \end{align*}
 
-Here, the intercept term was given by $\alpha_0$. The 
-remaining $\alpha_i$ terms represent the short-run relationships 
-between the variables, and $\phi_j$ represent the long-run relationships.
+The intercept term was given by $\alpha_0$. The 
+remaining $\alpha_i, \beta_i$ coefficients represent the long-run relationships 
+between the variables. The coefficients were computed using the ardl function in the 
+ARDL package in R.
 
-In the ARDL bounds test, the null and alternate hypothesis were:
+The short-term dynamics were modelled using an Error Correction Model (ECM)
 
 \begin{align*}
-    H_{0}: \phi_i &= \phi_j\\
-    H_{1}: \phi_i &\neq \phi_j
+    \Delta \text{FTSE}_{250_t} &= \gamma_0 + \sum_{i=1}^{p} \gamma_i \Delta \text{FTSE}_{250_{t-i}} + \sum_{j_{1}=0}^{q_1} \delta_{1j_{1}} \Delta \text{CPI}_{t-j_{1}} + \sum_{j_{2}=0}^{q_2} \delta_{2j_{2}} \Delta \text{INT}_{t-j_{2}} \\
+                               & + \sum_{j_{3}=0}^{q_3} \delta_{3j_{3}} \Delta \text{EXCHG}_{t-j_{3}} + \sum_{j_{4}=0}^{q_4} \delta_{4j_{4}} \Delta \text{M3}_{t-j_{4}} + \lambda \text{ECM}_{t-1} + \varepsilon_t
 \end{align*}
 
-for, $i,j = 1,2,3,4,5$, with $i\neq j$. The F-statistic was computed for the joint
-significance of the lagged variables and compared against the critical values
-given by \citeboth{pesaran2001} to determine whether cointegration exists, 
-and at what significance level. If cointegration exists, then 
-the long-run coefficients $\kappa_i$ are estimated to find the cointegration equation:
+Here, the intercept term was given by $\gamma_0$. The 
+remaining $\gamma_i, \delta_i$ coefficients represent the short-run 
+relationships between the variables. The $\lambda$ coefficient 
+describes the speed at which short-run deviations from the long-run equilibrium
+are adjusted. A negative $\gamma$ implies that there exists a correction mechanism that guides the deviation 
+to the long-run equilibrium. These coefficients, as well 
+as the error correction term ECT, were computed using the recm function in the 
+ARDL package in R.
 
-\begin{align}
-    FTSE_{250} = \kappa_0 + \kappa_1 CPI + \kappa_2 INT + \kappa_3 EXCHG + \kappa_4 M3 + \varepsilon \label{eq:two}
-\end{align}
-
-Then, an Error Correction Model (ECM) was constructed:
+Subsequently, cointegration was checked. The null and alternate hypothesis were:
 
 \begin{align*}
-    \Delta \text{FTSE}_{250_t} &= \alpha_0 + \sum_{i=1}^{p} \alpha_i \Delta \text{FTSE}_{250_{t-i}} + \sum_{j_{1}=0}^{q_1} \beta_{1j_{1}} \Delta \text{CPI}_{t-j_{1}} + \sum_{j_{2}=0}^{q_2} \beta_{2j_{2}} \Delta \text{INT}_{t-j_{2}} \\
-                               & \sum_{j_{3}=0}^{q_3} \beta_{3j_{3}} \Delta \text{EXCHG}_{t-j_{3}} + \sum_{j_{4}=0}^{q_4} \beta_{4j_{4}} \Delta \text{M3}_{t-j_{4}} + \gamma\text{ECT}_{t-1} + \varepsilon_t
+    H_{0}: \beta_i &= 0\\
+    H_{1}: \beta_i &\neq 0
 \end{align*}
 
-where the error correction term, ECT, was found by calculating the residual 
-produced by \eqref{eq:two}:
-
-\begin{align*}
-    ECT_{t-1} = FTSE_{250_{t-1}} - \biggl(\kappa_0 + \kappa_1 CPI + \kappa_2 INT + \kappa_3 EXCHG + \kappa_4 M3 + \varepsilon \biggr)
-\end{align*}
-
-The parameter $\gamma$ describes the speed at which short-run deviations from the long-run equilibrium
-are adjusted. A negative $\gamma$ implies that there exists a correction mechanism that converges the deviation 
-to the long-run equilibrium. To estimate the short-run coefficients, $\alpha_i$ and $\beta_k$, as 
-well as the short-run correction speed $\gamma$ Ordinary Least Squares (OLS) regression was used.
+for all $i = 1,2,3,4$. The bounds\_f\_test in 
+the ARDL library, based on the bounds test approach 
+outlined in \citeboth{pesaran2001}, was used to calculate the 
+F-statistic of the ARDL model and subsequently test for cointegration.
 
@@
 display autocorrelation, heteroskedascity, non-normality and that the cumulative 
-sum of residuals and squared residuals are both stable.
+sum of residuals and squared residuals are both stable. This is to 
+ensure our regression is not spurious.
 
-\subsection{Results}
+\subsection{RBF Neural Network}
 
-The null hypothesis of the ADF Unit Root test is that of non-stationarity, and the alternate hypothesis 
-is that the variables are stationary.
+Radial basis function neural networks are a type of artificial 
+neural network that use radial basis functions at each node in the 
+activation layer of the network. They are able to identify complex, 
+non-linear relationships between time-series variables and have been used
+extensively in this aspect (\ref{subsubsec: fore}). The advantages of radial basis function neural networks over other 
+neural networks in time-series forecasting lies in their 
+ability to operate effectively on small sample sizes.
 
-\begin{table}[h!]
-    \centering
-    \caption{Augmented Dickey-Fuller (ADF) Unit Root Test\tablefootnote{All of the variables are tested in logarithmic form. The asterisks *, **, ***, and **** denote statistical significance
-    at the 1$\%$, 2.5$\%$, 5$\%$ and 10$\%$ significance levels.}}
-    \begin{tabular}{lccc}
-        \toprule
-        \textbf{Variable} & \textbf{Level ADF} & \textbf{First Difference ADF} & \textbf{Stationarity} \\
-        \midrule
-        FTSE$_{250}$ & $-3.2646^{****}$ & $-6.6892^{*}$ & At first difference \\
-        CPI          & $-1.4594$ & $-4.705^{*}$ & At first difference \\
-        EXCHG        & $-2.2467$ & $-7.3855^{*}$ & At first difference \\
-        INT          & $-2.0946$ & $-4.1156^{*}$ & At first difference \\
-        M3           & $-0.70827$ & $-6.3806^{*}$ & At first difference \\
-        \bottomrule
-    \end{tabular}
-\end{table}
+\subsubsection{Model Architecture}
+The exact interpolation problem, which forms the motivation for radial basis function neural networks, 
+involves mapping a $d$-dimensional input space $\boldsymbol{\mathrm{X}}$ to
+a $c$-dimensional target space $\boldsymbol{\mathrm{y}}$. The dataset consists of $N$ input
+vectors $\boldsymbol{\mathrm{x_n}}$, along with corresponding target 
+values $y_n$. The objective is to find a function
+  $h(\boldsymbol{\mathrm{x}})$ such that:
 
-Subsequently, the optimum lag of the variables was calculated using the 
-Schwarz Information Criterion, to obtain the optimum model 
-of ARDL(5, 4, 7, 7, 4). The resulting F-statistic of $4.1$ was compared against the critical values
-given by \citeboth{pesaran2001}, providing evidence of cointegration 
-at the $10\%$, $5\%$ and $2.5\%$ significance levels, but not at the $1\%$ 
-level.
+\begin{align}
+  h(\boldsymbol{\mathrm{x_n}}) = y_n, \quad \text{for } n = 1, \ldots, N
+\end{align}
 
-\begin{table}[h!]
-    \centering
-    \caption{ARDL Bounds Test Critical Values}
-    \begin{tabular}{lcccc}
-        \toprule
-        \textbf{Order} & \textbf{10$\%$} & \textbf{5$\%$} & \textbf{2.5$\%$} & \textbf{1$\%$} \\
-        \midrule
-        I(0) & $2.20$ & $2.56$ & $2.88$ & $3.29$ \\
-        I(1) & $3.09$ & $3.49$ & $3.87$  & $4.37$ \\
-        \bottomrule
-    \end{tabular}
-\end{table}
+The radial basis function approach 
+introduces a set of $N$ basis functions, one for each 
+data point. The radial basis function, $\phi(\cdot)$, is some non-linear function of the Euclidean distance
+$||\boldsymbol{\mathrm{x}} - \boldsymbol{\mathrm{x_n}}||$. The output of the mapping is then expressed as a linear combination of these basis functions:
 
-The long-run and short-run coefficients were then estimated, and presented 
-alonsgide the diagnostic test results which confirm that the model does not
-display autocorrelation, heteroskedascity, non-normality and the cumulative 
-sum of residuals and squared residuals are both stable. The error
-correction term is also given, and it was negative, as expected.
+\begin{align}
+    h(\boldsymbol{\mathrm{x}}) = \sum_{n=1}^{N} w_n \phi(||\boldsymbol{\mathrm{x}} - \boldsymbol{\mathrm{x_n}}||)
+\end{align}
 
-\begin{table}[h!]
-    \centering
-    \caption{Coefficients Estimates}
-    \begin{tabular}{llll}
-        \toprule
-        \textbf{Variable} & \textbf{Long-Run Coefficient} & \textbf{Variable} & \textbf{Short-Run Coefficient} \\
-        \midrule
-        FTSE$_{250}$ & $-$ & FTSE$_{250}$ $(-5)$  & $-0.033089$ \\
-        CPI & $3.157179$ & CPI $(-4)$ & $-0.921235$ \\
-        INT & $-0.1341097^{***}$ & INT $(-7)$ & $-0.034572^{****}$\\
-        EXCHG &  $-0.890146^{****}$ & EXCHG $(-7)$ & $0.122318$ \\
-        M3 & $-0.1141114$ & M3 $(-4)$ & $-0.527798^{***}$ \\
-        Intercept & $-3.447767$ & ECT $(-1)$ & $-0.1794^{**}$ \\
-        \bottomrule
-    \end{tabular}
-\end{table}
+By introducing several modifications to the exact interpolation 
+procedure, the radial basis function neural network
+model is obtained, as presented in Figure \ref{fig:rbf_nn_architecture}. 
+The network mapping takes the following form:
 
-\begin{table}[h!]
-    \centering
-    \caption{Diagnostic Tests}
-    \begin{tabular}{llll}
-        \toprule
-        \textbf{Condition} & \textbf{Test} & \textbf{p-value} & \textbf{Result} \\
-        \midrule
-        Autocorrelation & Box-Ljung Test & $0.9926$ & No autocorrelation \\
-        Heteroskedascity & Breusch-Pagan Test & $0.3635$ & Homoskedascity \\
-        Normality & Shapiro-Wilk Test & $3.889$e-$08$ & Normality \\
-        CUSUM & CUSUM Test & $-$ & Stable \\
-        CUSUM$^2$ & CUSUM$^2$ Test & $-$ & Stable \\
-        \bottomrule
-    \end{tabular}
-\end{table}
+\begin{align}
+    y(\boldsymbol{\mathrm{x}}) =  \sum_{j=1}^{M} w_{j} \phi_j(\boldsymbol{\mathrm{x}}) + w_{0}
+\end{align}
 
-The long-term coefficients show that the interest rate and exchange rate
-interest rate have significant long-term effects on the stock market price
-while money supply and inflation are statistically insignificant. \citeboth{neifar2023}
-proposes that this is because deflation, that is the negative movements in the 
-CPI, is a statistically significant factor in the movement of the UK stock 
-market, but positive movements (inflation) are not. Nevertheless, the positive
-coefficient of 3.16, and the negative short-term adjustment, is in line with the expectation from the literature that 
-inflation exerts a positive effect on the stock market in the long-term, but a negative 
-one in the short-run.
+\begin{align}
+    \phi_j(\boldsymbol{\mathrm{x}})  &= f(||\boldsymbol{\mathrm{x}-c_j}||), \quad{j=1\ldots,M}
+\end{align}
 
-The insignificant long-term effect of the money supply can be explained by several reasons.
-Firstly, as discussed in \citeboth{sellin2001}, it is possible that the FTSE250 is efficient in processing changes in the 
-money supply, at least over certain intervals if not always. This would mean anticipated changes in the money supply would not affect the stock market price, only the
-unanticipated component of a change in money supply would, as established by \citeboth{sorensen1982} and \citeboth{bernanke2005}. 
-It is also possible that the M3 measure is too broad of a measure and includes overly illiquid components that are not as impactful 
-on the stock market dynamics. In the short-run it is observed that the money supply does have a 
-significantly negative effect on the stock market, as outlined by \citeboth{olawale2014}. According to \citeboth{sellin2001}, 
-this supports the Keynesian argument that a positive money supply shock will lead people to
-expect a contraction in future monetary policy. 
-This leads to an increase in the demand for funds, which drives up the current rate of interest, in turn
-reducing the present value of earnings in the Discounted Cash Flow (DCF) stock valuation model. The effect is a decline in the stock market prices.
-
-Unsurprisingly, the USD/GBP exchange rate was significantly related to the FTSE250's movements. It 
-should not be forgotten that, although it is considered to be a more domestic indicator of the LSE, 
-at times $50\%$ of the FTSE250's earnings have come from abroad, indicating that 
-the value of the Pound has a significant impact on this stock market. 
-In the long run, the appreciation of the Pound would increase the relative 
-cost of UK exports increases, thereby reducing the 
-competitiveness of FTSE250 companies in international markets. This in turn would lower their earnings
-and depress the FTSE250. Moreover, foreign investors may view the stronger Pound as a signal of reduced returns on their investments in the UK, which can lead to decreased investment in the FTSE250. 
-In the short-term, however, before structural shifts can materialise the
-weakening of the Pound has a positive effect. As the USD has been 
-the global reserve currency for many decades, the depreciation of the Pound 
-against it will lead to increased revenue in Pound terms for FTSE250 companies with substantial overseas operations.
-This is because their foreign earnings, when converted back into Pounds, will be worth more, potentially increasing profits.
-The findings
-are in line with \citeboth{neifar2023} and \citeboth{khan2018}.
-
-The interest rate was also found to have significant negative long-term and short-term effects on the stock market, confirming the findings of \citeboth{alam2009}, \citeboth{demir2019} and \citeboth{neifar2023}
-This can be explained by the fact that higher interest 
-rates increase the cost of borrowing for companies, leading to reduced investment 
-and lower future earnings potential, in turn depressing stock prices. 
-Furthermore, as per the flight-to-quality phenomenon, persistently higher interest rates make bonds and other fixed-income 
-investments more attractive relative to stocks, leading investors to shift 
-their portfolios away from equities and toward safer assets over time (\citeboth{asgharian2016}). 
-In the short term, interest rate hikes immediately decrease the present value of earnings, as per the 
-DCF valuation model, and depress prices.
-
-The error correction term 
-was negative and statistically significant ensuring that long-run equilibrium 
-can be achieved. The adjustment speed of $0.1794$ indicates that 
-roughly $18\%$ of the fluctuation in the previous month's price caused by the 
-macroeconomic variable movements is adjusted back to the long-run equilibrium in the subsequent month. This implies 
-that it takes approximately 6 months for the FTSE250 to return to its long-run equilibrium following 
-changes in the selected macroeconomic variables.
-
-\section{RBF Neural Network}
-
-\subsection{Model Architecture}
+where $\boldsymbol{c_j}$ is a radial basis function centre and $f$ is the radial basis function (\citeboth{bishop1995}).
 
@@
 
-The exact interpolation problem, which forms the motivation for radial basis function neural networks, 
-involves mapping a $d$-dimensional input space $\boldsymbol{\mathrm{X}}$ to
-a $c$-dimensional target space $\boldsymbol{\mathrm{y}}$. The dataset consists of $N$ input
-vectors $\boldsymbol{\mathrm{x_n}}$, along with corresponding target 
-values $y_n$. The objective is to find a function
-  $h(\boldsymbol{\mathrm{x}})$ such that:
+\subsubsection{Methodology}
+
+First, the dataset was normalised using a z-score normalisation, because neural networks work best with data
+that is between a specified range (\citeboth{mendelsohn1993}).
+
+The inputs into the network are 4-dimensional input vectors, \(\boldsymbol{\mathrm{x_i}}\),
 
 \begin{align}
-  h(\boldsymbol{\mathrm{x_n}}) = y_n, \quad \text{for } n = 1, \ldots, N
+    \boldsymbol{\mathrm{x_i}} = \{\Delta \text{CPI}_i, \Delta \text{EXCHG}_i, \Delta \text{INT}_i, \Delta \text{M3}_i\}
 \end{align}
 
-The radial basis function approach 
-introduces a set of $N$ basis functions, one for each 
-data point. The radial basis function, $\phi(\cdot)$, is some non-linear function of the Euclidean distance
-$||\boldsymbol{\mathrm{x}} - \boldsymbol{\mathrm{x_n}}||$. The output of the mapping is then expressed as a linear combination of these basis functions:
+where \(\Delta \text{CPI}_i\), \(\Delta \text{EXCHG}_i\), \(\Delta \text{INT}_i\), and \(\Delta \text{M3}_i\) 
+represent differenced monthly values in the CPI, USD/GBP exchange rate, interest rate, and M3 money supply, respectively.
+
+The output is a one-dimensional output vector, \(\boldsymbol{\mathrm{y_i}}\), representing differenced monthly values in the FTSE250.
 
 \begin{align}
-    h(\boldsymbol{\mathrm{x}}) = \sum_{n=1}^{N} w_n \phi(||\boldsymbol{\mathrm{x}} - \boldsymbol{\mathrm{x_n}}||)
+    \boldsymbol{\mathrm{y_i}} = \{\Delta \text{FTSE}_{250_i}\}
 \end{align}
 
-By introducing several modifications to the exact interpolation 
-procedure, the radial basis function neural network
-model is obtained, as presented in Figure \ref{fig:rbf_nn_architecture}. 
-The network mapping takes the following form:
+Following this 
+it was necessary to program the radial basis function neural network into Python. The Bayesian kernel regression approach outlined by 
+\citeboth{bishop1995}, using the GaussianProcessRegressor package in Python was employed to model the network. For a 
+detailed description of this construction please refer to the \hyperref[sec: appendix]{Appendix}.
 
-\begin{align}
-    y(\boldsymbol{\mathrm{x}}) =  \sum_{j=1}^{M} w_{j} \phi_j(\boldsymbol{\mathrm{x}}) + w_{0}
-\end{align}
-
-\begin{align}
-    \phi_j(\boldsymbol{\mathrm{x}})  &= f(||\boldsymbol{\mathrm{x}-c_j}||), \quad{j=1\ldots,M}
-\end{align}
-
-where $\boldsymbol{c_j}$ is a radial basis function centre and $f$ is the radial basis function (\citeboth{bishop1995}).
-
-\subsection{Methodology}
-
-The radial basis function neural network was utilised for both short-term and long-term 
-forecasting. For the short-term, 
-the model was tested over a 5-month horizon, based on the optimum lag calculated earlier. 
-The long-term forecst spanned 57 months. 
-The data was differenced based on the recommendations of the 
-unit root tests conducted earlier to achieve stationarity.
-
-The data was normalised using a z-score normalisation, because neural networks work best with data
-that is between a specified range (\citeboth{mendelsohn1993}). Following this 
-it was necessary to program the radial basis function neural network into Python. The kernel regression approach outlined by 
-\citeboth{bishop1995}, using the GaussianProcessRegressor package in Python was employed to model the network.
-
-\subsubsection{The Gaussian Process Regressor}
-
-The GaussianProcessRegressor function in Python's Scikit library is based on the 
-standard function
-
-\begin{align} 
-    f(x) = x^{T}w, \quad{y = f(x) + \varepsilon}
-\end{align}
-
-where $y$ differs from the function values $f(x)$ by noise which is assumed
-to follow an iid Gaussian distribution (\citeboth{rasmussen2006}). 
-$\mathcal{N}_1$
-\begin{align*}
-    \varepsilon \sim \mathcal{N}_1 \bigl(0, \sigma_{n}^2\bigr)
-\end{align*}
-This assumption together with the model produces the 
-probability density of the observations given the inputs, 
-which is factored over cases in the training set. This is
-permissible because of the independence assumption (\citeboth{rasmussen2006}).
-
-\begin{align}
-    p(y|X,w) = \prod_{i=1}^{n} p(y_i|x_i ,w) = \mathcal{N}\bigl(X^T, \sigma_{n}^2 I\bigr)
-\end{align}
-
-The weights \( w \) represent the parameters that define the relationship between the inputs and outputs. 
-They are treated as random variables defined by the Gaussian distribution \eqref{eq: weight}.
-
-A covariance function or kernel
-is specified according to the characteristics of the data. In this study the
-generalised inverse multiquadric is used.
-
-\begin{align}
-    k(x,x') = \biggl( ||x-x'|| + \sigma^2\biggr)^{-\beta} \label{eq: gimq}
-\end{align}
-and used to construct the covariance matrix $K$, which along with 
-a mean function $m(x)$ forms the distribution of the the weights
-
-\begin{align}
-    w \sim \mathcal{N}\bigl(m(x), K\bigr) \label{eq: weight}
-\end{align}
-
-The mean function $m(x)$ acts as a bias, 
-providing a baseline expectation of the function values. 
-This forms the prior distribution over the weights \eqref{eq: weight}.
-
-Then, by Bayes' Theorem,
-
-\begin{align}
-    p(w|y, X) = \frac{p(y|X,w) p(w)}{p(y|X)} \label{eq: post}
-\end{align}
-Using this, the posterior distribution \eqref{eq:nine} is obtained (\citeboth{rasmussen2006}).
-
-\begin{align}
-    p(w|y, X) \sim \mathcal{N}\bigl(\frac{A^{-1}Xy}{\sigma_{n}^{2}}, A^{-1}\bigr) \label{eq:nine}
-\end{align}
-
-The covariance matrix $A^{-1} = K^{-1} + \sigma_{n}^{-2}(XX^T$).
-All possible $w$, weighted by their probability, are then averaged to obtain the
-most likely functional estimate of the output (\citeboth{rasmussen2006}).
-
-\subsubsection{The Network}
-
-First, the optimal hyperparameters in the covariance function \eqref{eq: gimq} 
+Based on this, the train-test process is then conducted:
+\begin{enumerate}
+    \item First, the optimal hyperparameters in the covariance function \eqref{eq: gimq} 
     are computed using GridSearch Cross Validation; these hyperparameters are:
@@
     \end{enumerate}
-
-Then, training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
+    \item Then, training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
     and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) are successively passed through the 
@@
     all training observations, $K$.
-
-A $4$-dimensional input space $X$ takes in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
+    \item A $4$-dimensional input space $X$ takes in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
     calculates the posterior weights distribution. 
@@
     the predicted output generated.
-
-The testing process is conducted similarly. The 
+    \item The testing process is conducted similarly. The 
     posterior weight distribution that contains the information of the training phase becomes the new 
@@
     \end{align}
+\end{enumerate}
 
-\subsection{Results}
+A short-term and long-term forecast is conducted using this process. For the short-term forecast a test set of 5 months was used, and for the long-term forecast a test set of 57 months was used.
 
+\section{Results}
+
+\subsection{ARDL-ECM}
+
+\begin{table}[h!]
+    \centering
+    \caption{Augmented Dickey-Fuller (ADF) Unit Root Test\tablefootnote{All of the variables are tested in natural logarithmic form. The asterisks *, **, ***, and **** denote statistical significance
+    at the 1$\%$, 2.5$\%$, 5$\%$ and 10$\%$ significance levels.}}
+    \begin{tabular}{lccc}
+        \toprule
+        \textbf{Variable} & \textbf{Level ADF} & \textbf{First Difference ADF} & \textbf{Stationarity} \\
+        \midrule
+        FTSE$_{250}$ & $-3.2646^{****}$ & $-6.6892^{*}$ & At first difference \\
+        CPI          & $-1.4594$ & $-4.705^{*}$ & At first difference \\
+        EXCHG        & $-2.2467$ & $-7.3855^{*}$ & At first difference \\
+        INT          & $-2.0946$ & $-4.1156^{*}$ & At first difference \\
+        M3           & $-0.70827$ & $-6.3806^{*}$ & At first difference \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+The optimum lags for the ARDL model were ($FTSE_{250} = 5$, $CPI = 4$, $EXCHG=7$, $INT =7$, $M3=4$) for  The resulting F-statistic of $4.1$ was compared against the critical values
+given by \citeboth{pesaran2001}, providing evidence of cointegration 
+at the $10\%$, $5\%$ and $2.5\%$ significance levels, but not at the $1\%$ 
+level.
+
+\begin{table}[h!]
+    \centering
+    \caption{ARDL Bounds Test Critical Values}
+    \begin{tabular}{lcccc}
+        \toprule
+        \textbf{Order} & \textbf{10$\%$} & \textbf{5$\%$} & \textbf{2.5$\%$} & \textbf{1$\%$} \\
+        \midrule
+        I(0) & $2.20$ & $2.56$ & $2.88$ & $3.29$ \\
+        I(1) & $3.09$ & $3.49$ & $3.87$  & $4.37$ \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+\begin{table}[h!]
+    \centering
+    \caption{Coefficients Estimates}
+    \begin{tabular}{llll}
+        \toprule
+        \textbf{Variable} & \textbf{Long-Run Coefficient} & \textbf{Variable} & \textbf{Short-Run Coefficient} \\
+        \midrule
+        FTSE$_{250}$ & $-$ & FTSE$_{250}$ $(-5)$  & $-0.033089$ \\
+        CPI & $3.157179$ & CPI $(-4)$ & $-0.921235$ \\
+        INT & $-0.1341097^{***}$ & INT $(-7)$ & $-0.034572^{****}$\\
+        EXCHG &  $-0.890146^{****}$ & EXCHG $(-7)$ & $0.122318$ \\
+        M3 & $-0.1141114$ & M3 $(-4)$ & $-0.527798^{***}$ \\
+        Intercept & $-3.447767$ & ECT $(-1)$ & $-0.1794^{**}$ \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+\begin{table}[h!]
+    \centering
+    \caption{Diagnostic Tests}
+    \begin{tabular}{llll}
+        \toprule
+        \textbf{Condition} & \textbf{Test} & \textbf{p-value} & \textbf{Result} \\
+        \midrule
+        Autocorrelation & Box-Ljung Test & $0.9926$ & No autocorrelation \\
+        Heteroskedascity & Breusch-Pagan Test & $0.3635$ & Homoskedascity \\
+        Normality & Shapiro-Wilk Test & $3.889$e-$08$ & Normality \\
+        CUSUM & CUSUM Test & $-$ & Stable \\
+        CUSUM$^2$ & CUSUM$^2$ Test & $-$ & Stable \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+\subsubsection{Short-term Impacts}
+
+The interest rate has a significantly
+negative short-term impact on the FTSE250 price in the short term. 
+This can be attributed to the immediate effect of interest rate hikes, 
+which decrease the present value of future earnings according to the 
+Discounted Cash Flow (DCF) valuation model, thereby 
+affecting investor sentiment and depressing stock prices. 
+This finding is consistent with the studies by 
+\citeboth{alam2009}, \citeboth{demir2019}, and \citeboth{neifar2023}.
+
+The money supply has a significantly negative short-term effect on the 
+FTSE250 price, as noted by \citeboth{olawale2014}. 
+According to \citeboth{sellin2001}, this supports the Keynesian argument 
+that an increase in money supply leads to expectations of a future 
+monetary contraction. This expectation increases the demand for 
+funds, driving up current interest rates and reducing stock prices 
+through the same mechanism as described previously.
+
+In the short-term, before structural shifts can materialise, the
+weakening of the Pound against the Dollar has a positive, but 
+insignificant effect. As the USD has been 
+the global reserve currency for many decades, the depreciation of the Pound 
+against it will lead to increased revenue in Pound terms for FTSE250 
+companies with substantial overseas operations. 
+This is because their foreign earnings, 
+when converted back into Pounds, will be worth more, 
+potentially increasing profits. However, the insignificant impact suggests 
+that FTSE250 firms are resilient to short-term fluctuations in the stock market 
+price. The possible reasons for this are that they have diversified 
+operations, meaning fluctuations against the Dollar will not threaten 
+their profits and they use hedging strategies involving financial instruments 
+such as currency forwards which protect against volatility.
+
+The insignificance of inflation, defined as a positive 
+increase in the CPI, confirms the findings of 
+\citeboth{gultekin1983}, \citeboth{firth1979} and \citeboth{neifar2023}. 
+The CPI as a measure does not uniformly represent all sectors of the economy, 
+and may consequently be unrepresentative of inflationary changes in the 
+sectors that predominate the FTSE250. In addition, different sectors 
+are more vulnerable to inflation than others which may distort the effect 
+of inflation even further. Moreover, inflation may be anticipated 
+by investors and already be factored into stock prices.
+
+The error correction term 
+was negative and statistically significant ensuring that long-run equilibrium 
+can be achieved. The adjustment speed of $0.1794$ indicates that 
+roughly $18\%$ of the fluctuation in the previous month's price caused by the 
+macroeconomic variable movements is adjusted back to the long-run equilibrium in the subsequent month. This implies 
+that it takes approximately 6 months for the FTSE250 to return to its long-run equilibrium following 
+changes in the selected macroeconomic variables.
+
+\subsubsection{Long-term Impacts}
+
+The CPI did not have a significant effect on the FTSE250 price over the 
+long-run, lending support to the findings of 
+\citeboth{gultekin1983} and \citeboth{firth1979}
+that the Fisher hypothesis is valid for the UK. One of the 
+reasons for this may be that, since stock prices represent the 
+discounted value of future cash flows (which are real), 
+they are not directly impacted by inflation in the long run. Moreover, 
+firms are able to pass on nominal costs to consumers which reduces
+the impact of inflation.
+The freer role of the Bank of England in setting monetary policy could be 
+another factor in this, as a stricter monetary policy would keep inflation
+in check and obscure any long-term equilibrium relationship.
+
+The insignificant long-term effect of the money supply can be explained by several reasons.
+Firstly, as discussed in \citeboth{sellin2001}, it is possible that the FTSE250 is efficient in processing changes in the 
+money supply, at least over certain intervals if not always. This would mean anticipated changes in the money supply would not affect the stock market price, only the
+unanticipated component of a change in money supply would, as established by \citeboth{sorensen1982} and \citeboth{bernanke2005}. 
+It is also possible that the M3 measure is too broad of a measure and includes overly illiquid components that are not as impactful 
+on the stock market dynamics.
+
+The USD/GBP exchange rate significantly positively impacted the FTSE250 price, confirming the findings of 
+\citeboth{wong2022}. That there is a 
+significant relationship is not surprising, because whilst the FTSE250 is a more domestic indicator of the London Stock Exchange than the FTSE 100 Index, 
+at times nearly $50\%$ of its earnings have come from abroad, indicating that 
+the value of the Pound has a significant impact on this stock market. There are many reasons for why this effect is observed.
+In the long run, the appreciation of the Pound would decrease the 
+cost of imports, thereby increasing the 
+profits of FTSE250 companies and inflating their share prices.
+Moreover, a stronger Pound makes UK assets more appealing to foreign investors, as their investments would be worth more when converted back to their domestic currencies.
+This would drive up demand for UK stocks and positively influence the 
+FTSE250 price.
+
+The interest rate was also found to have significant negative long-term and short-term effects on the stock market, confirming the findings of \citeboth{alam2009}, \citeboth{demir2019} and \citeboth{neifar2023}
+This can be explained by the fact that higher interest 
+rates increase the cost of borrowing for companies, leading to reduced investment 
+and lower future earnings potential, in turn depressing stock prices. 
+Furthermore, as per the flight-to-quality phenomenon, persistently higher interest rates make bonds and other fixed-income 
+investments more attractive relative to stocks, leading investors to shift 
+their portfolios away from equities and toward safer assets over time (\citeboth{asgharian2016}). 
+In the short term, interest rate hikes immediately decrease the present value of earnings, as per the 
+DCF valuation model, and depress prices.
+
+\subsection{RBFNN}
+
+\subsubsection{Long-term Forecast}
 The model returned a mean absolute error of MAE $=0.16$ and 
@@
 $2.4\%$ increase in the $R^2$ obtained by the ARDL-ECM model. The results are presented 
-in Figures \ref{fig:longmonthly} and \ref{fig:longdaily}, against monthly and daily FTSE250 changes respectively.
+in Figure \ref{fig:longmonthly}.
 
@@
     \centering
+    \hspace*{-1.5cm}
     \input{long_term_monthly.tex}  
@@
     \label{fig:longmonthly}
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \input{long_term_daily.tex}  
-    \caption{Long-term predicted vs real daily FTSE250 changes}
-    \label{fig:longdaily}
 \end{figure}
@@
 
+\subsubsection{Short-term Forecast}
+
 The short-term prediction yielded MAE = $0.07$ and 
-$R^2 = 61\%$ for a 5-month time horizon. The results are presented in Figures \ref{fig:shortmonthly} 
-and \ref{fig:shortdaily}, against monthly and daily FTSE250 changes respectively.
+$R^2 = 61\%$ for a 5-month time horizon. The results are presented in Figure \ref{fig:shortmonthly}.
 The significantly better performance of the model in the short-run
@@
     \centering
+    \hspace*{-1.5cm}
     \input{short_term_monthly.tex}  
@@
     \label{fig:shortmonthly}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \input{short_term_daily.tex}  
-    \caption{Short-term predicted vs real daily FTSE250 changes}
-    \label{fig:shortdaily}
 \end{figure}
@@
 that dilute the relationship between these variables and the stock market in the 
-long-run. As the prediction window expands, 
+long-run. As the prediction window expands in regression problems, 
 more uncertainty and unknowns are introduced, which translates into 
@@
 
+\section{Appendix}
+\label{sec: appendix}
+
+The GaussianProcessRegressor function in Python's Scikit library is based on the 
+standard function
+
+\begin{align} 
+    f(x) = x^{T}w, \quad{y = f(x) + \varepsilon}
+\end{align}
+
+where $y$ differs from the function values $f(x)$ by noise which is assumed
+to follow an iid Gaussian distribution 
+$\mathcal{N}_1$ (\citeboth{rasmussen2006}).
+\begin{align*}
+    \varepsilon \sim \mathcal{N}_1 \bigl(0, \sigma_{n}^2\bigr)
+\end{align*}
+This assumption together with the model produces the 
+probability density of the observations given the inputs, 
+which is factored over cases in the training set. This is
+permissible because of the independence assumption (\citeboth{rasmussen2006}).
+
+\begin{align}
+    p(y|X,w) = \prod_{i=1}^{n} p(y_i|x_i ,w) = \mathcal{N}\bigl(X^T, \sigma_{n}^2 I\bigr)
+\end{align}
+
+The weights \( w \) represent the parameters that define the relationship between the inputs and outputs. 
+They are treated as random variables defined by the Gaussian distribution \eqref{eq: weight}.
+
+A covariance function or kernel
+is specified according to the characteristics of the data. In this study the
+generalised inverse multiquadric is used.
+
+\begin{align}
+    k(x,x') = \biggl( ||x-x'|| + \sigma^2\biggr)^{-\beta} \label{eq: gimq}
+\end{align}
+and used to construct the covariance matrix $K$, which along with 
+a mean function $m(x)$ forms the distribution of the the weights
+
+\begin{align}
+    w \sim \mathcal{N}\bigl(m(x), K\bigr) \label{eq: weight}
+\end{align}
+
+The mean function $m(x)$ acts as a bias, 
+providing a baseline expectation of the function values. 
+This forms the prior distribution over the weights \eqref{eq: weight}.
+
+Then, by Bayes' Theorem,
+
+\begin{align}
+    p(w|y, X) = \frac{p(y|X,w) p(w)}{p(y|X)} \label{eq: post}
+\end{align}
+Using this, the posterior distribution \eqref{eq:nine} is obtained (\citeboth{rasmussen2006}).
+
+\begin{align}
+    p(w|y, X) \sim \mathcal{N}\bigl(\frac{A^{-1}Xy}{\sigma_{n}^{2}}, A^{-1}\bigr) \label{eq:nine}
+\end{align}
+
+The covariance matrix $A^{-1} = K^{-1} + \sigma_{n}^{-2}(XX^T$).
+All possible $w$, weighted by their probability, are then averaged to obtain the
+most likely functional estimate of the output (\citeboth{rasmussen2006}).
+
 \bibliographystyle{plainnat}

### 2024-08-25 tex/paper.tex
final adjustments

@@
 
-Macroeconomic indicators can be effectively used in machine learning 
+Macroeconomic indicators can also be effectively used in machine learning 
 methods to forecast stock prices by serving as input features that capture 
-the economic environment’s influence on market behaviour. By carefully 
+the economic environment’s influence on stock market behaviour. By carefully 
 selecting the most relevant indicators and preprocessing data, 
@@
 Section \ref{sec:lit} provides a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables.
-Section \ref{sec:meth} presents the methodology of the paper. Section 4 discusses the results and Section 5 concludes
+Section \ref{sec:meth} presents the methodology of the paper. Section \ref{sec: results} discusses the results and Section \ref{sec: conc} concludes
 the study.
@@
 
-\subsection{Data}
-Monthly FTSE 250 Index price data between Oct 1992 to March 2024 has been utilised for this study giving us a dataset with 379 observations.
+\subsection{Data Sources}
+Monthly FTSE 250 Index price data, 
+monthly CPI data, monthly M3 money supply data, monthly USD/GBP exchange 
+rate data and interest rate data between Oct 1992 to March 2024 has been 
+utilised for this study giving a dataset with 379 observations. The sources 
+are presented in Table \ref{table: tabl}.
 
+\clearpage
 \begin{table}[h!]
@@
     \caption{Data sources}
+    \label{table: tabl}
     \begin{tabular}{lll}
@@
 
+\subsection{Data Preprocessing}
 For the interest rate data, it was necessary to address an issue raised by the fact that the Bank of England’s Monetary Policy Committee (MPC) 
 does not convene on a monthly basis, and are not obligated to 
-announce a rate on a regular basis. To address the resulting irregularities the data was adjusted using \ref{alg:interest_rate_adjustment}.
+announce a rate on a regular basis. To address the resulting irregularities the data was adjusted using Algorithm \ref{alg:interest_rate_adjustment}.
 
@@
     \KwData{Interest rate data adjustments by the Bank of England’s MPC}
-    \KwResult{Monthly interest rate data}
+    \KwResult{Monthly interest rate data (INT)}
 
@@
         \If{$\forall d \in D_m \; \text{(interest rate unchanged)}$}{
-            \textbf{Set:} $\text{Rate}_m \leftarrow \text{Rate}_{m-1}$\;
+            \textbf{Set:} $\text{INT}_m \leftarrow \text{INT}_{m-1}$\;
         }
@@
             }
-            \textbf{Set:} $\text{Rate}_m \leftarrow W_{int}$\;
+            \textbf{Set:} $\text{INT}_m \leftarrow W_{int}$\;
         }
@@
 \end{algorithm}
+
+The result is a dataset $X$ composed of monthly values for the FTSE250
+price, CPI value, USD/GBP exchange rate, interest rate, and M3 money supply.
+
+\subsection{Descriptive Statistics}
 
@@
 non-linear relationships between time-series variables and have been used
-extensively in this aspect (\ref{subsubsec: fore}). The advantages of radial basis function neural networks over other 
+extensively in this aspect. The advantages of radial basis function neural networks over other 
 neural networks in time-series forecasting lies in their 
@@
 
-\subsubsection{Model Architecture}
+\subsubsection{Network Architecture}
 The exact interpolation problem, which forms the motivation for radial basis function neural networks, 
@@
 
-\subsubsection{Methodology}
+\subsubsection{Network Construction}
 
-First, the dataset was normalised using a z-score normalisation, because neural networks work best with data
-that is between a specified range (\citeboth{mendelsohn1993}).
+Neural networks work best with data that is specified between a specific 
+range (\citeboth{mendelsohn1993}). Also, to achieve stationarity
+it was required to take the first difference all of the variables. Therefore, all of the 
+variables in the dataset described in were normalised using a z-score normalisation
+and differenced, as described in Algorithm \ref{alg:z_score_differencing}.
 
-The inputs into the network are 4-dimensional input vectors, \(\boldsymbol{\mathrm{x_i}}\),
+\begin{algorithm}[H]
+    \caption{Z-Score Normalisation and Differencing of Variables}
+    \label{alg:z_score_differencing}
+    \KwData{Dataset $X$ with variables $X_1 = FTSE_{250}, X_2 = CPI, X_3 = EXCHG, X_4 = INT, X_5 = M3$}
+    \KwResult{Normalised and differenced dataset $Z$}
+
+    \For{each variable $X_i$ in $X$}{
+        \textbf{Compute:} $\mu_i \leftarrow \frac{1}{m} \sum_{j=1}^{m} X_{ij}$\;
+        \textbf{Compute:} $\sigma_i \leftarrow \sqrt{\frac{1}{m} \sum_{j=1}^{m} (X_{ij} - \mu_i)^2}$\;
+        \For{each observation $X_{ij}$ in $X_i$}{
+            \textbf{Normalise:} $Z_{ij} \leftarrow \frac{X_{ij} - \mu_i}{\sigma_i}$\;
+        }
+        \For{each observation $Z_{ij}$ from the second observation to the last}{
+            \textbf{Difference:} $D_{ij} \leftarrow Z_{ij} - Z_{i(j-1)}$\;
+        }
+    }
+    \textbf{Update:} $Z_i \leftarrow D_i$ \;
+\end{algorithm}
+
+The inputs into the network are 4-dimensional input vectors, \(\boldsymbol{\mathrm{x_i}}\), from the differenced, normalised dataset $Z$
 
@@
 
-The output is a one-dimensional output vector, \(\boldsymbol{\mathrm{y_i}}\), representing differenced monthly values in the FTSE250.
+The output is a one-dimensional output vector, \(\boldsymbol{\mathrm{y_i}}\) from the dataset $Z$, representing differenced monthly values in the FTSE250.
 
@@
 
-Following this 
-it was necessary to program the radial basis function neural network into Python. The Bayesian kernel regression approach outlined by 
+The radial basis function neural network was programmed into Python. The Bayesian kernel regression approach outlined by 
 \citeboth{bishop1995}, using the GaussianProcessRegressor package in Python was employed to model the network. For a 
@@
 
-Based on this, the train-test process is then conducted:
+Based on this, the train-test process was conducted:
 \begin{enumerate}
-    \item First, the optimal hyperparameters in the covariance function \eqref{eq: gimq} 
-    are computed using GridSearch Cross Validation; these hyperparameters are:
+    \item The optimal hyperparameters in the covariance function \eqref{eq: gimq} 
+    are computed using GridSearch Cross Validation. These hyperparameters are:
     \begin{enumerate}
@@
     \end{enumerate}
-    \item Then, training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
+    \item Training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
     and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) are successively passed through the 
@@
     the predicted output generated.
-    \item The testing process is conducted similarly. The 
+    \item The testing process was conducted similarly. The 
     posterior weight distribution that contains the information of the training phase becomes the new 
     prior distribution. From this the posterior weight distributions for the test observations \eqref{eq: testw}
-    are calculated, and then the predictive output is found (\citeboth{rasmussen2006}).
+    were calculated, and then the predictive output was found (\citeboth{rasmussen2006}).
 
@@
 \section{Results}
+\label{sec: results}
 
@@
 
-\subsubsection{Long-term Forecast}
-The model returned a mean absolute error of MAE $=0.16$ and 
-$R^2 = 9\%$ for the long-term prediction, providing a 
-$2.4\%$ increase in the $R^2$ obtained by the ARDL-ECM model. The results are presented 
-in Figure \ref{fig:longmonthly}.
-
-\begin{figure}[H]
-    \centering
-    \hspace*{-1.5cm}
-    \input{long_term_monthly.tex}  
-    \caption{Long-term predicted vs real monthly FTSE250 changes}
-    \label{fig:longmonthly}
-\end{figure}
-
-The interpretation of this is that, over the long-run (4-5 years), the variations in the 
-Consumer Price Index, interest rate, USD/GBP exchange rate, and M3 money 
-supply explain or account for $9\%$ of the fluctuations of the
-monthly FTSE250 value. This relates closely to the findings of \citeboth{conover1999}, 
-who used an econometric model to establish that $4\%$ of the variation in monthly UK stock returns can be 
-explained by variations in the money supply. It also suggests a higher level of influence 
-than the ARDL-ECM model provided, that only $R^2=3.4\%$ of the monthly FTSE250 variations 
-are affected by the selected macroeconomic variables, demonstrating the ability of machine learning
-methods to capture subtle patterns in the data unavailable to econometric models.
-
-In the context of a phenomenon as complex as the stock market,
-which is influenced by an inumerable quantity of factors such as investor sentiment
-- inherently irrational, 
-political instability, geopolitical shocks, corporate governance, and financial regulation, this result 
-unsurprising. Furthermore, the selection of economic variables in this study
-is by no means exhaustive; commodity prices, international trade dynamics,
-economic growth, and many other variables also exert a non-insignificant 
-effect on the stock market.
-
 \subsubsection{Short-term Forecast}
@@
 
+\subsubsection{Long-term Forecast}
+The model returned a mean absolute error of MAE $=0.16$ and 
+$R^2 = 9\%$ for the long-term prediction, providing a 
+$2.4\%$ increase in the $R^2$ obtained by the ARDL-ECM model. The results are presented 
+in Figure \ref{fig:longmonthly}.
+
+The interpretation of this is that, over the long-run (4-5 years), the variations in the 
+Consumer Price Index, interest rate, USD/GBP exchange rate, and M3 money 
+supply explain or account for $9\%$ of the fluctuations of the
+monthly FTSE250 value. This relates closely to the findings of \citeboth{conover1999}, 
+who used an econometric model to establish that $4\%$ of the variation in monthly UK stock returns can be 
+explained by variations in the money supply. It also suggests a higher level of influence 
+than the ARDL-ECM model provided, that only $R^2=3.4\%$ of the monthly FTSE250 variations 
+are affected by the selected macroeconomic variables, demonstrating the ability of machine learning
+methods to capture subtle patterns in the data unavailable to econometric models.
+
+\begin{figure}[H]
+    \centering
+    \hspace*{-1.5cm}
+    \input{long_term_monthly.tex}  
+    \caption{Long-term predicted vs real monthly FTSE250 changes}
+    \label{fig:longmonthly}
+\end{figure}
+
+In the context of a phenomenon as complex as the stock market,
+which is influenced by an inumerable quantity of factors such as investor sentiment
+- inherently irrational, 
+political instability, geopolitical shocks, corporate governance, and financial regulation, this result 
+unsurprising. Furthermore, the selection of economic variables in this study
+is by no means exhaustive; commodity prices, international trade dynamics,
+economic growth, and many other variables also exert a non-insignificant 
+effect on the stock market.
+
 \section{Conclusion}
+\label{sec: conc}
 

### 2024-08-26 tex/paper.tex
final revisions

@@
     This study investigates the impact of the Consumer Price Index (CPI), Bank of England Base Rate, US Dollar (USD) and British Pound Sterling (GBP) exchange rate, and the money supply (M3) measure on the price of the FTSE 250 Index (FTSE250).
-    In addition, it assesses the effectiveness of radial basis function neural networks as a tool to forecast the price of the FTSE 250 Index based on the same set of macroeconomic variables.
+    In addition, it assesses the effectiveness of radial basis function neural networks (RBFNN) as a tool to forecast the price of the FTSE 250 Index based on the same set of macroeconomic variables.
 This study finds quantitative estimates for the short-term and long-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 Index.
@@
 the economic environment’s influence on stock market behaviour. By carefully 
-selecting the most relevant indicators and preprocessing data, 
-as well as choosing appropriate models, machine learning can enhance the accuracy of stock price forecasts (\citeboth{Prasad2023}).
+selecting the most relevant indicators, 
+as well as choosing the appropriate models, machine learning can enhance the accuracy of stock price forecasts (\citeboth{Prasad2023}).
 
@@
 
-In order to analyse the dynamic relationship between variables in time series data, this study
+In order to analyse the dynamic relationship between the variables in the time series data, this study
 will employ the Autoregressive Distributed Lag (ARDL) model in conjunction with an Error
-Correction Model (ECM). This research also employs a radial basis function neural network
-(RBFNN) to predict the price movements of the FTSE250.
+Correction Model (ECM). This research also employs a radial basis function neural network to predict the price movements of the FTSE250.
 
 The remainder of this paper will be organised as follows. 
-Section \ref{sec:lit} provides a brief review of the literature in this problem space, and explains the rationale behind the selection of macroeconomic variables.
-Section \ref{sec:meth} presents the methodology of the paper. Section \ref{sec: results} discusses the results and Section \ref{sec: conc} concludes
+Section \ref{sec:lit} provides a brief review of the literature in this problem space and explains the rationale behind the selection of macroeconomic variables.
+Section \ref{sec:meth} presents the methodology. Section \ref{sec: results} discusses the results and Section \ref{sec: conc} concludes
 the study.
@@
 \subsection{Data Sources}
-Monthly FTSE 250 Index price data, 
-monthly CPI data, monthly M3 money supply data, monthly USD/GBP exchange 
-rate data and interest rate data between Oct 1992 to March 2024 has been 
-utilised for this study giving a dataset with 379 observations. The sources 
+
+The study uses monthly FTSE250 price data between October 1992 and March 2024,
+thus providing 379 observations. The data for the macroeconomic indicators 
+was also captured on a monthly basis in the same time frame, except for the 
+interest rate. This is because the Bank of England's Monetary Policy 
+Committee, 
+which is responsible for setting interest rates, meets on an \textit{ad hoc} 
+basis\footnote{The MPC typically meets roughly every six weeks, however, in uncertain economic 
+climes it may convene far more frequently.}
+to decide on whether the interest rate should be changed. The sources 
 are presented in Table \ref{table: tabl}.
@@
     \label{table: tabl}
-    \begin{tabular}{lll}
+    \begin{tabular}{llll}
         \toprule
-        \textbf{Variable} & \textbf{Notation} & \textbf{Source} \\
+        \textbf{Variable} & \textbf{Frequency} & \textbf{Notation} & \textbf{Source} \\
         \midrule
-        FTSE 250 Index & FTSE$_{250}$ & Yahoo Finance \\
-        Consumer Price Index & CPI & Office for National Statistics (ONS) \\
-        USD/GBP Exchange Rate (£) & EXCHG & Federal Reserve of St Louis \\
-        Interest Rate ($\%$) & INT & Bank of England \\
-        Money Supply (M3) (millions £) & M3 & Bank of England \\
+        FTSE 250 Index & Monthly & FTSE$_{250}$ & Yahoo Finance \\
+        Consumer Price Index & Monthly & CPI & Office for National Statistics (ONS) \\
+        USD/GBP Exchange Rate (£) & Monthly & EXCHG & Federal Reserve of St Louis \\
+        Interest Rate ($\%$) & Variable & INT & Bank of England \\
+        Money Supply (M3) (millions £) & Monthly & M3 & Bank of England \\
         \bottomrule
@@
 \subsection{Data Preprocessing}
-For the interest rate data, it was necessary to address an issue raised by the fact that the Bank of England’s Monetary Policy Committee (MPC) 
-does not convene on a monthly basis, and are not obligated to 
-announce a rate on a regular basis. To address the resulting irregularities the data was adjusted using Algorithm \ref{alg:interest_rate_adjustment}.
+\label{section: prep}
+
+To address the variability in the interest data, Algorithm \ref{alg:interest_rate_adjustment}
+was used to convert the data into a monthly format, ensuring a uniform
+frequency across the variables.
 
@@
 Interest rates, have varied widely from $0.1\%$ to $8.4\%$, with a standard deviation of $2.52\%$, reflecting the 
-increasing role played by monetary policy in the UK since May 1997 when the Bank of England was given operational responsibility for setting interest rates (\citeboth{king1997changes}).
+increasing role played by monetary policy in the UK since May 1997 when operational responsibility for setting interest rates was tranferred from the Chancellor of the Exchequer to the Bank of England (\citeboth{king1997changes}).
 The Bank of England was forced to adjust interest rates
@@
 of the variables was tested using the Augmented Dickey-Fuller 
-Unit Root Test found within the tseries package in R. The optimal lags for each of the 
+Unit Root Test found within the \texttt{tseries} package in R. The optimal lags for each of the 
 variables were calculated using the Schwarz Information Criterion, contained
@@
 The intercept term was given by $\alpha_0$. The 
-remaining $\alpha_i, \beta_i$ coefficients represent the long-run relationships 
-between the variables. The coefficients were computed using the ardl function in the 
-ARDL package in R.
+remaining $\alpha_i, \beta_i$ coefficients represented the long-run relationships 
+between the variables. The coefficients were computed using the \texttt{ardl} function in the 
+\texttt{ARDL} library in R.
 
@@
 
-Here, the intercept term was given by $\gamma_0$. The 
-remaining $\gamma_i, \delta_i$ coefficients represent the short-run 
+The intercept term was given by $\gamma_0$. The 
+remaining $\gamma_i, \delta_i$ coefficients represented the short-run 
 relationships between the variables. The $\lambda$ coefficient 
-describes the speed at which short-run deviations from the long-run equilibrium
-are adjusted. A negative $\gamma$ implies that there exists a correction mechanism that guides the deviation 
-to the long-run equilibrium. These coefficients, as well 
-as the error correction term ECT, were computed using the recm function in the 
-ARDL package in R.
+described the speed at which short-run deviations from the long-run equilibrium
+were adjusted. A negative $\gamma$ implies that there existed a correction mechanism that guided the deviation 
+back to the long-run equilibrium. These coefficients, as well 
+as the error correction term ECT, were computed using the \texttt{recm} function in the 
+\texttt{ARDL} library in R.
 
-Subsequently, cointegration was checked. The null and alternate hypothesis were:
+The \texttt{{bounds\_f\_test}} in 
+the \texttt{ARDL} library, based on the bounds test approach 
+outlined in \citeboth{pesaran2001}, was used to calculate the 
+F-statistic of the ARDL model and subsequently test for cointegration.
 
@@
 
-for all $i = 1,2,3,4$. The bounds\_f\_test in 
-the ARDL library, based on the bounds test approach 
-outlined in \citeboth{pesaran2001}, was used to calculate the 
-F-statistic of the ARDL model and subsequently test for cointegration.
+for all $i = 1,2,3,4$.
 
-Finally, diagnostic tests are conducted to confirm that the model does not
-display autocorrelation, heteroskedascity, non-normality and that the cumulative 
-sum of residuals and squared residuals are both stable. This is to 
-ensure our regression is not spurious.
+To confirm that the model did not display autocorrelation, heteroskedasticity, non-normality, or structural instability, the Box-Ljung, Breusch-Pagan, Shapiro-Wilk, and CUSUM tests were conducted, respectively
 
@@
 
-Radial basis function neural networks are a type of artificial 
-neural network that use radial basis functions at each node in the 
-activation layer of the network. They are able to identify complex, 
-non-linear relationships between time-series variables and have been used
-extensively in this aspect. The advantages of radial basis function neural networks over other 
-neural networks in time-series forecasting lies in their 
-ability to operate effectively on small sample sizes.
+This study employed radial basis function neural networks, 
+a type of artificial neural network that utilise radial basis functions 
+at each node in the activation layer. 
+This was because radial basis function neural networks 
+are well-suited for identifying complex, non-linear relationships 
+between time-series variables, making them particularly valuable in 
+forecasting tasks. In addition, radial basis function neural networks
+have been shown to be effective in operating on small sample sizes, 
+which was critical for the data constraints of this study.
 
@@
 Neural networks work best with data that is specified between a specific 
-range (\citeboth{mendelsohn1993}). Also, to achieve stationarity
-it was required to take the first difference all of the variables. Therefore, all of the 
-variables in the dataset described in were normalised using a z-score normalisation
-and differenced, as described in Algorithm \ref{alg:z_score_differencing}.
+range (\citeboth{mendelsohn1993}).
+Also, based on the unit root tests in Table \ref{table: unitroot} the variables in the dataset $X$, described in \ref{section: prep},
+only achieve stationary at first difference.
+Therefore, all of the 
+variables in the dataset $X$ were normalised using a z-score normalisation and differenced,
+as described in Algorithm \ref{alg:z_score_differencing}.
 
@@
 
-The inputs into the network are 4-dimensional input vectors, \(\boldsymbol{\mathrm{x_i}}\), from the differenced, normalised dataset $Z$
+The inputs into the network were 4-dimensional input vectors, \(\boldsymbol{\mathrm{x_i}}\), from the differenced, normalised dataset $Z$
 
@@
 where \(\Delta \text{CPI}_i\), \(\Delta \text{EXCHG}_i\), \(\Delta \text{INT}_i\), and \(\Delta \text{M3}_i\) 
-represent differenced monthly values in the CPI, USD/GBP exchange rate, interest rate, and M3 money supply, respectively.
+represented differenced monthly values in the CPI, USD/GBP exchange rate, interest rate, and M3 money supply, respectively.
 
-The output is a one-dimensional output vector, \(\boldsymbol{\mathrm{y_i}}\) from the dataset $Z$, representing differenced monthly values in the FTSE250.
+The output was a one-dimensional output vector, \(\boldsymbol{\mathrm{y_i}}\) from the dataset $Z$, representing differenced monthly values in the FTSE250.
 
@@
 
-Based on this, the train-test process was conducted:
+The approach outlined in the \hyperref[sec: appendix]{Appendix} was used to conduct the train-test process:
 \begin{enumerate}
     \item The optimal hyperparameters in the covariance function \eqref{eq: gimq} 
-    are computed using GridSearch Cross Validation. These hyperparameters are:
+    were computed using GridSearch Cross Validation. These hyperparameters were:
     \begin{enumerate}
         \item $\alpha$: a regularisation parameter representing a small quantity
-        that is added to the diagonals of the covariance matrix to ensure it is 
+        that was added to the diagonals of the covariance matrix to ensure it was 
         invertible. 
-        \item $\beta$: the power in the covariance function, which controls the shape of the function.
-        \item $\sigma$: a parameter which controls how much of the noise in the data the kernel accounts for. It moderates the risk of overfitting and underfitting.
-        \item Length scale: a parameter that controls the influence that each kernel is to have. Analogous to 
-        the width each radial basis function neuron possesses.
+        \item $\beta$: the power in the covariance function, which controlled the shape of the function.
+        \item $\sigma$: a parameter which controlled how much of the noise in the data the kernel accounted for. It moderated the risk of overfitting and underfitting.
+        \item Length scale: a parameter that controlled the influence that each kernel was to have. Analogous to 
+        the width that each radial basis function neuron possesses.
     \end{enumerate}
     \item Training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
-    and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) are successively passed through the 
+    and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) were successively passed through the 
     covariance function $(x,x')$, to obtain the matrix of the pairwise covariance of 
     all training observations, $K$.
-    \item A $4$-dimensional input space $X$ takes in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
-    calculates the posterior weights distribution. 
-    Each possible weight vector $w$ represents a different function, explicitly 
+    \item A $4$-dimensional input space $X$ took in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
+    calculated the posterior weights distribution. 
+    Each possible weight vector $w$ represented a different function, explicitly 
     based on the radial basis function covariance function, that could fit the data. In the 
-    hidden layer the expectation is then taken, which
-    combines all weights and finds the mean, weighted by the 
-    respective probabilities. This represents the best functional estimate of the price movement. Based on this estimate 
-    the predicted output generated.
+    hidden layer the expectation of this distribution was then taken, which
+    combined all the weights and found the mean, weighted by the 
+    respective probabilities. This represented the best functional estimate of the price movement. Based on this estimate 
+    the predicted output was generated.
     \item The testing process was conducted similarly. The 
-    posterior weight distribution that contains the information of the training phase becomes the new 
+    posterior weight distribution that contained the information of the training phase became the new 
     prior distribution. From this the posterior weight distributions for the test observations \eqref{eq: testw}
@@
 
-A short-term and long-term forecast is conducted using this process. For the short-term forecast a test set of 5 months was used, and for the long-term forecast a test set of 57 months was used.
+A short-term and long-term forecast was conducted using this process. For the short-term forecast a test set of 5 months was used, and for the long-term forecast a test set of 57 months was used.
 
@@
     \centering
+    \label{table: unitroot}
     \caption{Augmented Dickey-Fuller (ADF) Unit Root Test\tablefootnote{All of the variables are tested in natural logarithmic form. The asterisks *, **, ***, and **** denote statistical significance
@@
 
-The optimum lags for the ARDL model were ($FTSE_{250} = 5$, $CPI = 4$, $EXCHG=7$, $INT =7$, $M3=4$) for  The resulting F-statistic of $4.1$ was compared against the critical values
+The optimum lags for the ARDL model were ($FTSE_{250} = 5$, $CPI = 4$, $EXCHG=7$, $INT =7$, $M3=4$). The resulting F-statistic of $4.1$ was compared against the critical values
 given by \citeboth{pesaran2001}, providing evidence of cointegration 
@@
 
-The interest rate has a significantly
+The interest rate had a significantly
 negative short-term impact on the FTSE250 price in the short term. 
@@
 affecting investor sentiment and depressing stock prices. 
-This finding is consistent with the studies by 
+This finding was consistent with the studies by 
 \citeboth{alam2009}, \citeboth{demir2019}, and \citeboth{neifar2023}.
 
-The money supply has a significantly negative short-term effect on the 
+The money supply had a significantly negative short-term effect on the 
 FTSE250 price, as noted by \citeboth{olawale2014}. 
-According to \citeboth{sellin2001}, this supports the Keynesian argument 
+According to \citeboth{sellin2001}, this aligned with the Keynesian argument 
 that an increase in money supply leads to expectations of a future 
@@
 In the short-term, before structural shifts can materialise, the
-weakening of the Pound against the Dollar has a positive, but 
+weakening of the Pound against the Dollar had a positive, but 
 insignificant effect. As the USD has been 
 the global reserve currency for many decades, the depreciation of the Pound 
-against it will lead to increased revenue in Pound terms for FTSE250 
+against it would lead to increased revenue in Pound terms for FTSE250 
 companies with substantial overseas operations. 
@@
 The insignificance of inflation, defined as a positive 
-increase in the CPI, confirms the findings of 
+increase in the CPI, confirmed the findings of 
 \citeboth{gultekin1983}, \citeboth{firth1979} and \citeboth{neifar2023}. 
@@
 was negative and statistically significant ensuring that long-run equilibrium 
-can be achieved. The adjustment speed of $0.1794$ indicates that 
+can be achieved. The adjustment speed of $0.1794$ indicated that 
 roughly $18\%$ of the fluctuation in the previous month's price caused by the 
-macroeconomic variable movements is adjusted back to the long-run equilibrium in the subsequent month. This implies 
+macroeconomic variable movements was adjusted back to the long-run equilibrium in the subsequent month. This implies 
 that it takes approximately 6 months for the FTSE250 to return to its long-run equilibrium following 
@@
 
-The insignificant long-term effect of the money supply can be explained by several reasons.
+The insignificant long-term effect of the money supply can be explained by several factors.
 Firstly, as discussed in \citeboth{sellin2001}, it is possible that the FTSE250 is efficient in processing changes in the 
@@
 The USD/GBP exchange rate significantly positively impacted the FTSE250 price, confirming the findings of 
-\citeboth{wong2022}. That there is a 
-significant relationship is not surprising, because whilst the FTSE250 is a more domestic indicator of the London Stock Exchange than the FTSE 100 Index, 
+\citeboth{wong2022}. That there is a significant relationship is not surprising, because whilst the FTSE250 is a more domestic indicator of the London Stock Exchange than the FTSE 100 Index, 
 at times nearly $50\%$ of its earnings have come from abroad, indicating that 
@@
 The short-term prediction yielded MAE = $0.07$ and 
-$R^2 = 61\%$ for a 5-month time horizon. The results are presented in Figure \ref{fig:shortmonthly}.
+$R^2 = 61\%$ for a 5-month time horizon. The results obtained are presented in Figure \ref{fig:shortmonthly}.
 The significantly better performance of the model in the short-run
@@
 the FTSE250 over the October 1993 – April 2024 period.
-The results suggest a cointegrated long-run
+The results suggested a cointegrated long-run
 relationship between the macroeconomic variables and the stock market index. 
-The long-run coefficients imply that
+The long-run coefficients implied that
 increases in the interest rate and depreciation in the exchange rate

### 2024-08-26 tex/paper.tex
revised writeup

@@
 \usepackage{parskip}
+\usepackage{titlesec}
 \usepackage{float}
@@
 
-\title{Investigating the Impact of Macroeconomic Indicators
+\titleformat{\title}[block]{\normalfont\bfseries}{\thetitle}{1em}{}
 
-on the FTSE 250 Index Using Statistical and Machine Learning Methods}
-\author{Mateen Khan}
-\date{\today}
-
+\title{{Investigating the Impact of Macroeconomic Indicators
+on the FTSE 250 Index Using Statistical and Machine Learning Methods}}
+\author{
+  \textbf{Mateen Khan} \\
+  Mathematics Institute, University of Warwick  \\ 
+  \\
+  \textbf{Supervisor: Dr. Randa Herzallah} \\
+  Mathematics Institute, University of Warwick
+}
 \begin{document}
@@
 Section \ref{sec:lit} provides a brief review of the literature in this problem space and explains the rationale behind the selection of macroeconomic variables.
-Section \ref{sec:meth} presents the methodology. Section \ref{sec: results} discusses the results and Section \ref{sec: conc} concludes
+Section \ref{sec:meth} presents the methodology of this study. Section \ref{sec: results} discusses the results and Section \ref{sec: conc} concludes
 the study.
@@
 
-The study uses monthly FTSE250 price data between October 1992 and March 2024,
+This study used monthly FTSE250 price data between October 1992 and March 2024,
 thus providing 379 observations. The data for the macroeconomic indicators 
 was also captured on a monthly basis in the same time frame, except for the 
-interest rate. This is because the Bank of England's Monetary Policy 
-Committee, 
-which is responsible for setting interest rates, meets on an \textit{ad hoc} 
+interest rate. This is due to the
+\textit{ad hoc} 
 basis\footnote{The MPC typically meets roughly every six weeks, however, in uncertain economic 
-climes it may convene far more frequently.}
-to decide on whether the interest rate should be changed. The sources 
+conditions it may convene far more frequently.} on which the
+Bank of England's Monetary Policy 
+Committee decides to set interest rates. The sources 
 are presented in Table \ref{table: tabl}.
 
-\clearpage
 \begin{table}[h!]
@@
 
-To address the variability in the interest data, Algorithm \ref{alg:interest_rate_adjustment}
-was used to convert the data into a monthly format, ensuring a uniform
-frequency across the variables.
+To address the variable frequency of the interest rate data, Algorithm \ref{alg:interest_rate_adjustment}
+was used to convert the data into a monthly format.
 
@@
 
-The result is a dataset $X$ composed of monthly values for the FTSE250
+The result was a dataset $X$ composed of monthly values for the FTSE250
 price, CPI value, USD/GBP exchange rate, interest rate, and M3 money supply.
@@
 experienced significant fluctuations, highlighting periods of high inflationary pressure such as the onset of the Russia-Ukraine conflict in 2022, and periods of stable, low
-inflation such as the early to late 2010s.
+inflation such as the period between the early to late 2010s.
 
-Interest rates, have varied widely from $0.1\%$ to $8.4\%$, with a standard deviation of $2.52\%$, reflecting the 
-increasing role played by monetary policy in the UK since May 1997 when operational responsibility for setting interest rates was tranferred from the Chancellor of the Exchequer to the Bank of England (\citeboth{king1997changes}).
+Interest rates have ranged from $0.1\%$ to $8.4\%$ with a standard deviation of $2.52\%$, reflecting the 
+increasing role played by monetary policy in the UK since May 1997.
+This was when operational responsibility for setting interest rates was tranferred from the Chancellor of the Exchequer to the Bank of England (\citeboth{king1997changes}).
 The Bank of England was forced to adjust interest rates
@@
 in large part influenced by the various Quantitative Easing (QE) programs implemented by the Bank of England 
-in response to the recessions onset by the 2008 financial crisis and the COVID-19 pandemic.
+in response to the recessions caused by the 2008 financial crisis and the COVID-19 pandemic.
 These programs aimed to inject liquidity into the economy to stimulate economic activity, thus inflating the overall money supply.
@@
 
-The study uses an autoregressive distributed lag (ARDL) model in conjunction with 
+The study used an autoregressive distributed lag (ARDL) model in conjunction with 
 an error correction model (ECM) to examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250
@@
 lagged values and the lagged values of independent variables. 
-The variables are permitted to be of differing stationarity. The stationarity 
+In this model the variables are permitted to be of differing stationarity. The stationarity 
 of the variables was tested using the Augmented Dickey-Fuller 
@@
 variables were calculated using the Schwarz Information Criterion, contained
-within the auto\_ardl function in the ARDL package in R.
+within the \texttt{auto\_ardl} function in the \texttt{ARDL} library in R.
 
@@
 described the speed at which short-run deviations from the long-run equilibrium
-were adjusted. A negative $\gamma$ implies that there existed a correction mechanism that guided the deviation 
+were adjusted. A negative $\gamma$ implied that there existed a correction mechanism that guided the deviation 
 back to the long-run equilibrium. These coefficients, as well 
@@
 
-The \texttt{{bounds\_f\_test}} in 
+The \texttt{{bounds\_f\_test}} function in 
 the \texttt{ARDL} library, based on the bounds test approach 
 outlined in \citeboth{pesaran2001}, was used to calculate the 
-F-statistic of the ARDL model and subsequently test for cointegration.
+F-statistic of the ARDL model and then test for cointegration.
 
@@
 
-To confirm that the model did not display autocorrelation, heteroskedasticity, non-normality, or structural instability, the Box-Ljung, Breusch-Pagan, Shapiro-Wilk, and CUSUM tests were conducted, respectively
+To confirm that the model did not display autocorrelation, heteroskedasticity, {non-normality}, or structural instability, the Box-Ljung, Breusch-Pagan, Shapiro-Wilk, and CUSUM tests were conducted, respectively
 
@@
 range (\citeboth{mendelsohn1993}).
-Also, based on the unit root tests in Table \ref{table: unitroot} the variables in the dataset $X$, described in \ref{section: prep},
+Moreover, based on the unit root tests in Table \ref{table: unitroot} the variables in the dataset $X$, described in \ref{section: prep},
 only achieve stationary at first difference.
 Therefore, all of the 
-variables in the dataset $X$ were normalised using a z-score normalisation and differenced,
+variables in the dataset $X$ were normalised using a z-score normalisation and differenced
 as described in Algorithm \ref{alg:z_score_differencing}.
@@
 
-The inputs into the network were 4-dimensional input vectors, \(\boldsymbol{\mathrm{x_i}}\), from the differenced, normalised dataset $Z$
+The inputs into the network were 4-dimensional input vectors, \(\boldsymbol{\mathrm{x_i}}\), from the differenced and normalised dataset $Z$
 
@@
 
-The output was a one-dimensional output vector, \(\boldsymbol{\mathrm{y_i}}\) from the dataset $Z$, representing differenced monthly values in the FTSE250.
+The output was a one-dimensional output vector, \(\boldsymbol{\mathrm{y_i}}\) from the dataset $Z$, representing differenced, normalised monthly values in the FTSE250.
 
@@
 
-The radial basis function neural network was programmed into Python. The Bayesian kernel regression approach outlined by 
-\citeboth{bishop1995}, using the GaussianProcessRegressor package in Python was employed to model the network. For a 
-detailed description of this construction please refer to the \hyperref[sec: appendix]{Appendix}.
+The radial basis function neural network was programmed in Python. The Bayesian kernel regression approach outlined by 
+\citeboth{bishop1995}, using the \texttt{GaussianProcessRegressor} package in Python was employed to model the network\footnote{See \hyperref[sec: appendix]{Appendix}}.
 
-The approach outlined in the \hyperref[sec: appendix]{Appendix} was used to conduct the train-test process:
+This approach was then used to conduct the train-test process as follows:
 \begin{enumerate}
@@
     prior distribution. From this the posterior weight distributions for the test observations \eqref{eq: testw}
-    were calculated, and then the predictive output was found (\citeboth{rasmussen2006}).
+    were calculated, and then the predictive output was found by taking the expectation (\citeboth{rasmussen2006}).
 
@@
 \end{table}
+\clearpage
 
@@
 
-The interest rate had a significantly
+The interest rate has a significantly
 negative short-term impact on the FTSE250 price in the short term. 
@@
 
-The money supply had a significantly negative short-term effect on the 
+The money supply has a significantly negative short-term effect on the 
 FTSE250 price, as noted by \citeboth{olawale2014}. 
@@
 In the short-term, before structural shifts can materialise, the
-weakening of the Pound against the Dollar had a positive, but 
+weakening of the Pound against the Dollar has a positive, but 
 insignificant effect. As the USD has been 
@@
 The insignificance of inflation, defined as a positive 
-increase in the CPI, confirmed the findings of 
+increase in the CPI, confirms the findings of 
 \citeboth{gultekin1983}, \citeboth{firth1979} and \citeboth{neifar2023}. 
@@
 The error correction term 
-was negative and statistically significant ensuring that long-run equilibrium 
-can be achieved. The adjustment speed of $0.1794$ indicated that 
+is negative and statistically significant ensuring that long-run equilibrium 
+can be achieved. The adjustment speed of $0.1794$ indicates that 
 roughly $18\%$ of the fluctuation in the previous month's price caused by the 
@@
 
-The CPI did not have a significant effect on the FTSE250 price over the 
+The CPI does not have a significant effect on the FTSE250 price over the 
 long-run, lending support to the findings of 
@@
 
-The USD/GBP exchange rate significantly positively impacted the FTSE250 price, confirming the findings of 
+The USD/GBP exchange rate significantly positively impacts the FTSE250 price, confirming the findings of 
 \citeboth{wong2022}. That there is a significant relationship is not surprising, because whilst the FTSE250 is a more domestic indicator of the London Stock Exchange than the FTSE 100 Index, 
@@
 
-The interest rate was also found to have significant negative long-term and short-term effects on the stock market, confirming the findings of \citeboth{alam2009}, \citeboth{demir2019} and \citeboth{neifar2023}
+The interest rate is also found to have significant negative long-term and short-term effects on the stock market, confirming the findings of \citeboth{alam2009}, \citeboth{demir2019} and \citeboth{neifar2023}
 This can be explained by the fact that higher interest 
@@
 
-The short-term prediction yielded MAE = $0.07$ and 
-$R^2 = 61\%$ for a 5-month time horizon. The results obtained are presented in Figure \ref{fig:shortmonthly}.
+The model yields a mean absolute error of MAE = $0.07$ and 
+$R^2 = 61\%$ for the short-term forecast. The results obtained are presented in Figure \ref{fig:shortmonthly}.
 The significantly better performance of the model in the short-run
@@
 \subsubsection{Long-term Forecast}
-The model returned a mean absolute error of MAE $=0.16$ and 
-$R^2 = 9\%$ for the long-term prediction, providing a 
+The model returns a mean absolute error of MAE $=0.16$ and 
+$R^2 = 9\%$ for the long-term forecast, providing a 
 $2.4\%$ increase in the $R^2$ obtained by the ARDL-ECM model. The results are presented 
@@
 forthcoming fluctuations based on policy changes that affect the macroeconomy.
-Within this context, the study examined the 
+Within this context, this study examined the 
 impacts of the interest rate, USD/GBP
@@
 M3 money supply market activity on
-the FTSE250 over the October 1993 – April 2024 period.
+the FTSE250 over the October 1992 – March 2024 period.
 The results suggested a cointegrated long-run
@@
 
-\section{Appendix}
+\newpage
+
+\section*{Appendix}
 \label{sec: appendix}
 
-The GaussianProcessRegressor function in Python's Scikit library is based on the 
+The \texttt{GaussianProcessRegressor} function in Python's \texttt{scikit-learn} library is based on the 
 standard function

### 2024-08-26 tex/paper.tex
revisions

@@
   \textbf{Mateen Khan} \\
-  Mathematics Institute, University of Warwick  \\ 
+  University of Warwick  \\ 
   \\
@@
 This study finds quantitative estimates for the short-term and long-term impacts of the aforementioned macroeconomic indicators on the price of the FTSE 250 Index.
-It also demonstrates the effectiveness of radial basis function neural networks as a short-term forecasting tool for the price of the FTSE 250 Index.
+It also demonstrates the effectiveness of RBFNN as a short-term forecasting tool for the price of the FTSE 250 Index.
     \keywords{macroeconomic indicators, FTSE 250, autoregressive distributed lag model, radial basis function neural network, time-series forecasting}
@@
 makers take the stock market into account when making fiscal policy 
-decisions using a combination of quantitative and qualitative methods. 
+decisions using a combination of quantitative and qualitative methods (\citeboth{DemirgucKunt1996}). 
 These methods such as econometric modelling and forecasting, and empirical 
@@
 macroeconomic indicators
-on the wider economy through the stock market (\citeboth{DemirgucKunt1996}).
+on the wider economy through the stock market.
 
@@
 The significance of this study is to quantify the effects of macroeconomic indicators on the stock market as well as assessing the use of these indicators to forecast stock market prices over the short-term and long-term.
-In addition, the study seeks to understand which macroeconomic indicators are most relevant in capturing the dynamics of a stock market to aid forecasting.
+In addition, this study seeks to understand which macroeconomic indicators are most relevant in capturing the dynamics of a stock market to aid forecasting.
 
@@
 will employ the Autoregressive Distributed Lag (ARDL) model in conjunction with an Error
-Correction Model (ECM). This research also employs a radial basis function neural network to predict the price movements of the FTSE250.
+Correction Model (ECM). This research also employs an RBFNN to predict the price movements of the FTSE250.
 
@@
 Section \ref{sec:meth} presents the methodology of this study. Section \ref{sec: results} discusses the results and Section \ref{sec: conc} concludes
-the study.
+this study.
 
@@
 
-Radial basis function neural networks are a type of artificial 
+RBFNN are a type of artificial 
 neural network that use radial basis functions at each node in the 
 activation layer of the network.
-The advantage of radial basis function neural networks lies in their ability to 
+The advantage of RBFNN lies in their ability to 
 operate effectively in volatile financial time-series 
@@
 
-Radial basis function neural networks have been used considerably for stock market forecasting. \citeboth{cao2004} 
-used a radial basis function network based on the Optimal Partition Algorithm (OPA) to 
-forecast S$\&$P 500 prices. \citeboth{dass2019} use a radial basis function neural network, utilising the Back-propagation 
+RBFNN have been used considerably for stock market forecasting. \citeboth{cao2004} 
+used a RBFNN based on the Optimal Partition Algorithm (OPA) to 
+forecast S$\&$P 500 prices. \citeboth{dass2019} use a RBFNN, utilising the Back-propagation 
 Algorithm (BPA) instead to tune parameters; their results demonstrate 
-that the radial basis function network outperformed a standard feed-forward neural network. 
-\citeboth{ji2014} also found that the radial basis function network was able to predict the value 
+that the RBFNN outperformed a standard feed-forward neural network. 
+\citeboth{ji2014} also found that the RBFNN was able to predict the value 
 of the Shanghai Composite Index on a 45 day time-horizon using the closing price, with a mean prediction error of 
@@
 
-Whilst \citeboth{abotaleb2024}, used quarterly data and a relatively minor economic indicator, the effectiveness of radial basis function neural networks in stock market forecasting using 
+Whilst \citeboth{abotaleb2024}, used quarterly data and a relatively minor economic indicator, the effectiveness of RBFNN in stock market forecasting using 
 major macroeconomic indicators and coarse-grain time-series has not yet been investigated and presents an area for further 
@@
 
-Over the period covered by the dataset, the FTSE250 has 
-shown substantial growth, reflecting the structural transition of the UK economy.
+Over the period covered by the dataset, the FTSE250 
+showed substantial growth, reflecting the structural transition of the UK economy.
 It has also experienced significant volatility, highlighting significant market shifts driven by 
@@
 
-The USD/GBP exchange rate has exhibited moderate variability, exhibiting the stable relations between 
+The USD/GBP exchange rate exhibited moderate variability, highlighting the stable relations between 
 the two countries. Nonetheless, geopolitical shocks such as the 2008 financial crisis and Brexit both greatly affected the value of the Pound, 
-which accounts for the large range.
+which accounted for the large range.
 
-The CPI, with an average of 88.36, has 
-experienced significant fluctuations, highlighting periods of high inflationary pressure such as the onset of the Russia-Ukraine conflict in 2022, and periods of stable, low
+The CPI, with an average of 88.36, 
+experienced significant fluctuations, which highlighted periods of high inflationary pressure such as the onset of the Russia-Ukraine conflict in 2022, and periods of stable, low
 inflation such as the period between the early to late 2010s.
 
-Interest rates have ranged from $0.1\%$ to $8.4\%$ with a standard deviation of $2.52\%$, reflecting the 
+Interest rates ranged from $0.1\%$ to $8.4\%$ with a standard deviation of $2.52\%$, which reflected the 
 increasing role played by monetary policy in the UK since May 1997.
-This was when operational responsibility for setting interest rates was tranferred from the Chancellor of the Exchequer to the Bank of England (\citeboth{king1997changes}).
+It was at this time when operational responsibility for setting interest rates was tranferred from the Chancellor of the Exchequer to the Bank of England (\citeboth{king1997changes}).
 The Bank of England was forced to adjust interest rates 
-several times in response to worsening economic conditions, examples being the steep rate cuts during the 2008 financial crisis to stimulate the economy and the gradual rate hikes in response to 
+several times in response to worsening economic conditions. For example, 
+it made steep rate cuts during the 2008 financial crisis
+to stimulate the economy and a series gradual rate hikes in response to 
 post-COVID inflationary pressures.
 
-The M3 money supply has shown substantial variability, ranging from 494,181 to 3,704,873, 
+The M3 money supply showed substantial variability, ranging from 494,181 to 3,704,873, 
 in large part influenced by the various Quantitative Easing (QE) programs implemented by the Bank of England 
@@
 
-The study used an autoregressive distributed lag (ARDL) model in conjunction with 
+This study used an autoregressive distributed lag (ARDL) model in conjunction with 
 an error correction model (ECM) to examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250
-and to quantify these relationships. Prior to constructing the model a natural logarithmic transform was applied
-on the variables. This was done so that the coefficients can be interpreted easily as percentage changes.
+and to quantify these relationships. Prior to constructing the model,
+a natural logarithmic transform was applied
+on the variables to enable the coefficients to be interpreted 
+easily as percentage changes.
 
@@
 The \texttt{{bounds\_f\_test}} function in 
-the \texttt{ARDL} library, based on the bounds test approach 
+the \texttt{ARDL} library, which is based on the bounds test approach 
 outlined in \citeboth{pesaran2001}, was used to calculate the 
-F-statistic of the ARDL model and then test for cointegration.
+F-statistic and then test for cointegration. The hypotheses were:
 
@@
 
-for all $i = 1,2,3,4$.
+where $\beta_{1\leq i\leq4}$ represented the long-term coefficients in 
+the ARDL model.
 
-To confirm that the model did not display autocorrelation, heteroskedasticity, {non-normality}, or structural instability, the Box-Ljung, Breusch-Pagan, Shapiro-Wilk, and CUSUM tests were conducted, respectively
+To confirm that the model did not display autocorrelation, heteroskedasticity, {non-normality}, or structural instability, the Box-Ljung, Breusch-Pagan, Shapiro-Wilk, and CUSUM tests were conducted respectively
 
-\subsection{RBF Neural Network}
+\subsection{RBFNN}
 
-This study employed radial basis function neural networks, 
+This study employed RBFNN, 
 a type of artificial neural network that utilise radial basis functions 
 at each node in the activation layer. 
-This was because radial basis function neural networks 
+The primary reason for this is that they
 are well-suited for identifying complex, non-linear relationships 
 between time-series variables, making them particularly valuable in 
-forecasting tasks. In addition, radial basis function neural networks
+forecasting tasks. In addition, RBFNN
 have been shown to be effective in operating on small sample sizes, 
-which was critical for the data constraints of this study.
+which was a significant issue in this study.
 
 \subsubsection{Network Architecture}
-The exact interpolation problem, which forms the motivation for radial basis function neural networks, 
+The exact interpolation problem, which forms the motivation for RBFNN, 
 involves mapping a $d$-dimensional input space $\boldsymbol{\mathrm{X}}$ to
@@
 By introducing several modifications to the exact interpolation 
-procedure, the radial basis function neural network
-model is obtained, as presented in Figure \ref{fig:rbf_nn_architecture}. 
+procedure, the RBFNN
+model as shown in Figure \ref{fig:rbf_nn_architecture} is obtained.
+
 The network mapping takes the following form:
@@
     \end{tikzpicture}
-    \caption{Radial basis function neural network}
+    \caption{RBFNN Architecture}
     \label{fig:rbf_nn_architecture}
@@
 
-The radial basis function neural network was programmed in Python. The Bayesian kernel regression approach outlined by 
-\citeboth{bishop1995}, using the \texttt{GaussianProcessRegressor} package in Python was employed to model the network\footnote{See \hyperref[sec: appendix]{Appendix}}.
+The RBFNN was programmed in Python. The Bayesian kernel regression approach outlined by 
+\citeboth{bishop1995}, using the \texttt{GaussianProcessRegressor} 
+package in the \texttt{sci-kit} library was employed to construct the
+network\footnote{See \hyperref[sec: appendix]{Appendix}}.
 
-This approach was then used to conduct the train-test process as follows:
+\subsubsection{Network Training and Testing}
+
+The Bayesian kernel regression 
+framework was then used to conduct the train-test process as follows:
 \begin{enumerate}
@@
     and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) were successively passed through the 
-    covariance function $(x,x')$, to obtain the matrix of the pairwise covariance of 
+    covariance function $(x,x')$ to obtain the matrix of the pairwise covariance of 
     all training observations, $K$.
@@
 M3 money supply market activity on
-the FTSE250 over the October 1992 – March 2024 period.
+the FTSE250 between October 1992 and March 2024.
 The results suggested a cointegrated long-run
@@
 Policymakers dealing with forecasting the UK stock markets
-should consider the factors in the study, and to raise the stock
+should consider the factors in this study, and in order to raise the stock
 market's value, a lower interest rate regime should be adopted. This will 
@@
 
-The study also investigated the efficacy of machine learning methods in 
-the field of stock-econometric forecasting. The radial basis function network performed well in
+This study also investigated the efficacy of machine learning methods in 
+the field of stock-econometric forecasting. The RBFNN performed well in
 short-term forecasting of the FTSE250, and offers investors with an accurate forecasting tool, 
@@
 Brownian bridge algorithm to generate daily macroeconomic data. This would allow 
-the testing of the radial basis function neural network against other neural networks 
+the testing of the RBFNN against other neural networks 
 such as the LSTM. Other machine learning methods suited for 
@@
 
-\section*{Appendix}
+\section{Appendix}
 \label{sec: appendix}
@@
 \begin{align} 
-    f(x) = x^{T}w, \quad{y = f(x) + \varepsilon}
+    f(\boldsymbol{\mathrm{x}}) = \boldsymbol{\mathrm{x}}^{T}\boldsymbol{\mathrm{w}}, \quad{\boldsymbol{\mathrm{y}} = f(\boldsymbol{\mathrm{x}}) + \varepsilon}
 \end{align}
 
-where $y$ differs from the function values $f(x)$ by noise which is assumed
+where $\boldsymbol{\mathrm{y}}$ differs from the function values $f(x)$ by noise which is assumed
 to follow an iid Gaussian distribution 
@@
 \begin{align}
-    p(y|X,w) = \prod_{i=1}^{n} p(y_i|x_i ,w) = \mathcal{N}\bigl(X^T, \sigma_{n}^2 I\bigr)
+    p(\boldsymbol{\mathrm{y}}|X,\boldsymbol{\mathrm{w}}) = \prod_{i=1}^{n} p(\boldsymbol{\mathrm{y}}_i|\boldsymbol{\mathrm{x}}_i ,\boldsymbol{\mathrm{w}}) = \mathcal{N}\bigl(X^T, \sigma_{n}^2 I\bigr)
 \end{align}
 
-The weights \( w \) represent the parameters that define the relationship between the inputs and outputs. 
+The weights \( \boldsymbol{\mathrm{w}} \) represent the parameters that define the relationship between the inputs and outputs. 
 They are treated as random variables defined by the Gaussian distribution \eqref{eq: weight}.
@@
 \begin{align}
-    k(x,x') = \biggl( ||x-x'|| + \sigma^2\biggr)^{-\beta} \label{eq: gimq}
+    k(\boldsymbol{\mathrm{x}},\boldsymbol{\mathrm{x}}') = \biggl( ||\boldsymbol{\mathrm{x}}-\boldsymbol{\mathrm{x}}'|| + \sigma^2\biggr)^{-\beta} \label{eq: gimq}
 \end{align}
 and used to construct the covariance matrix $K$, which along with 
-a mean function $m(x)$ forms the distribution of the the weights
+a mean function $m(\boldsymbol{\mathrm{x}})$ forms the distribution of the the weights
 
 \begin{align}
-    w \sim \mathcal{N}\bigl(m(x), K\bigr) \label{eq: weight}
+    \boldsymbol{\mathrm{w}} \sim \mathcal{N}\bigl(m(\boldsymbol{\mathrm{x}}), K\bigr) \label{eq: weight}
 \end{align}
 
-The mean function $m(x)$ acts as a bias, 
+The mean function $m(\boldsymbol{\mathrm{x}})$ acts as a bias, 
 providing a baseline expectation of the function values. 
@@
 \begin{align}
-    p(w|y, X) = \frac{p(y|X,w) p(w)}{p(y|X)} \label{eq: post}
+    p(\boldsymbol{\mathrm{w}}|\boldsymbol{\mathrm{y}}, X) = \frac{p(\boldsymbol{\mathrm{y}}|X,\boldsymbol{\mathrm{w}}) p(\boldsymbol{\mathrm{w}})}{p(\boldsymbol{\mathrm{y}}|X)} \label{eq: post}
 \end{align}
@@
 \begin{align}
-    p(w|y, X) \sim \mathcal{N}\bigl(\frac{A^{-1}Xy}{\sigma_{n}^{2}}, A^{-1}\bigr) \label{eq:nine}
+    p(\boldsymbol{\mathrm{w}}|\boldsymbol{\mathrm{y}}, X) \sim \mathcal{N}\bigl(\frac{A^{-1}X\boldsymbol{\mathrm{y}}}{\sigma_{n}^{2}}, A^{-1}\bigr) \label{eq:nine}
 \end{align}

### 2025-07-17 latex/paper.tex
refactored ardl-ecm code

--- 2024-08-26 tex/paper.tex
@@
 macroeconomic indicators
-on the wider economy through the stock market.
+on the wider economy.
 
@@
 outlined in \citeboth{pesaran2001}, was used to calculate the 
-F-statistic and then test for cointegration. The hypotheses were:
+F-statistic and subsequently tested for cointegration. The hypotheses were:
 
@@
 The Bayesian kernel regression 
-framework was then used to conduct the train-test process as follows:
+framework was used to conduct the train-test process as follows:
 \begin{enumerate}
@@
 Policymakers dealing with forecasting the UK stock markets
-should consider the factors in this study, and in order to raise the stock
-market's value, a lower interest rate regime should be adopted. This will 
+should consider the factors in this study and in order to raise the stock
+market's value they ought to adopt a lower interest rate regime. This will 
 stimulate consumer activity, reduce operating costs for firms, and make 
@@
 
+\addcontentsline{toc}{section}{References}
 \bibliographystyle{plainnat}

### 2025-07-17 latex/short_term_monthly.tex
refactored ardl-ecm code

% Created by tikzDevice version 0.12.6 on 2024-08-15 14:40:36
% !TEX encoding = UTF-8 Unicode
\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,289.08);
\begin{scope}
\path[clip] ( 34.16, 76.79) rectangle (500.39,266.42);
\definecolor{drawColor}{gray}{0.90}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16, 87.71) --
	(500.39, 87.71);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,147.06) --
	(500.39,147.06);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,206.41) --
	(500.39,206.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,265.76) --
	(500.39,265.76);

\path[draw=drawColor,line width= 0.3pt,line join=round] (113.15, 76.79) --
	(113.15,266.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (218.23, 76.79) --
	(218.23,266.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (323.32, 76.79) --
	(323.32,266.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (430.16, 76.79) --
	(430.16,266.42);
\definecolor{drawColor}{gray}{0.80}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,117.38) --
	(500.39,117.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,176.73) --
	(500.39,176.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,236.08) --
	(500.39,236.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 58.85, 76.79) --
	( 58.85,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (167.44, 76.79) --
	(167.44,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (269.02, 76.79) --
	(269.02,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (377.61, 76.79) --
	(377.61,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (482.70, 76.79) --
	(482.70,266.42);
\definecolor{drawColor}{RGB}{0,255,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 55.35,257.80) --
	(163.94, 85.41) --
	(265.52, 88.16) --
	(374.11,197.41) --
	(479.20,125.17);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 55.35,229.43) --
	(163.94,102.42) --
	(265.52, 90.46) --
	(374.11,144.72) --
	(479.20,194.60);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 1.1pt,line join=round,line cap=round] ( 34.16, 76.79) rectangle (500.39,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,114.35) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,173.70) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,233.05) {0.2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 63.14, 67.55) {Jan};

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (171.73, 67.55) {Feb};

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (273.31, 67.55) {Mar};

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (381.90, 67.55) {Apr};

\node[text=drawColor,rotate= 45.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (486.99, 67.55) {May};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (267.27, 44.09) {$t$ = time};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,171.61) {$\delta$ = FTSE250 price change};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.75, 18.23) -- (233.31, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,255,0}

\path[draw=drawColor,line width= 1.1pt,line join=round] (284.01, 18.23) -- (295.57, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (240.26, 15.20) {Predicted};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (302.51, 15.20) {Real};
\end{scope}
\end{tikzpicture}

### 2025-07-17 ardl/adrl-bounds.R

@@ -1,125 +1,0 @@
-library(tseries)
-library(vars)
-library(lmtest)
-library(ARDL)
-library(dynlm)
-
-columns_to_transform <- c("CPI", "INT", "UNEM", "EXCHG", "FTSE.250")
-data[columns_to_transform] <- lapply(data[columns_to_transform], function(x) log(x))
-
-adf.test(data$UNEM)
-dunem = diff(data$UNEM)
-adf.test(dunem)
-
-adf.test(data$INT)
-dint = diff(data$INT)
-adf.test(dint)
-
-adf.test(data$EXCHG)
-dexchg = diff(data$EXCHG)
-adf.test(dexchg)
-
-
-adf.test(data$CPI)
-dcpi = diff(data$CPI)
-adf.test(dcpi)
-
-max_lags <- 9
-results <- list()
-
-for (p in 1:max_lags) {
-  model <- ardl(FTSE.250 ~ UNEM + INT + CPI + EXCHG, data = data, order = c(p,p,p,p,p))
-  
-  aic <- AIC(model)
-  sc <- BIC(model)
-  
-  bg_test <- bgtest(model, order = p)
-  
-  results[[p]] <- list(
-    AIC = aic,
-    SC = sc,
-    Serial_Correlation_p_value = bg_test$p.value
-  )
-}
-
-results_df <- do.call(rbind, lapply(results, as.data.frame))
-print(results_df)
-
-
-
-VARselect(data$FTSE.250)
-VARselect(data$CPI)
-VARselect(data$UNEM)
-VARselect(data$INT)
-VARselect(data$EXCHG)
-
-model3 <- ardl(FTSE.250 ~ UNEM + INT + CPI + EXCHG, data = data, order = c(4,7,7,7,7))
-summary(model)
-
-test3 <- bounds_f_test(model3, case=3)
-test3
-
-model4 <- ardl(FTSE.250 ~ INT + CPI + EXCHG, data = data, order = c(4,7,7,7))
-summary(model)
-
-test4 <- bounds_f_test(model4, case=2)
-test4
-
-install.packages("dLagM")
-library(dLagM)
-
-coefs <- coef(model4)
-
-
-
-# LR_coeff = sum(short-run coefficients of the variable) / (1 - sum(autoregressive coefficients))
-LR_INT <- sum(coefs[grepl("INT", names(coefs))]) / (1 - sum(coefs[grepl("FTSE.250", names(coefs))]))
[…]

### 2025-07-17 ardl/ardl.R

@@ -1,122 +1,0 @@
-library(tseries)
-library(vars)
-library(lmtest)
-library(ARDL)
-library(dynlm)
-library(dplyr)
-library(dLagM)
-library(strucchange)
-library(zoo)
-
-
-columns_to_transform <- c("CPI", "INT", "UNEM", "EXCHG", "M2", "M3", "FTSE.250")
-monthly.data.m3[columns_to_transform] <- lapply(monthly.data.m3[columns_to_transform], function(x) log(x))
-
-adf.test(data$UNEM)
-dunem <- diff(data$UNEM)
-adf.test(dunem)
-
-adf.test(data$INT)
-dint <- diff(data$INT)
-adf.test(dint)
-
-adf.test(data$EXCHG)
-dexchg <- diff(data$EXCHG)
-adf.test(dexchg)
-
-
-adf.test(data$CPI)
-dcpi <- diff(data$CPI)
-adf.test(dcpi)
-
-adf.test(data$M3)
-dm3 <- diff(data$M3)
-adf.test(dm3)
-
-adf.test(data$FTSE.250)
-dftse <- diff(data$FTSE.250)
-adf.test(dftse)
-
-max_lags <- 9
-results <- list()
-
-for (p in 1:max_lags) {
-  model <- ardl(FTSE.250 ~ INT + CPI + EXCHG + M3, data = monthly.data.m3, order = c(p, p, p, p, p))
-
-  aic <- AIC(model)
-  sc <- BIC(model)
-
-  bg_test <- bgtest(model, order = p)
-
-  results[[p]] <- list(
-    AIC = aic,
-    SC = sc,
-    Serial_Correlation_p_value = bg_test$p.value
-  )
-}
-
-
-VARselect(data$FTSE.250)
-VARselect(data$CPI)
-VARselect(data$UNEM)
-VARselect(data$INT)
-VARselect(monthly.data.m3$M3)
-VARselect(data$EXCHG)
-
-
-model5 <- ardl(FTSE.250 ~ INT + CPI + EXCHG + M3, data = monthly.data.m3, order = c(5, 4, 7, 7, 4))
-summary(model5)
-
-
-test5 <- bounds_f_test(model5, case = 2)
-test5
-
-
-coefs <- coef(model5)
-
-
-LR_INT <- sum(coefs[grepl("INT", names(coefs))]) / (1 - sum(coefs[grepl("FTSE.250", names(coefs))]))
[…]

### 2025-07-17 r-scripts/ardl-alt.R

@@ -1,0 +1,123 @@
+# Load required packages
+library(tseries)      # For ADF test
+library(vars)         # For VAR selection criteria
+library(lmtest)       # For Breusch-Godfrey test
+library(ARDL)         # For ARDL bounds testing
+library(dynlm)        # For dynamic linear models (ECM)
+library(dLagM)        # For alternative ARDL implementation
+library(dplyr)        # For data wrangling
+library(strucchange)  # For structural break tests
+library(zoo)          # For time series manipulation
+library(memochange)   # For CUSUM testing
+
+# Log-transform relevant macroeconomic variables
+columns_to_log <- c("CPI", "INT", "UNEM", "EXCHG", "M2", "M3", "FTSE.250")
+monthly.data.m3[columns_to_log] <- lapply(../data/monthly.data.m3[columns_to_log], log)
+
+# Perform ADF tests to check stationarity
+macro_vars <- c("UNEM", "INT", "EXCHG", "CPI", "M3", "FTSE.250")
+for (var in macro_vars) {
+  cat("ADF test for", var, "\n")
+  print(adf.test(monthly.data.m3[[var]]))
+  
+  # Test first difference if non-stationary
+  diff_var <- diff(monthly.data.m3[[var]])
+  cat("ADF test for Δ", var, "\n")
+  print(adf.test(diff_var))
+}
+
+# Select maximum lag length for ARDL order search
+max_lags <- 9
+results <- list()
+
+# Grid search to choose ARDL order by AIC, BIC and BG serial correlation test
+for (p in 1:max_lags) {
+  model <- ardl(FTSE.250 ~ INT + CPI + EXCHG + M3, data = monthly.data.m3, order = c(p, p, p, p, p))
+  bg_test <- bgtest(model, order = p)
+  
+  results[[p]] <- list(
+    AIC = AIC(model),
+    BIC = BIC(model),
+    BG_pvalue = bg_test$p.value
+  )
+}
+# Optional: View results as a data frame
+# do.call(rbind, results)
+
+# Select optimal ARDL order manually or based on information criteria
+# Run VARselect for information purposes
+VARselect(monthly.data.m3$FTSE.250)
+VARselect(monthly.data.m3$CPI)
[…]

### 2025-07-17 r-scripts/ardl-ecm.R

@@ -1,0 +1,117 @@
+# Load required libraries
+library(tseries)       # ADF test
+library(vars)          # VAR selection
+library(lmtest)        # Breusch-Godfrey and Breusch-Pagan tests
+library(ARDL)          # ARDL estimation and bounds testing
+library(dynlm)         # Dynamic linear models
+library(dLagM)         # Lag models
+library(dplyr)         # Data wrangling
+library(strucchange)   # Structural stability
+library(zoo)           # Rolling residuals
+library(here)          # Robust file path handling
+
+# Load data from parent directory's data folder
+data <- read.csv(here("..", "data", "monthly_data.csv"))
+
+# Log-transform selected macroeconomic variables
+columns_to_log <- c("CPI", "INT", "UNEM", "EXCHG", "FTSE.250")
+data[columns_to_log] <- lapply(data[columns_to_log], log)
+
+# Perform ADF unit root tests for level and first differences
+for (var in columns_to_log) {
+  cat("ADF test for:", var, "\n")
+  print(adf.test(data[[var]]))
+  
+  diff_var <- diff(data[[var]])
+  cat("ADF test for Δ", var, "\n")
+  print(adf.test(diff_var))
+}
+
+# ARDL model lag order selection (grid search up to max_lags)
+max_lags <- 9
+results <- list()
+
+for (p in 1:max_lags) {
+  model <- ardl(FTSE.250 ~ UNEM + INT + CPI + EXCHG, data = data, order = c(p, p, p, p, p))
+  
+  results[[p]] <- list(
+    AIC = AIC(model),
+    BIC = BIC(model),
+    BG_pval = bgtest(model, order = p)$p.value
+  )
+}
+
+results_df <- do.call(rbind, lapply(results, as.data.frame))
+print(results_df)
+
+# VAR selection (information only)
+vars_to_check <- c("FTSE.250", "CPI", "UNEM", "INT", "EXCHG")
+
+# Loop through variables and apply VARselect individually
+for (var_name in vars_to_check) {
+  cat("\n===== VARselect for", var_name, "=====\n")
+  var_data <- data[[var_name]]
+  
+  # VARselect expects a multivariate time series; we convert to a ts object
[…]

### 2025-07-17 results/rbf-nn-plots.R
refactor rbf-nn code

@@ -1,0 +1,47 @@
+library(ggplot2)
+
+# --- Load Data ---
+rbf_predictions_dataframe <- read.csv("../data/rbf-predictions.csv")
+long_data <- read.csv("../data/long_data.csv")
+
+# Convert date columns
+rbf_predictions_dataframe$daily.date <- as.Date(rbf_predictions_dataframe$daily.date)
+long_data$Date <- as.Date(long_data$Date)
+
+# --- Plot 1: Daily Price Changes (Short-Term Model Output) ---
+ggplot(rbf_predictions_dataframe, aes(x = daily.date)) +
+  geom_line(aes(y = DNFTSE, color = "Real"), linewidth = 1) +
+  geom_line(aes(y = daily.predicted.price, color = "Predicted"), linewidth = 0.5) +
+  scale_color_manual(values = c("Real" = "green", "Predicted" = "blue")) +
+  labs(
+    title = "Daily FTSE 250: Real vs Predicted Changes",
+    x = "Date",
+    y = "Price Change",
+    color = ""
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    legend.position = "bottom",
+    panel.grid.major = element_line(colour = "grey80"),
+    panel.grid.minor = element_line(colour = "grey90")
+  )
+
+# --- Plot 2: Monthly Trend (Long-Term Model Output) ---
+ggplot(long_data, aes(x = Date)) +
+  geom_line(aes(y = Real, color = "Real"), linewidth = 1) +
+  geom_line(aes(y = Predicted, color = "Predicted"), linewidth = 0.5) +
+  scale_color_manual(values = c("Real" = "green", "Predicted" = "blue")) +
+  labs(
+    title = "Monthly FTSE 250: Real vs Predicted Trend",
+    x = "Date",
+    y = "Price Change",
+    color = ""
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    legend.position = "bottom",
+    panel.grid.major = element_line(colour = "grey80"),
+    panel.grid.minor = element_line(colour = "grey90")
+  )

### 2025-07-17 results/results-plots.R

@@ -1,37 +1,0 @@
-library(ggplot2)
-updated_dataframe$daily.date <- as.Date(updated_dataframe$daily.date)
-
-ggplot(updated_dataframe, aes(x = daily.date)) +
-  geom_line(aes(y = DNFTSE, color = "Real"), linewidth = 1) +    # Plot DNFTSE first
-  geom_line(aes(y = daily.predicted.price, color = "Predicted"), linewidth = 0.5) +    # Plot Predicted second (on top)
-  scale_color_manual(values = c("Predicted" = "blue", "Real" = "green")) +
-  labs(title = "",
-       x = "",
-       y = "Price changes",
-       color = "") +
-  theme_minimal() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    legend.position = "bottom",
-    panel.grid.major = element_line(color = "grey80"),
-    panel.grid.minor = element_line(color = "grey90")
-  )
-
-long_data$Date <- as.Date(long_data$Date)
-
-# Create the plot
-ggplot(long_data, aes(x = Date)) +
-  geom_line(aes(y = Real, color = "Real"), linewidth = 1) +    # Plot Real first
-  geom_line(aes(y = Predicted, color = "Predicted"), linewidth = 0.5) +    # Plot Predicted second (on top)
-  scale_color_manual(values = c("Predicted" = "blue", "Real" = "green")) +
-  labs(title = "",
-       x = "Date",
-       y = "Price Changes",
-       color = "") +
-  theme_minimal() +
-  theme(
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    legend.position = "bottom",
-    panel.grid.major = element_line(color = "grey80"),
-    panel.grid.minor = element_line(color = "grey90")
-  )
